---
title: "Breast Cancer Risk Prediction in Indian Population"
author: "Parichoy Pal Choudhury"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

## Install the required packages and libraries
```{r, message = FALSE}
# if (!require("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# 
# BiocManager::install("iCARE")
library(readstata13)
library(plotrix)
library(gdata)
library(openxlsx)
library(iCARE)
```

## Set the working directory to the location of the data and read the data
```{r}

setwd("/Users/anaghchattopadhyay/Documents/Nilanjan\ Project/TMC_BCProject/Data")

data = read.dta13("Questionnaire_28.04.2017.dta",
                  generate.factors = TRUE, nonint.factors = TRUE)
```

## Categories for age at first full term pregnancy
```{r}
data$ageatfirstfulltermpreg_cat = data$ageatfirstfulltermpreg
data$ageatfirstfulltermpreg_cat[which(data$ageatfirstfulltermpreg < 20)] = 1
data$ageatfirstfulltermpreg_cat[which(data$ageatfirstfulltermpreg == 20 
                                      | data$ageatfirstfulltermpreg == 21)] = 2
data$ageatfirstfulltermpreg_cat[which(data$ageatfirstfulltermpreg == 22 
                                      | data$ageatfirstfulltermpreg == 23)] = 3
data$ageatfirstfulltermpreg_cat[which(data$ageatfirstfulltermpreg == 24 
                                      | data$ageatfirstfulltermpreg == 25)] = 4
data$ageatfirstfulltermpreg_cat[which(data$ageatfirstfulltermpreg >= 26)] = 5
```

## Categories for interval between menarche and full term pregnancy
```{r}
data$interval_menarche_firstbirth = data$ageatfirstfulltermpreg - 
  data$whtagestartmensurating

data$interval_menarche_firstbirth_cat = data$interval_menarche_firstbirth
data$interval_menarche_firstbirth_cat[data$interval_menarche_firstbirth < 10] = 0
data$interval_menarche_firstbirth_cat[data$interval_menarche_firstbirth >= 10] = 1

```

## Categories for number of full term pregnancies
```{r}
data$fulltermpreg_cat = data$fulltermpreg
data$fulltermpreg_cat[data$fulltermpreg >= 4] = 4
```

## Categories for number of induced abortions
```{r}
data$totalabortion_cat = data$totalabortion
data$totalabortion_cat[data$totalabortion >= 2] = 2
```

## Categories for number of spontaneous abortions
```{r}
data$totalmiscarriage_cat = data$totalmiscarriage
data$totalmiscarriage_cat[data$totalmiscarriage >= 2] = 2
```

## Categories for duration of breastfeeding
```{r}
data$mean_brfeed_duration_cat = data$mean_brfeed_duration
data$mean_brfeed_duration_cat[data$mean_brfeed_duration < 12] = 1
data$mean_brfeed_duration_cat[data$mean_brfeed_duration >= 12 
                              & data$mean_brfeed_duration < 24] = 2
data$mean_brfeed_duration_cat[data$mean_brfeed_duration >= 24 
                              & data$mean_brfeed_duration < 36] = 3
data$mean_brfeed_duration_cat[data$mean_brfeed_duration >= 36] = 4
```

## Categories for duration of diabetes
```{r}
data$diabetic_years_cat = data$diabetic_years
data$diabetic_years_cat[data$diabetes_YN == 0] = 0
data$diabetic_years_cat[data$diabetic_years <= 5] = 1
data$diabetic_years_cat[data$diabetic_years >= 6 
                        & data$diabetic_years <= 10] = 2
data$diabetic_years_cat[data$diabetic_years >= 11] = 3
```

## Collapsing the <=18.5 and 18.5-24.9 categories for BMI
```{r}
data$bmicat[data$bmicat == 0 | data$bmicat == 1] = 1
```

## Subset the data to pre-menopausal women
```{r}
data_premeno = data[which(as.character(data$menopause) == "premenopausal"),]
```

## Number of cases and controls
```{r}
table(data_premeno$br)
```

## Number and proportion of risk factors among controls and cases

## Age at first full term pregnancy
```{r}
table(data_premeno$ageatfirstfulltermpreg_cat
      [as.character(data_premeno$br) == "Control"])
sum(table(data_premeno$ageatfirstfulltermpreg_cat
          [as.character(data_premeno$br) == "Control"]))
prop.table(table(data_premeno$ageatfirstfulltermpreg_cat
                 [as.character(data_premeno$br) == "Control"]))

table(data_premeno$ageatfirstfulltermpreg_cat
      [as.character(data_premeno$br) == "Breast Cancer Case"])
sum(table(data_premeno$ageatfirstfulltermpreg_cat
          [as.character(data_premeno$br) == "Breast Cancer Case"]))
prop.table(table(data_premeno$ageatfirstfulltermpreg_cat
                 [as.character(data_premeno$br) == "Breast Cancer Case"]))
```

## Interval between menarche and first full term pregnancy
```{r}
table(data_premeno$interval_menarche_firstbirth_cat
      [as.character(data_premeno$br) == "Control"])
sum(table(data_premeno$interval_menarche_firstbirth_cat
          [as.character(data_premeno$br) == "Control"]))
prop.table(table(data_premeno$interval_menarche_firstbirth_cat
                 [as.character(data_premeno$br) == "Control"]))

table(data_premeno$interval_menarche_firstbirth_cat
      [as.character(data_premeno$br) == "Breast Cancer Case"])
sum(table(data_premeno$interval_menarche_firstbirth_cat
          [as.character(data_premeno$br) == "Breast Cancer Case"]))
prop.table(table(data_premeno$interval_menarche_firstbirth_cat
                 [as.character(data_premeno$br) == "Breast Cancer Case"]))
```

## Number of full term pregnancies
```{r}
table(data_premeno$fulltermpreg_cat
      [as.character(data_premeno$br) == "Control"])
sum(table(data_premeno$fulltermpreg_cat
          [as.character(data_premeno$br) == "Control"]))
prop.table(table(data_premeno$fulltermpreg_cat
                 [as.character(data_premeno$br) == "Control"]))

table(data_premeno$fulltermpreg_cat
      [as.character(data_premeno$br) == "Breast Cancer Case"])
sum(table(data_premeno$fulltermpreg_cat
          [as.character(data_premeno$br) == "Breast Cancer Case"]))
prop.table(table(data_premeno$fulltermpreg_cat
                 [as.character(data_premeno$br) == "Breast Cancer Case"]))
```

## OC Use
```{r}
table(data_premeno$oc
      [as.character(data_premeno$br) == "Control"])
sum(table(data_premeno$oc
          [as.character(data_premeno$br) == "Control"]))
prop.table(table(data_premeno$oc
                 [as.character(data_premeno$br) == "Control"]))

table(data_premeno$oc
      [as.character(data_premeno$br) == "Breast Cancer Case"])
sum(table(data_premeno$oc
          [as.character(data_premeno$br) == "Breast Cancer Case"]))
prop.table(table(data_premeno$oc
                 [as.character(data_premeno$br) == "Breast Cancer Case"]))
```

## Duration of breast-feeding
```{r}
table(data_premeno$mean_brfeed_duration_cat
      [as.character(data_premeno$br) == "Control"])
sum(table(data_premeno$mean_brfeed_duration_cat
          [as.character(data_premeno$br) == "Control"]))
prop.table(table(data_premeno$mean_brfeed_duration_cat
                 [as.character(data_premeno$br) == "Control"]))

table(data_premeno$mean_brfeed_duration_cat
      [as.character(data_premeno$br) == "Breast Cancer Case"])
sum(table(data_premeno$mean_brfeed_duration_cat
          [as.character(data_premeno$br) == "Breast Cancer Case"]))
prop.table(table(data_premeno$mean_brfeed_duration_cat
                 [as.character(data_premeno$br) == "Breast Cancer Case"]))
```

## Number of induced abortions
```{r}
table(data_premeno$totalabortion_cat
      [as.character(data_premeno$br) == "Control"])
sum(table(data_premeno$totalabortion_cat
          [as.character(data_premeno$br) == "Control"]))
prop.table(table(data_premeno$totalabortion_cat
                 [as.character(data_premeno$br) == "Control"]))

table(data_premeno$totalabortion_cat
      [as.character(data_premeno$br) == "Breast Cancer Case"])
sum(table(data_premeno$totalabortion_cat
          [as.character(data_premeno$br) == "Breast Cancer Case"]))
prop.table(table(data_premeno$totalabortion_cat
                 [as.character(data_premeno$br) == "Breast Cancer Case"]))
```

## Number of spontaneous abortions
```{r}
table(data_premeno$totalmiscarriage_cat
      [as.character(data_premeno$br) == "Control"])
sum(table(data_premeno$totalmiscarriage_cat
          [as.character(data_premeno$br) == "Control"]))
prop.table(table(data_premeno$totalmiscarriage_cat
                 [as.character(data_premeno$br) == "Control"]))

table(data_premeno$totalmiscarriage_cat
      [as.character(data_premeno$br) == "Breast Cancer Case"])
sum(table(data_premeno$totalmiscarriage_cat
          [as.character(data_premeno$br) == "Breast Cancer Case"]))
prop.table(table(data_premeno$totalmiscarriage_cat
                 [as.character(data_premeno$br) == "Breast Cancer Case"]))
```
Q. (Doesn't make sense!)
## Family history of breast, endometrial and ovarian cancer 
```{r}
table(data_premeno$family_sitory_of_beo
      [as.character(data_premeno$br) == "Control"])
sum(table(data_premeno$family_sitory_of_beo
          [as.character(data_premeno$br) == "Control"]))
prop.table(table(data_premeno$family_sitory_of_beo
                 [as.character(data_premeno$br) == "Control"]))

table(data_premeno$family_sitory_of_beo
      [as.character(data_premeno$br) == "Breast Cancer Case"])
sum(table(data_premeno$family_sitory_of_beo
          [as.character(data_premeno$br) == "Breast Cancer Case"]))
prop.table(table(data_premeno$family_sitory_of_beo
                 [as.character(data_premeno$br) == "Breast Cancer Case"]))
```

## BMI
```{r}
table(data_premeno$bmicat
      [as.character(data_premeno$br) == "Control"])
sum(table(data_premeno$bmicat
          [as.character(data_premeno$br) == "Control"]))
prop.table(table(data_premeno$bmicat
                 [as.character(data_premeno$br) == "Control"]))

table(data_premeno$bmicat
      [as.character(data_premeno$br) == "Breast Cancer Case"])
sum(table(data_premeno$bmicat
          [as.character(data_premeno$br) == "Breast Cancer Case"]))
prop.table(table(data_premeno$bmicat
                 [as.character(data_premeno$br) == "Breast Cancer Case"]))
```

## Height
```{r}
table(data_premeno$catht
      [as.character(data_premeno$br) == "Control"])
sum(table(data_premeno$catht
          [as.character(data_premeno$br) == "Control"]))
prop.table(table(data_premeno$catht
                 [as.character(data_premeno$br) == "Control"]))

table(data_premeno$catht
      [as.character(data_premeno$br) == "Breast Cancer Case"])
sum(table(data_premeno$catht
          [as.character(data_premeno$br) == "Breast Cancer Case"]))
prop.table(table(data_premeno$catht
                 [as.character(data_premeno$br) == "Breast Cancer Case"]))
```

## Waist-to-hip ratio
```{r}
table(data_premeno$ratio
      [as.character(data_premeno$br) == "Control"])
sum(table(data_premeno$ratio
          [as.character(data_premeno$br) == "Control"]))
prop.table(table(data_premeno$ratio
                 [as.character(data_premeno$br) == "Control"]))

table(data_premeno$ratio
      [as.character(data_premeno$br) == "Breast Cancer Case"])
sum(table(data_premeno$ratio
          [as.character(data_premeno$br) == "Breast Cancer Case"]))
prop.table(table(data_premeno$ratio
                 [as.character(data_premeno$br) == "Breast Cancer Case"]))
```

## BMI and WHR interaction
```{r}
table(data_premeno$bmicat
      [as.character(data_premeno$br) == "Control"], 
      data_premeno$ratio
      [as.character(data_premeno$br) == "Control"])
sum(table(data_premeno$bmicat
          [as.character(data_premeno$br) == "Control"], 
          data_premeno$ratio
          [as.character(data_premeno$br) == "Control"]))
prop.table(table(data_premeno$bmicat
                 [as.character(data_premeno$br) == "Control"], 
                 data_premeno$ratio
                 [as.character(data_premeno$br) == "Control"]))


table(data_premeno$bmicat
      [as.character(data_premeno$br) == "Breast Cancer Case"], 
      data_premeno$ratio
      [as.character(data_premeno$br) == "Breast Cancer Case"])
sum(table(data_premeno$bmicat
          [as.character(data_premeno$br) == "Breast Cancer Case"], 
          data_premeno$ratio
          [as.character(data_premeno$br) == "Breast Cancer Case"]))
prop.table(table(data_premeno$bmicat
                 [as.character(data_premeno$br) == "Breast Cancer Case"], 
                 data_premeno$ratio
                 [as.character(data_premeno$br) == "Breast Cancer Case"]))
```

## Categorical age at enrollment
```{r}
table(data_premeno$agecode
      [as.character(data_premeno$br) == "Control"])
sum(table(data_premeno$agecode
          [as.character(data_premeno$br) == "Control"]))
prop.table(table(data_premeno$agecode
                 [as.character(data_premeno$br) == "Control"]))

table(data_premeno$agecode
      [as.character(data_premeno$br) == "Breast Cancer Case"])
sum(table(data_premeno$agecode
          [as.character(data_premeno$br) == "Breast Cancer Case"]))
prop.table(table(data_premeno$agecode
                 [as.character(data_premeno$br) == "Breast Cancer Case"]))
```

## Geographical location
```{r}
table(data_premeno$area
      [as.character(data_premeno$br) == "Control"])
sum(table(data_premeno$area
          [as.character(data_premeno$br) == "Control"]))
prop.table(table(data_premeno$area
                 [as.character(data_premeno$br) == "Control"]))

table(data_premeno$area
      [as.character(data_premeno$br) == "Breast Cancer Case"])
sum(table(data_premeno$area
          [as.character(data_premeno$br) == "Breast Cancer Case"]))
prop.table(table(data_premeno$area
                 [as.character(data_premeno$br) == "Breast Cancer Case"]))
```
(Important factor to avoid biased sample)
## Rural/urban based on first 20 years
```{r}
table(data_premeno$residancefirsttwentyyears
      [as.character(data_premeno$br) == "Control"])
sum(table(data_premeno$residancefirsttwentyyears
          [as.character(data_premeno$br) == "Control"]))
prop.table(table(data_premeno$residancefirsttwentyyears
                 [as.character(data_premeno$br) == "Control"]))

table(data_premeno$residancefirsttwentyyears
      [as.character(data_premeno$br) == "Breast Cancer Case"])
sum(table(data_premeno$residancefirsttwentyyears
          [as.character(data_premeno$br) == "Breast Cancer Case"]))
prop.table(table(data_premeno$residancefirsttwentyyears
                 [as.character(data_premeno$br) == "Breast Cancer Case"]))
```

## Binary education status: <8 years and >= 8 years of schooling
```{r}
table(data_premeno$ed
      [as.character(data_premeno$br) == "Control"])
sum(table(data_premeno$ed
          [as.character(data_premeno$br) == "Control"]))
prop.table(table(data_premeno$ed
                 [as.character(data_premeno$br) == "Control"]))

table(data_premeno$ed
      [as.character(data_premeno$br) == "Breast Cancer Case"])
sum(table(data_premeno$ed
          [as.character(data_premeno$br) == "Breast Cancer Case"]))
prop.table(table(data_premeno$ed
                 [as.character(data_premeno$br) == "Breast Cancer Case"]))
```

## Relative risk model for premenopausal women
```{r}
model_premeno = glm(br ~ as.factor(agecode) + as.factor(area) 
                    + as.factor(ed) + as.factor(residancefirsttwentyyears)
                    + as.factor(ageatfirstfulltermpreg_cat)
                    + interval_menarche_firstbirth_cat   
                    + as.factor(fulltermpreg_cat)
                    + oc + as.factor(mean_brfeed_duration_cat)
                    + as.factor(totalabortion_cat)
                    + as.factor(totalmiscarriage_cat)
                    + as.factor(catht) 
                    + as.factor(bmicat) * as.factor(ratio),
                    data = data_premeno, family = binomial(link=logit), 
                    na.action = na.omit)

cbind(exp(coef(model_premeno)),exp(confint(model_premeno)))
```

##Joint association of BMI and WHR#####
```{r}
#Log relative risks and variances
estimate = coef(model_premeno)[c("as.factor(bmicat)2","as.factor(bmicat)3",
                                 "as.factor(ratio)1","as.factor(ratio)2",
                                 "as.factor(bmicat)2:as.factor(ratio)1",
                                 "as.factor(bmicat)3:as.factor(ratio)1",
                                 "as.factor(bmicat)2:as.factor(ratio)2",
                                 "as.factor(bmicat)3:as.factor(ratio)2")]

variance = vcov(model_premeno)[c("as.factor(bmicat)2","as.factor(bmicat)3",
                                 "as.factor(ratio)1","as.factor(ratio)2",
                                 "as.factor(bmicat)2:as.factor(ratio)1",
                                 "as.factor(bmicat)3:as.factor(ratio)1",
                                 "as.factor(bmicat)2:as.factor(ratio)2",
                                 "as.factor(bmicat)3:as.factor(ratio)2"),
                               c("as.factor(bmicat)2","as.factor(bmicat)3",
                                 "as.factor(ratio)1","as.factor(ratio)2",
                                 "as.factor(bmicat)2:as.factor(ratio)1",
                                 "as.factor(bmicat)3:as.factor(ratio)1",
                                 "as.factor(bmicat)2:as.factor(ratio)2",
                                 "as.factor(bmicat)3:as.factor(ratio)2")]

#BMI < 24.9, WHR < 0.84

RR1 = 1

ci1 = c(1,1)

#BMI < 24.9, WHR 0.85-0.94

RR2 = exp(coef(model_premeno))["as.factor(ratio)1"]

ci2 = exp(confint(model_premeno))["as.factor(ratio)1",]
#BMI < 24.9, WHR >= 0.95

RR3 = exp(coef(model_premeno))["as.factor(ratio)2"]

ci3 = exp(confint(model_premeno))["as.factor(ratio)2",]

#BMI 25-29.9, WHR < 0.84

RR4 = exp(coef(model_premeno))["as.factor(bmicat)2"]

ci4 = exp(confint(model_premeno))["as.factor(bmicat)2",]

#BMI 25-29.9, WHR 0.85-0.94

RR5 = exp(coef(model_premeno))["as.factor(bmicat)2"] * 
  exp(coef(model_premeno))["as.factor(ratio)1"] * 
  exp(coef(model_premeno))["as.factor(bmicat)2:as.factor(ratio)1"]

v5 = c(1,0,1,0,1,0,0,0)

RR5 = exp(v5 %*% estimate)

sd5 = sqrt(v5 %*% variance %*% v5)

ci5 = exp(c(log(RR5) - 1.96 * sd5, log(RR5) + 1.96 * sd5))

#BMI 25-29.9, WHR >= 0.95

RR6 = exp(coef(model_premeno))["as.factor(bmicat)2"] * 
  exp(coef(model_premeno))["as.factor(ratio)2"] * 
  exp(coef(model_premeno))["as.factor(bmicat)2:as.factor(ratio)2"]

v6 = c(1,0,0,1,0,0,1,0)

RR6 = exp(v6 %*% estimate)

sd6 = sqrt(v6 %*% variance %*% v6)

ci6 = exp(c(log(RR6) - 1.96 * sd6, log(RR6) + 1.96 * sd6))

#BMI >= 30, WHR < 0.84

RR7 = exp(coef(model_premeno))["as.factor(bmicat)3"]

ci7 = exp(confint(model_premeno))["as.factor(bmicat)3",]

#BMI >= 30, WHR 0.85-0.94

RR8 = exp(coef(model_premeno))["as.factor(bmicat)3"] * 
  exp(coef(model_premeno))["as.factor(ratio)1"] * 
  exp(coef(model_premeno))["as.factor(bmicat)3:as.factor(ratio)1"]

v8 = c(0,1,1,0,0,1,0,0)

RR8 = exp(v8 %*% estimate)

sd8 = sqrt(v8 %*% variance %*% v8)

ci8 = exp(c(log(RR8) - 1.96 * sd8, log(RR8) + 1.96 * sd8))

#BMI >= 30, WHR >= 0.95

RR9 = exp(coef(model_premeno))["as.factor(bmicat)3"] * 
  exp(coef(model_premeno))["as.factor(ratio)2"] * 
  exp(coef(model_premeno))["as.factor(bmicat)3:as.factor(ratio)2"]

v9 = c(0,1,0,1,0,0,0,1)

RR9 = exp(v9 %*% estimate)

sd9 = sqrt(v9 %*% variance %*% v9)

ci9 = exp(c(log(RR9) - 1.96 * sd9, log(RR9) + 1.96 * sd9))


BMI_WHR_Premeno = cbind(c(RR1,RR2,RR3,RR4,RR5,RR6,RR7,RR8,RR9),
                        rbind(ci1,ci2,ci3,ci4,ci5,ci6,ci7,ci8,ci9))

rownames(BMI_WHR_Premeno)[c(1,5,6,8,9)] = c("Ref","as.factor(bmicat)2:as.factor(ratio)1",
                                         "as.factor(bmicat)2:as.factor(ratio)2",
                                         "as.factor(bmicat)3:as.factor(ratio)1",
                                         "as.factor(bmicat)2:as.factor(ratio)2")

#pdf(file = "BMI_WHR_Interaction_Premeno")
plotCI(x = BMI_WHR_Premeno[,1], ui = BMI_WHR_Premeno[,3], li = BMI_WHR_Premeno[,2],xlab = "",
       ylab = "Joint Relative Risk",pch=16,lwd = 2, xaxt = "n", 
       col = c(rep("red",3),rep("blue",3),rep("green",3)))

mtext(expression("BMI"<="24.9"), side = 3, line = 1, at = 2, col = "red")
mtext(expression("BMI 25-29.9"), side = 3, line = 1, at = 5, col = "blue")
mtext(expression("BMI">="30"), side = 3, line = 1, at = 8, col = "green")

mtext(rep(expression("WHR"<=0.84),3), side = 1, line = 0.2, las = 2, 
      at = c(1,4,7), col = c("red","blue","green"), cex = 0.7)
mtext(rep("WHR 0.85-0.94",3), side = 1, line = 0.2, las = 2, 
      at = c(2,5,8), col = c("red","blue","green"), cex = 0.7)
mtext(rep(expression("WHR">=0.95),3), side = 1, line = 0.2, las = 2, 
      at = c(3,6,9), col = c("red","blue","green"), cex = 0.7)
#dev.off()

```

## Subset the data to post-menopausal women
```{r}
data_postmeno = data[which(as.character(data$menopause) == "postmenopausal"),]
```
## Number of cases and controls
```{r}
table(data_postmeno$br)
```

## Number and proportion of risk factors among controls

## Age at first full term pregnancy
```{r}
table(data_postmeno$ageatfirstfulltermpreg_cat
      [as.character(data_postmeno$br) == "Control"])
sum(table(data_postmeno$ageatfirstfulltermpreg_cat
          [as.character(data_postmeno$br) == "Control"]))
prop.table(table(data_postmeno$ageatfirstfulltermpreg_cat
                 [as.character(data_postmeno$br) == "Control"]))

table(data_postmeno$ageatfirstfulltermpreg_cat
      [as.character(data_postmeno$br) == "Breast Cancer Case"])
sum(table(data_postmeno$ageatfirstfulltermpreg_cat
          [as.character(data_postmeno$br) == "Breast Cancer Case"]))
prop.table(table(data_postmeno$ageatfirstfulltermpreg_cat
                 [as.character(data_postmeno$br) == "Breast Cancer Case"]))
```

## Interval between menarche and first full term pregnancy
```{r}
table(data_postmeno$interval_menarche_firstbirth_cat
      [as.character(data_postmeno$br) == "Control"])
sum(table(data_postmeno$interval_menarche_firstbirth_cat
          [as.character(data_postmeno$br) == "Control"]))
prop.table(table(data_postmeno$interval_menarche_firstbirth_cat
                 [as.character(data_postmeno$br) == "Control"]))

table(data_postmeno$interval_menarche_firstbirth_cat
      [as.character(data_postmeno$br) == "Breast Cancer Case"])
sum(table(data_postmeno$interval_menarche_firstbirth_cat
          [as.character(data_postmeno$br) == "Breast Cancer Case"]))
prop.table(table(data_postmeno$interval_menarche_firstbirth_cat
                 [as.character(data_postmeno$br) == "Breast Cancer Case"]))
```

## Number of full term pregnancies
```{r}
table(data_postmeno$fulltermpreg_cat
      [as.character(data_postmeno$br) == "Control"])
sum(table(data_postmeno$fulltermpreg_cat
          [as.character(data_postmeno$br) == "Control"]))
prop.table(table(data_postmeno$fulltermpreg_cat
                 [as.character(data_postmeno$br) == "Control"]))

table(data_postmeno$fulltermpreg_cat
      [as.character(data_postmeno$br) == "Breast Cancer Case"])
sum(table(data_postmeno$fulltermpreg_cat
          [as.character(data_postmeno$br) == "Breast Cancer Case"]))
prop.table(table(data_postmeno$fulltermpreg_cat
                 [as.character(data_postmeno$br) == "Breast Cancer Case"]))
```

## OC Use
```{r}
table(data_postmeno$oc
      [as.character(data_postmeno$br) == "Control"])
sum(table(data_postmeno$oc
          [as.character(data_postmeno$br) == "Control"]))
prop.table(table(data_postmeno$oc
                 [as.character(data_postmeno$br) == "Control"]))

table(data_postmeno$oc
      [as.character(data_postmeno$br) == "Breast Cancer Case"])
sum(table(data_postmeno$oc
          [as.character(data_postmeno$br) == "Breast Cancer Case"]))
prop.table(table(data_postmeno$oc
                 [as.character(data_postmeno$br) == "Breast Cancer Case"]))
```

##Age at menopause
```{r}
table(data_postmeno$catmenopause
      [as.character(data_postmeno$br) == "Control"])
sum(table(data_postmeno$catmenopause
          [as.character(data_postmeno$br) == "Control"]))
prop.table(table(data_postmeno$catmenopause
                 [as.character(data_postmeno$br) == "Control"]))

table(data_postmeno$catmenopause
      [as.character(data_postmeno$br) == "Breast Cancer Case"])
sum(table(data_postmeno$catmenopause
          [as.character(data_postmeno$br) == "Breast Cancer Case"]))
prop.table(table(data_postmeno$catmenopause
                 [as.character(data_postmeno$br) == "Breast Cancer Case"]))
```

## Duration of breast-feeding
```{r}
table(data_postmeno$mean_brfeed_duration_cat
      [as.character(data_postmeno$br) == "Control"])
sum(table(data_postmeno$mean_brfeed_duration_cat
          [as.character(data_postmeno$br) == "Control"]))
prop.table(table(data_postmeno$mean_brfeed_duration_cat
                 [as.character(data_postmeno$br) == "Control"]))

table(data_postmeno$mean_brfeed_duration_cat
      [as.character(data_postmeno$br) == "Breast Cancer Case"])
sum(table(data_postmeno$mean_brfeed_duration_cat
          [as.character(data_postmeno$br) == "Breast Cancer Case"]))
prop.table(table(data_postmeno$mean_brfeed_duration_cat
                 [as.character(data_postmeno$br) == "Breast Cancer Case"]))
```

## Number of induced abortions
```{r}
table(data_postmeno$totalabortion_cat
      [as.character(data_postmeno$br) == "Control"])
sum(table(data_postmeno$totalabortion_cat
          [as.character(data_postmeno$br) == "Control"]))
prop.table(table(data_postmeno$totalabortion_cat
                 [as.character(data_postmeno$br) == "Control"]))

table(data_postmeno$totalabortion_cat
      [as.character(data_postmeno$br) == "Breast Cancer Case"])
sum(table(data_postmeno$totalabortion_cat
          [as.character(data_postmeno$br) == "Breast Cancer Case"]))
prop.table(table(data_postmeno$totalabortion_cat
                 [as.character(data_postmeno$br) == "Breast Cancer Case"]))
```

## Number of spontaneous abortions
```{r}
table(data_postmeno$totalmiscarriage_cat
      [as.character(data_postmeno$br) == "Control"])
sum(table(data_postmeno$totalmiscarriage_cat
          [as.character(data_postmeno$br) == "Control"]))
prop.table(table(data_postmeno$totalmiscarriage_cat
                 [as.character(data_postmeno$br) == "Control"]))

table(data_postmeno$totalmiscarriage_cat
      [as.character(data_postmeno$br) == "Breast Cancer Case"])
sum(table(data_postmeno$totalmiscarriage_cat
          [as.character(data_postmeno$br) == "Breast Cancer Case"]))
prop.table(table(data_postmeno$totalmiscarriage_cat
                 [as.character(data_postmeno$br) == "Breast Cancer Case"]))
```
(why no differential family history)
## Family history of breast, endometrial and ovarian cancer 
```{r}
table(data_postmeno$family_sitory_of_beo
      [as.character(data_postmeno$br) == "Control"])
sum(table(data_postmeno$family_sitory_of_beo
          [as.character(data_postmeno$br) == "Control"]))
prop.table(table(data_postmeno$family_sitory_of_beo
                 [as.character(data_postmeno$br) == "Control"]))

table(data_postmeno$family_sitory_of_beo
      [as.character(data_postmeno$br) == "Breast Cancer Case"])
sum(table(data_postmeno$family_sitory_of_beo
          [as.character(data_postmeno$br) == "Breast Cancer Case"]))
prop.table(table(data_postmeno$family_sitory_of_beo
                 [as.character(data_postmeno$br) == "Breast Cancer Case"]))
```

## BMI
```{r}
table(data_postmeno$bmicat
      [as.character(data_postmeno$br) == "Control"])
sum(table(data_postmeno$bmicat
          [as.character(data_postmeno$br) == "Control"]))
prop.table(table(data_postmeno$bmicat
                 [as.character(data_postmeno$br) == "Control"]))

table(data_postmeno$bmicat
      [as.character(data_postmeno$br) == "Breast Cancer Case"])
sum(table(data_postmeno$bmicat
          [as.character(data_postmeno$br) == "Breast Cancer Case"]))
prop.table(table(data_postmeno$bmicat
                 [as.character(data_postmeno$br) == "Breast Cancer Case"]))
```

## Height
```{r}
table(data_postmeno$catht
      [as.character(data_postmeno$br) == "Control"])
sum(table(data_postmeno$catht
          [as.character(data_postmeno$br) == "Control"]))
prop.table(table(data_postmeno$catht
                 [as.character(data_postmeno$br) == "Control"]))

table(data_postmeno$catht
      [as.character(data_postmeno$br) == "Breast Cancer Case"])
sum(table(data_postmeno$catht
          [as.character(data_postmeno$br) == "Breast Cancer Case"]))
prop.table(table(data_postmeno$catht
                 [as.character(data_postmeno$br) == "Breast Cancer Case"]))
```

## Waist-to-hip ratio
```{r}
table(data_postmeno$ratio
      [as.character(data_postmeno$br) == "Control"])
sum(table(data_postmeno$ratio
          [as.character(data_postmeno$br) == "Control"]))
prop.table(table(data_postmeno$ratio
                 [as.character(data_postmeno$br) == "Control"]))

table(data_postmeno$ratio
      [as.character(data_postmeno$br) == "Breast Cancer Case"])
sum(table(data_postmeno$ratio
          [as.character(data_postmeno$br) == "Breast Cancer Case"]))
prop.table(table(data_postmeno$ratio
                 [as.character(data_postmeno$br) == "Breast Cancer Case"]))
```

## Duration of diabetes
```{r}
table(data_postmeno$diabetic_years_cat
      [as.character(data_postmeno$br) == "Control"])
sum(table(data_postmeno$diabetic_years_cat
          [as.character(data_postmeno$br) == "Control"]))
prop.table(table(data_postmeno$diabetic_years_cat
                 [as.character(data_postmeno$br) == "Control"]))

table(data_postmeno$diabetic_years_cat
      [as.character(data_postmeno$br) == "Breast Cancer Case"])
sum(table(data_postmeno$diabetic_years_cat
          [as.character(data_postmeno$br) == "Breast Cancer Case"]))
prop.table(table(data_postmeno$diabetic_years_cat
                 [as.character(data_postmeno$br) == "Breast Cancer Case"]))
```
## New variable for duration of diabetes collapsing the levels 0 and 1
```{r}
data_postmeno$diabetes_duration = data_postmeno$diabetic_years_cat
data_postmeno$diabetes_duration[data_postmeno$diabetes_duration <= 1] = 1
```

## Diabetes
```{r}
table(data_postmeno$diabetes_YN
      [as.character(data_postmeno$br) == "Control"])
sum(table(data_postmeno$diabetes_YN
          [as.character(data_postmeno$br) == "Control"]))
prop.table(table(data_postmeno$diabetes_YN
                 [as.character(data_postmeno$br) == "Control"]))

table(data_postmeno$diabetes_YN
      [as.character(data_postmeno$br) == "Breast Cancer Case"])
sum(table(data_postmeno$diabetes_YN
          [as.character(data_postmeno$br) == "Breast Cancer Case"]))
prop.table(table(data_postmeno$diabetes_YN
                 [as.character(data_postmeno$br) == "Breast Cancer Case"]))
```

## Family history of diabetes
```{r}
table(data_postmeno$F_H_diabetes
      [as.character(data_postmeno$br) == "Control"])
sum(table(data_postmeno$F_H_diabetes
          [as.character(data_postmeno$br) == "Control"]))
prop.table(table(data_postmeno$F_H_diabetes
                 [as.character(data_postmeno$br) == "Control"]))

table(data_postmeno$F_H_diabetes
      [as.character(data_postmeno$br) == "Breast Cancer Case"])
sum(table(data_postmeno$F_H_diabetes
          [as.character(data_postmeno$br) == "Breast Cancer Case"]))
prop.table(table(data_postmeno$F_H_diabetes
                 [as.character(data_postmeno$br) == "Breast Cancer Case"]))
```


## BMI and WHR interaction
```{r}
table(data_postmeno$bmicat
      [as.character(data_postmeno$br) == "Control"], 
      data_postmeno$ratio
      [as.character(data_postmeno$br) == "Control"])
sum(table(data_postmeno$bmicat
          [as.character(data_postmeno$br) == "Control"], 
          data_postmeno$ratio
          [as.character(data_postmeno$br) == "Control"]))
prop.table(table(data_postmeno$bmicat
                 [as.character(data_postmeno$br) == "Control"], 
                 data_postmeno$ratio
                 [as.character(data_postmeno$br) == "Control"]))


table(data_postmeno$bmicat
      [as.character(data_postmeno$br) == "Breast Cancer Case"], 
      data_postmeno$ratio
      [as.character(data_postmeno$br) == "Breast Cancer Case"])
sum(table(data_postmeno$bmicat
          [as.character(data_postmeno$br) == "Breast Cancer Case"], 
          data_postmeno$ratio
          [as.character(data_postmeno$br) == "Breast Cancer Case"]))
prop.table(table(data_postmeno$bmicat
                 [as.character(data_postmeno$br) == "Breast Cancer Case"], 
                 data_postmeno$ratio
                 [as.character(data_postmeno$br) == "Breast Cancer Case"]))
```

## BMI and diabetes interaction
```{r}
table(data_postmeno$bmicat
      [as.character(data_postmeno$br) == "Control"], 
      data_postmeno$diabetes_YN
      [as.character(data_postmeno$br) == "Control"])
sum(table(data_postmeno$bmicat
          [as.character(data_postmeno$br) == "Control"], 
          data_postmeno$diabetes_YN
          [as.character(data_postmeno$br) == "Control"]))
prop.table(table(data_postmeno$bmicat
                 [as.character(data_postmeno$br) == "Control"], 
                 data_postmeno$diabetes_YN
                 [as.character(data_postmeno$br) == "Control"]))


table(data_postmeno$bmicat
      [as.character(data_postmeno$br) == "Breast Cancer Case"], 
      data_postmeno$diabetes_YN
      [as.character(data_postmeno$br) == "Breast Cancer Case"])
sum(table(data_postmeno$bmicat
          [as.character(data_postmeno$br) == "Breast Cancer Case"], 
          data_postmeno$diabetes_YN
          [as.character(data_postmeno$br) == "Breast Cancer Case"]))
prop.table(table(data_postmeno$bmicat
                 [as.character(data_postmeno$br) == "Breast Cancer Case"], 
                 data_postmeno$diabetes_YN
                 [as.character(data_postmeno$br) == "Breast Cancer Case"]))

```

## Categorical age at enrollment
```{r}
table(data_postmeno$agecode
      [as.character(data_postmeno$br) == "Control"])
sum(table(data_postmeno$agecode
          [as.character(data_postmeno$br) == "Control"]))
prop.table(table(data_postmeno$agecode
                 [as.character(data_postmeno$br) == "Control"]))

table(data_postmeno$agecode
      [as.character(data_postmeno$br) == "Breast Cancer Case"])
sum(table(data_postmeno$agecode
          [as.character(data_postmeno$br) == "Breast Cancer Case"]))
prop.table(table(data_postmeno$agecode
                 [as.character(data_postmeno$br) == "Breast Cancer Case"]))
```

## Geographical location
```{r}
table(data_postmeno$area
      [as.character(data_postmeno$br) == "Control"])
sum(table(data_postmeno$area
          [as.character(data_postmeno$br) == "Control"]))
prop.table(table(data_postmeno$area
                 [as.character(data_postmeno$br) == "Control"]))

table(data_postmeno$area
      [as.character(data_postmeno$br) == "Breast Cancer Case"])
sum(table(data_postmeno$area
          [as.character(data_postmeno$br) == "Breast Cancer Case"]))
prop.table(table(data_postmeno$area
                 [as.character(data_postmeno$br) == "Breast Cancer Case"]))
```

## Rural/urban based on first 20 years
```{r}
table(data_postmeno$residancefirsttwentyyears
      [as.character(data_postmeno$br) == "Control"])
sum(table(data_postmeno$residancefirsttwentyyears
          [as.character(data_postmeno$br) == "Control"]))
prop.table(table(data_postmeno$residancefirsttwentyyears
                 [as.character(data_postmeno$br) == "Control"]))

table(data_postmeno$residancefirsttwentyyears
      [as.character(data_postmeno$br) == "Breast Cancer Case"])
sum(table(data_postmeno$residancefirsttwentyyears
          [as.character(data_postmeno$br) == "Breast Cancer Case"]))
prop.table(table(data_postmeno$residancefirsttwentyyears
                 [as.character(data_postmeno$br) == "Breast Cancer Case"]))
```

## Binary education status: <8 years and >= 8 years of schooling
```{r}
table(data_postmeno$ed
      [as.character(data_postmeno$br) == "Control"])
sum(table(data_postmeno$ed
          [as.character(data_postmeno$br) == "Control"]))
prop.table(table(data_postmeno$ed
                 [as.character(data_postmeno$br) == "Control"]))

table(data_postmeno$ed
      [as.character(data_postmeno$br) == "Breast Cancer Case"])
sum(table(data_postmeno$ed
          [as.character(data_postmeno$br) == "Breast Cancer Case"]))
prop.table(table(data_postmeno$ed
                 [as.character(data_postmeno$br) == "Breast Cancer Case"]))
```

##Relative risk model for postmenopausal women
```{r}
model_postmeno = glm(br ~ as.factor(agecode) + as.factor(area) 
                    + as.factor(ed) + as.factor(residancefirsttwentyyears)
                    + as.factor(ageatfirstfulltermpreg_cat)
                    + interval_menarche_firstbirth_cat
                    + as.factor(fulltermpreg_cat)
                    + as.factor(catmenopause)
                    + oc + as.factor(mean_brfeed_duration_cat)
                    + as.factor(totalabortion_cat)
                    + as.factor(totalmiscarriage_cat)
                    + as.factor(catht) + as.factor(bmicat) * as.factor(ratio),
                    data = data_postmeno, family = binomial(link=logit),
                    na.action = na.omit)

cbind(exp(coef(model_postmeno)),exp(confint(model_postmeno)))
```

## Joint association of BMI and WHR#####
```{r}
#Log relative risks and variances
estimate = coef(model_postmeno)[c("as.factor(bmicat)2","as.factor(bmicat)3",
                                 "as.factor(ratio)1","as.factor(ratio)2",
                                 "as.factor(bmicat)2:as.factor(ratio)1",
                                 "as.factor(bmicat)3:as.factor(ratio)1",
                                 "as.factor(bmicat)2:as.factor(ratio)2",
                                 "as.factor(bmicat)3:as.factor(ratio)2")]

variance = vcov(model_postmeno)[c("as.factor(bmicat)2","as.factor(bmicat)3",
                                 "as.factor(ratio)1","as.factor(ratio)2",
                                 "as.factor(bmicat)2:as.factor(ratio)1",
                                 "as.factor(bmicat)3:as.factor(ratio)1",
                                 "as.factor(bmicat)2:as.factor(ratio)2",
                                 "as.factor(bmicat)3:as.factor(ratio)2"),
                               c("as.factor(bmicat)2","as.factor(bmicat)3",
                                 "as.factor(ratio)1","as.factor(ratio)2",
                                 "as.factor(bmicat)2:as.factor(ratio)1",
                                 "as.factor(bmicat)3:as.factor(ratio)1",
                                 "as.factor(bmicat)2:as.factor(ratio)2",
                                 "as.factor(bmicat)3:as.factor(ratio)2")]

#BMI < 24.9, WHR < 0.84

RR1 = 1

ci1 = c(1,1)

#BMI < 24.9, WHR 0.85-0.94

RR2 = exp(coef(model_postmeno))["as.factor(ratio)1"]

ci2 = exp(confint(model_postmeno))["as.factor(ratio)1",]

#BMI < 24.9, WHR >= 0.95

RR3 = exp(coef(model_postmeno))["as.factor(ratio)2"]

ci3 = exp(confint(model_postmeno))["as.factor(ratio)2",]

#BMI 25-29.9, WHR < 0.84

RR4 = exp(coef(model_postmeno))["as.factor(bmicat)2"]

ci4 = exp(confint(model_postmeno))["as.factor(bmicat)2",]

#BMI 25-29.9, WHR 0.85-0.94

RR5 = exp(coef(model_postmeno))["as.factor(bmicat)2"] * 
  exp(coef(model_postmeno))["as.factor(ratio)1"] * 
  exp(coef(model_postmeno))["as.factor(bmicat)2:as.factor(ratio)1"]

v5 = c(1,0,1,0,1,0,0,0)

RR5 = exp(v5 %*% estimate)

sd5 = sqrt(v5 %*% variance %*% v5)

ci5 = exp(c(log(RR5) - 1.96 * sd5, log(RR5) + 1.96 * sd5))


#BMI 25-29.9, WHR >= 0.95

RR6 = exp(coef(model_postmeno))["as.factor(bmicat)2"] * 
  exp(coef(model_postmeno))["as.factor(ratio)2"] * 
  exp(coef(model_postmeno))["as.factor(bmicat)2:as.factor(ratio)2"]

v6 = c(1,0,0,1,0,0,1,0)

RR6 = exp(v6 %*% estimate)

sd6 = sqrt(v6 %*% variance %*% v6)

ci6 = exp(c(log(RR6) - 1.96 * sd6, log(RR6) + 1.96 * sd6))

#BMI >= 30, WHR < 0.84

RR7 = exp(coef(model_postmeno))["as.factor(bmicat)3"]

ci7 = exp(confint(model_postmeno))["as.factor(bmicat)3",]

#BMI >= 30, WHR 0.85-0.94

RR8 = exp(coef(model_postmeno))["as.factor(bmicat)3"] * 
  exp(coef(model_postmeno))["as.factor(ratio)1"] * 
  exp(coef(model_postmeno))["as.factor(bmicat)3:as.factor(ratio)1"]

v8 = c(0,1,1,0,0,1,0,0)

RR8 = exp(v8 %*% estimate)

sd8 = sqrt(v8 %*% variance %*% v8)

ci8 = exp(c(log(RR8) - 1.96 * sd8, log(RR8) + 1.96 * sd8))

#BMI >= 30, WHR >= 0.95

RR9 = exp(coef(model_postmeno))["as.factor(bmicat)3"] * 
  exp(coef(model_postmeno))["as.factor(ratio)2"] * 
  exp(coef(model_postmeno))["as.factor(bmicat)3:as.factor(ratio)2"]

v9 = c(0,1,0,1,0,0,0,1)

RR9 = exp(v9 %*% estimate)

sd9 = sqrt(v9 %*% variance %*% v9)

ci9 = exp(c(log(RR9) - 1.96 * sd9, log(RR9) + 1.96 * sd9))

BMI_WHR_Postmeno = cbind(c(RR1,RR2,RR3,RR4,RR5,RR6,RR7,RR8,RR9),
                        rbind(ci1,ci2,ci3,ci4,ci5,ci6,ci7,ci8,ci9))

rownames(BMI_WHR_Postmeno)[c(1,5,6,8,9)] = c("Ref","as.factor(bmicat)2:as.factor(ratio)1",
                                            "as.factor(bmicat)2:as.factor(ratio)2",
                                            "as.factor(bmicat)3:as.factor(ratio)1",
                                            "as.factor(bmicat)2:as.factor(ratio)2")


#pdf(file = "BMI_WHR_Interaction_Postmeno")
plotCI(x = BMI_WHR_Postmeno[,1], ui = BMI_WHR_Postmeno[,3], li = BMI_WHR_Postmeno[,2],
       xlab = "", ylab = "Joint Relative Risk",pch=16,lwd = 2,xaxt = "n",
       col = c(rep("red",3),rep("blue",3),rep("green",3)))

mtext(expression("BMI"<="24.9"), side = 3, line = 1, at = 2, col = "red")
mtext(expression("BMI 25-29.9"), side = 3, line = 1, at = 5, col = "blue")
mtext(expression("BMI">="30"), side = 3, line = 1, at = 8, col = "green")

mtext(rep(expression("WHR"<=0.84),3), side = 1, line = 0.2, las = 2, 
      at = c(1,4,7), col = c("red","blue","green"), cex = 0.7)
mtext(rep("WHR 0.85-0.94",3), side = 1, line = 0.2, las = 2, 
      at = c(2,5,8), col = c("red","blue","green"), cex = 0.7)
mtext(rep(expression("WHR">=0.95),3), side = 1, line = 0.2, las = 2, 
      at = c(3,6,9), col = c("red","blue","green"), cex = 0.7)
#dev.off()
```


# Absolute risk prediction using iCARE

## Function to format incidence and mortality rates
```{r}
formatted_rates = function(mat){
  if(ncol(mat)==3){
    start <- mat[,1]
    end   <- mat[,2]
    rate  <- mat[,3]
    
    integer_ages <- seq(min(start), max(end))
    formatted    <- cbind( integer_ages, rep(NA, length(integer_ages)))
    for(i in 1:nrow(mat)){
      these              <- which(formatted[,1]>=start[i] & 
                                    formatted[,1]<=end[i])
      #formatted[these,2] <- rate[i]/length(these)
      formatted[these,2] <- rate[i]
    }
    colnames(formatted) <- c("ages", "rates")
    formatted
  }else{
    formatted = mat
  }
  formatted
}
```

## Formatting breast cancer incidence rate 
```{r}
setwd("/Users/anaghchattopadhyay/Documents/Nilanjan\ Project/TMC_BCProject/Data")

bc_inci = read.xlsx("BC_Incidence_Rates.xlsx")
bc_inci = formatted_rates(bc_inci)
bc_inci[,2] = bc_inci[,2]/100000
```

## Formatting overall mortality rate
```{r}
setwd("/Users/anaghchattopadhyay/Documents/Nilanjan\ Project/TMC_BCProject/Data")
mort_inci = read.xlsx("Overall_Mortality_Rates.xlsx")
mort_inci = formatted_rates(mort_inci)
mort_inci[,2] = mort_inci[,2]/1000
```

## Creating the reference dataset
```{r}
setwd("/Users/anaghchattopadhyay/Documents/Nilanjan\ Project/TMC_BCProject/Data")
ref_data = read.xlsx("Breast_1.12.22.xlsx")
test_data = read.xlsx("parichoy_br_data_6.2.23.xlsx")
```

## Categories for age at first full term pregnancy
```{r}
ref_data$ageatfirstfulltermpreg_cat = 
  ref_data$age_at_first_fullterm_preg
ref_data$ageatfirstfulltermpreg_cat[which(ref_data$age_at_first_fullterm_preg < 20)] = 1
ref_data$ageatfirstfulltermpreg_cat[which(ref_data$age_at_first_fullterm_preg == 20 
                                          | ref_data$age_at_first_fullterm_preg == 21)] = 2
ref_data$ageatfirstfulltermpreg_cat[which(ref_data$age_at_first_fullterm_preg == 22 
                                          | ref_data$age_at_first_fullterm_preg == 23)] = 3
ref_data$ageatfirstfulltermpreg_cat[which(ref_data$age_at_first_fullterm_preg == 24 
                                          | ref_data$age_at_first_fullterm_preg == 25)] = 4
ref_data$ageatfirstfulltermpreg_cat[which(ref_data$age_at_first_fullterm_preg >= 26)] = 5


test_data$ageatfirstfulltermpreg_cat = 
  test_data$age_at_first_fullterm_preg
test_data$ageatfirstfulltermpreg_cat[which(test_data$age_at_first_fullterm_preg < 20)] = 1
test_data$ageatfirstfulltermpreg_cat[which(test_data$age_at_first_fullterm_preg == 20 
                                          | test_data$age_at_first_fullterm_preg == 21)] = 2
test_data$ageatfirstfulltermpreg_cat[which(test_data$age_at_first_fullterm_preg == 22 
                                          | test_data$age_at_first_fullterm_preg == 23)] = 3
test_data$ageatfirstfulltermpreg_cat[which(test_data$age_at_first_fullterm_preg == 24 
                                          | test_data$age_at_first_fullterm_preg == 25)] = 4
test_data$ageatfirstfulltermpreg_cat[which(test_data$age_at_first_fullterm_preg >= 26)] = 5

table(test_data$ageatfirstfulltermpreg_cat)
sum(is.na(test_data$age_at_first_fullterm_preg))

```

##Categories of interval between menarche and full term pregnancy
```{r}
ref_data$interval_menarche_firstbirth_cat = 
  ref_data$durn_menarch_first_fullterm_preg
ref_data$interval_menarche_firstbirth_cat[ref_data$durn_menarch_first_fullterm_preg < 10] = 0
ref_data$interval_menarche_firstbirth_cat[ref_data$durn_menarch_first_fullterm_preg >= 10] = 1

test_data$interval_menarche_firstbirth_cat = 
  test_data$durn_menarch_first_fullterm_preg
test_data$interval_menarche_firstbirth_cat[test_data$durn_menarch_first_fullterm_preg < 10] = 0
test_data$interval_menarche_firstbirth_cat[test_data$durn_menarch_first_fullterm_preg >= 10] = 1

table(test_data$interval_menarche_firstbirth_cat)
sum(is.na(test_data$interval_menarche_firstbirth_cat))
```

## Categories for number of full term pregnancies
```{r}
ref_data$fulltermpreg_cat = ref_data$No_of_live_birth
ref_data$fulltermpreg_cat[ref_data$No_of_live_birth == 0] = 1
ref_data$fulltermpreg_cat[ref_data$No_of_live_birth >= 4] = 4

test_data$fulltermpreg_cat = test_data$No_of_live_birth
test_data$fulltermpreg_cat[test_data$No_of_live_birth == 0] = 1
test_data$fulltermpreg_cat[test_data$No_of_live_birth >= 4] = 4

table(test_data$fulltermpreg_cat)
sum(is.na(test_data$fulltermpreg_cat))

```

## OC use
```{r}
ref_data$oc = NA
ref_data$oc[ref_data$oc_use_yn == 2] = 0
ref_data$oc[ref_data$oc_use_yn == 1] = 1

test_data$oc = NA
test_data$oc[test_data$oc_use_yn == 2] = 0
test_data$oc[test_data$oc_use_yn == 1] = 1

table(test_data$oc)
sum(is.na(test_data$oc))
```

## Categories for number of induced abortions
```{r}
ref_data$totalabortion_cat = ref_data$count_abortion
ref_data$totalabortion_cat[ref_data$count_abortion >= 2] = 2

test_data$totalabortion_cat = test_data$count_abortion
test_data$totalabortion_cat[test_data$count_abortion >= 2] = 2

table(test_data$totalabortion_cat)
sum(is.na(test_data$totalabortion_cat))
```

## Categories for number of spontaneous abortions
```{r}
ref_data$totalmiscarriage_cat = ref_data$count_miscarriage
ref_data$totalmiscarriage_cat[ref_data$count_miscarriage >= 2] = 2

test_data$totalmiscarriage_cat = test_data$count_miscarriage
test_data$totalmiscarriage_cat[test_data$count_miscarriage >= 2] = 2

table(test_data$totalmiscarriage_cat)
sum(is.na(test_data$totalmiscarriage_cat))
```

## Categories for duration of breastfeeding
```{r}
ref_data$mean_brfeed_duration_cat = ref_data$avg_dur_breastfeeding
ref_data$mean_brfeed_duration_cat[ref_data$avg_dur_breastfeeding < 12] = 1
ref_data$mean_brfeed_duration_cat[ref_data$avg_dur_breastfeeding >= 12 
                                  & ref_data$avg_dur_breastfeeding < 24] = 2
ref_data$mean_brfeed_duration_cat[ref_data$avg_dur_breastfeeding >= 24 
                                  & ref_data$avg_dur_breastfeeding < 36] = 3
ref_data$mean_brfeed_duration_cat[ref_data$avg_dur_breastfeeding >= 36] = 4

test_data$mean_brfeed_duration_cat = test_data$avg_dur_breastfeeding
test_data$mean_brfeed_duration_cat[test_data$avg_dur_breastfeeding < 12] = 1
test_data$mean_brfeed_duration_cat[test_data$avg_dur_breastfeeding >= 12 
                                  & test_data$avg_dur_breastfeeding < 24] = 2
test_data$mean_brfeed_duration_cat[test_data$avg_dur_breastfeeding >= 24 
                                  & test_data$avg_dur_breastfeeding < 36] = 3
test_data$mean_brfeed_duration_cat[test_data$avg_dur_breastfeeding >= 36] = 4

table(test_data$mean_brfeed_duration_cat)
sum(is.na(test_data$mean_brfeed_duration_cat))
```

## BMI
```{r}
ref_data$bmicat = ref_data$BMI
ref_data$bmicat[ref_data$BMI < 25] = 1
ref_data$bmicat[ref_data$BMI >= 25 & ref_data$BMI < 30] = 2
ref_data$bmicat[ref_data$BMI >= 30] = 3

test_data$bmicat = test_data$BMI
test_data$bmicat[test_data$BMI < 25] = 1
test_data$bmicat[test_data$BMI >= 25 & test_data$BMI < 30] = 2
test_data$bmicat[test_data$BMI >= 30] = 3

table(test_data$bmicat )
sum(is.na(test_data$bmicat))
```

## Height
```{r}
ref_data$catht = ref_data$Avg_Height
ref_data$catht[ref_data$Avg_Height <= 150] = 0
ref_data$catht[ref_data$Avg_Height > 150 & 
                 ref_data$Avg_Height <= 155] = 1
ref_data$catht[ref_data$Avg_Height > 155 & 
                 ref_data$Avg_Height <= 160] = 2
ref_data$catht[ref_data$Avg_Height > 160] = 3

test_data$catht = test_data$Avg_Height
test_data$catht[test_data$Avg_Height <= 150] = 0
test_data$catht[test_data$Avg_Height > 150 & 
                 test_data$Avg_Height <= 155] = 1
test_data$catht[test_data$Avg_Height > 155 & 
                 test_data$Avg_Height <= 160] = 2
test_data$catht[test_data$Avg_Height > 160] = 3

table(test_data$catht)
sum(is.na(test_data$catht))
```

## WHR
```{r}
ref_data$ratio = ref_data$waist_to_hip
ref_data$ratio[ref_data$waist_to_hip <= 0.84] = 0
ref_data$ratio[ref_data$waist_to_hip > 0.84 & 
                 ref_data$waist_to_hip <= 0.94] = 1
ref_data$ratio[ref_data$waist_to_hip > 0.94] = 2

test_data$ratio = test_data$waist_to_hip
test_data$ratio[test_data$waist_to_hip <= 0.84] = 0
test_data$ratio[test_data$waist_to_hip > 0.84 & 
                 test_data$waist_to_hip <= 0.94] = 1
test_data$ratio[test_data$waist_to_hip > 0.94] = 2

table(test_data$ratio)
sum(is.na(test_data$ratio))
```

## Age at menopause
```{r}
ref_data$catmenopause = ref_data$Age_at_men_stop
ref_data$catmenopause[ref_data$Age_at_men_stop <= 40] = 0
ref_data$catmenopause[ref_data$Age_at_men_stop > 40 & 
                        ref_data$Age_at_men_stop <= 45] = 1
ref_data$catmenopause[ref_data$Age_at_men_stop > 45 & 
                        ref_data$Age_at_men_stop <= 50] = 2
ref_data$catmenopause[ref_data$Age_at_men_stop > 50] = 3

test_data$catmenopause = test_data$Age_at_men_stop
test_data$catmenopause[test_data$Age_at_men_stop <= 40] = 0
test_data$catmenopause[test_data$Age_at_men_stop > 40 & 
                        test_data$Age_at_men_stop <= 45] = 1
test_data$catmenopause[test_data$Age_at_men_stop > 45 & 
                        test_data$Age_at_men_stop <= 50] = 2
test_data$catmenopause[test_data$Age_at_men_stop > 50] = 3

table(test_data$catmenopause)
sum(is.na(test_data$catmenopause))
```

## Reference dataset for premenopausal women
```{r}
ref_data_premeno = ref_data[ref_data$mentruation_status == 1,]
ref_data_premeno = ref_data_premeno[,c("age","ageatfirstfulltermpreg_cat",
                                       "interval_menarche_firstbirth_cat",
                                       "fulltermpreg_cat",
                                       "oc", "mean_brfeed_duration_cat",
                                       "totalabortion_cat",
                                       "totalmiscarriage_cat",
                                       "catht", "bmicat", "ratio")]

ref_data_premeno = na.omit(ref_data_premeno)

#Study entry age for the premenopausal women
age_premeno = ref_data_premeno$age

ref_data_premeno = ref_data_premeno[,!names(ref_data_premeno) %in% "age"]

dim(ref_data_premeno)


##Setting up the test data premenopausal women
test_data_premeno = test_data[test_data$mentruation_status == 1,]
test_data_premeno = test_data_premeno[,c("age","ageatfirstfulltermpreg_cat",
                                       "interval_menarche_firstbirth_cat",
                                       "fulltermpreg_cat",
                                       "oc", "mean_brfeed_duration_cat",
                                       "totalabortion_cat",
                                       "totalmiscarriage_cat",
                                       "catht", "bmicat", "ratio")]

#test_data_premeno = na.omit(test_data_premeno)

#Study entry age for the premenopausal women
age_premeno_test = test_data_premeno$age

test_data_premeno = test_data_premeno[,!names(test_data_premeno) %in% "age"]
```

## Model formula for premenopausal women
```{r}
model.formula.premeno = br ~ as.factor(ageatfirstfulltermpreg_cat) + 
  interval_menarche_firstbirth_cat + as.factor(fulltermpreg_cat) + 
  oc + as.factor(mean_brfeed_duration_cat) +
  as.factor(totalabortion_cat) + as.factor(totalmiscarriage_cat) + 
  as.factor(catht) + as.factor(bmicat) * as.factor(ratio)
```

## Model covariate information for premenopausal women
```{r}
v1 = list()
v1$name = "ageatfirstfulltermpreg_cat"
v1$type = "factor"
v1$levels = c("1","2","3","4","5")

v2 = list()
v2$name = "interval_menarche_firstbirth_cat"
v2$type = "continuous"

v3 = list()
v3$name = "fulltermpreg_cat"
v3$type = "factor"
v3$levels = c("1","2","3","4")

v4 = list()
v4$name = "oc"
v4$type = "continuous"

v5 = list()
v5$name = "mean_brfeed_duration_cat"
v5$type = "factor"
v5$levels = c("1","2","3","4")

v6 = list()
v6$name = "totalabortion_cat"
v6$type = "factor"
v6$levels = c("0","1","2")

v7 = list()
v7$name = "totalmiscarriage_cat"
v7$type = "factor"
v7$levels = c("0","1","2")

v8 = list()
v8$name = "catht"
v8$type = "factor"
v8$levels = c("0","1","2","3")

v9 = list()
v9$name = "bmicat"
v9$type = "factor"
v9$levels = c("1","2","3")

v10 = list()
v10$name = "ratio"
v10$type = "factor"
v10$levels = c("0","1","2")

cov.info.premeno = list(v1,v2,v3,v4,v5,v6,v7,v8,v9,v10)
```

##Model log relative risks for premenopausal women
```{r}
logRR_premeno = coef(model_premeno)[-c(1:10)]
```


## Absolute risk distribution for premenopausal women
```{r}
## 5-year risk
output_premeno_5y = computeAbsoluteRisk(model.formula = model.formula.premeno,
                                        model.cov.info = cov.info.premeno,
                                        model.snp.info = NULL,
                                        model.log.RR = logRR_premeno,
                                        model.ref.dataset = ref_data_premeno,
                                        model.disease.incidence.rates = bc_inci,
                                        model.competing.incidence.rates = mort_inci,
                                        model.bin.fh.name = NA,
                                        apply.age.start = age_premeno,
                                        apply.age.interval.length = 5,
                                        apply.cov.profile = ref_data_premeno,
                                        apply.snp.profile = NULL,
                                        return.lp = TRUE,
                                        return.refs.risk = TRUE)

output_premeno_test_5y = computeAbsoluteRisk(model.formula = model.formula.premeno,
                                             model.cov.info = cov.info.premeno,
                                             model.snp.info = NULL,
                                             model.log.RR = logRR_premeno,
                                             model.ref.dataset = ref_data_premeno,
                                             model.disease.incidence.rates = bc_inci,
                                             model.competing.incidence.rates = mort_inci,
                                             model.bin.fh.name = NA,
                                             apply.age.start = age_premeno_test,
                                             apply.age.interval.length = 5,
                                             apply.cov.profile = test_data_premeno,
                                             apply.snp.profile = NULL,
                                             return.lp = TRUE,
                                             return.refs.risk = TRUE)

summary(output_premeno_5y$risk * 100)
summary(output_premeno_test_5y$risk * 100)

#pdf("Five_year_AR_premenopausal.pdf")
plot(density(output_premeno_5y$risk * 100),
     xlim = c(0,4), ylim = c(0,4),
     main = "Premenopausal Women",
     xlab = "5-year absolute risk (%)", lwd = 3)
abline(v = output_premeno_test_5y$risk * 100, col = "red")
text(x = 3, y = 1, 
     labels = paste("Mean AR (%):", 
                    round(mean(output_premeno_5y$risk * 100),1)))
#dev.off()

## 10-year risk
output_premeno_10y = computeAbsoluteRisk(model.formula = model.formula.premeno,
                                         model.cov.info = cov.info.premeno,
                                         model.snp.info = NULL,
                                         model.log.RR = logRR_premeno,
                                         model.ref.dataset = ref_data_premeno,
                                         model.disease.incidence.rates = bc_inci,
                                         model.competing.incidence.rates = mort_inci,
                                         model.bin.fh.name = NA,
                                         apply.age.start = age_premeno,
                                         apply.age.interval.length = 10,
                                         apply.cov.profile = ref_data_premeno,
                                         apply.snp.profile = NULL,
                                         return.lp = TRUE,
                                         return.refs.risk = TRUE)

output_premeno_test_10y = computeAbsoluteRisk(model.formula = model.formula.premeno,
                                              model.cov.info = cov.info.premeno,
                                              model.snp.info = NULL,
                                              model.log.RR = logRR_premeno,
                                              model.ref.dataset = ref_data_premeno,
                                              model.disease.incidence.rates = bc_inci,
                                              model.competing.incidence.rates = mort_inci,
                                              model.bin.fh.name = NA,
                                              apply.age.start = age_premeno_test,
                                              apply.age.interval.length = 10,
                                              apply.cov.profile = test_data_premeno,
                                              apply.snp.profile = NULL,
                                              return.lp = TRUE,
                                              return.refs.risk = TRUE)

summary(output_premeno_10y$risk * 100)
summary(output_premeno_test_10y$risk * 100)

#pdf("Five_year_AR_premenopausal.pdf")
plot(density(output_premeno_10y$risk * 100),
     xlim = c(0,4), ylim = c(0,2),
     main = "Premenopausal Women",
     xlab = "10-year absolute risk (%)", lwd = 3)
abline(v = output_premeno_test_10y$risk * 100, col = "red")
text(x = 3, y = 1, 
     labels = paste("Mean AR (%):", 
                    round(mean(output_premeno_10y$risk * 100),1)))
#dev.off()
```


## Reference dataset for postmenopausal women
```{r}
ref_data_postmeno = ref_data[ref_data$mentruation_status == 2,]
ref_data_postmeno = ref_data_postmeno[,c("age","ageatfirstfulltermpreg_cat",
                                         "interval_menarche_firstbirth_cat",
                                         "fulltermpreg_cat", "catmenopause",
                                         "oc", "mean_brfeed_duration_cat",
                                         "totalabortion_cat",
                                         "totalmiscarriage_cat",
                                         "catht", "bmicat", "ratio")]

ref_data_postmeno = na.omit(ref_data_postmeno)

#Study entry age for postmenopausal women
age_postmeno = ref_data_postmeno$age

ref_data_postmeno = ref_data_postmeno[,!names(ref_data_postmeno) %in% "age"]


test_data_postmeno = test_data[test_data$mentruation_status == 2,]
test_data_postmeno = test_data_postmeno[,c("age","ageatfirstfulltermpreg_cat",
                                         "interval_menarche_firstbirth_cat",
                                         "fulltermpreg_cat", "catmenopause",
                                         "oc", "mean_brfeed_duration_cat",
                                         "totalabortion_cat",
                                         "totalmiscarriage_cat",
                                         "catht", "bmicat", "ratio")]

#test_data_postmeno = na.omit(test_data_postmeno)

#Study entry age for postmenopausal women
age_postmeno_test = test_data_postmeno$age

test_data_postmeno = test_data_postmeno[,!names(test_data_postmeno) %in% "age"]
```

## Model formula for postmenopausal women
```{r}
model.formula.postmeno = br ~ as.factor(ageatfirstfulltermpreg_cat) + 
  interval_menarche_firstbirth_cat + as.factor(fulltermpreg_cat) + 
  as.factor(catmenopause) + oc + as.factor(mean_brfeed_duration_cat) + 
  as.factor(totalabortion_cat) + as.factor(totalmiscarriage_cat) + 
  as.factor(catht) + as.factor(bmicat) * as.factor(ratio)
```

## Model covariate information for postmenopausal women
```{r}
v1 = list()
v1$name = "ageatfirstfulltermpreg_cat"
v1$type = "factor"
v1$levels = c("1","2","3","4","5")

v2 = list()
v2$name = "interval_menarche_firstbirth_cat"
v2$type = "continuous"

v3 = list()
v3$name = "fulltermpreg_cat"
v3$type = "factor"
v3$levels = c("1","2","3","4")


v4 = list()
v4$name = "catmenopause"
v4$type = "factor"
v4$levels = c("0","1","2","3")

v5 = list()
v5$name = "oc"
v5$type = "continuous"

v6 = list()
v6$name = "mean_brfeed_duration_cat"
v6$type = "factor"
v6$levels = c("1","2","3","4")

v7 = list()
v7$name = "totalabortion_cat"
v7$type = "factor"
v7$levels = c("0","1","2")

v8 = list()
v8$name = "totalmiscarriage_cat"
v8$type = "factor"
v8$levels = c("0","1","2")

v9 = list()
v9$name = "catht"
v9$type = "factor"
v9$levels = c("0","1","2","3")

v10 = list()
v10$name = "bmicat"
v10$type = "factor"
v10$levels = c("1","2","3")

v11 = list()
v11$name = "ratio"
v11$type = "factor"
v11$levels = c("0","1","2")


cov.info.postmeno = list(v1,v2,v3,v4,v5,v6,v7,v8,v9,v10,v11)
```

## Model log relative risks for postmenopausal women
```{r}
logRR_postmeno = coef(model_postmeno)[-c(1:10)]
```

## Absolute risk distribution for postmenopausal women
```{r}
## 5-year risk
output_postmeno_5y = computeAbsoluteRisk(model.formula = model.formula.postmeno,
                                         model.cov.info = cov.info.postmeno,
                                         model.snp.info = NULL,
                                         model.log.RR = logRR_postmeno,
                                         model.ref.dataset = ref_data_postmeno,
                                         model.disease.incidence.rates = bc_inci,
                                         model.competing.incidence.rates = mort_inci,
                                         model.bin.fh.name = NA,
                                         apply.age.start = age_postmeno,
                                         apply.age.interval.length = 5,
                                         apply.cov.profile = ref_data_postmeno,
                                         apply.snp.profile = NULL,
                                         return.lp = TRUE,
                                         return.refs.risk = TRUE)

output_postmeno_test_5y = computeAbsoluteRisk(model.formula = model.formula.postmeno,
                                              model.cov.info = cov.info.postmeno,
                                              model.snp.info = NULL,
                                              model.log.RR = logRR_postmeno,
                                              model.ref.dataset = ref_data_postmeno,
                                              model.disease.incidence.rates = bc_inci,
                                              model.competing.incidence.rates = mort_inci,
                                              model.bin.fh.name = NA,
                                              apply.age.start = age_postmeno_test,
                                              apply.age.interval.length = 5,
                                              apply.cov.profile = test_data_postmeno,
                                              apply.snp.profile = NULL,
                                              return.lp = TRUE,
                                              return.refs.risk = TRUE)

summary(output_postmeno_5y$risk * 100)
summary(output_postmeno_test_5y$risk * 100)

#pdf("Five_year_AR_postmenopausal.pdf")
plot(density(output_postmeno_5y$risk * 100),
     xlim = c(0,4), ylim = c(0,4),
     main = "Postmenopausal Women",
     xlab = "5-year absolute risk (%)", lwd = 3)
abline(v = output_postmeno_test_5y$risk * 100, col = "red")
text(x = 3, y = 1, 
     labels = paste("Mean AR (%):", 
                    round(mean(output_postmeno_5y$risk * 100),1)))
#dev.off()

##10-year risk
output_postmeno_10y = computeAbsoluteRisk(model.formula = model.formula.postmeno,
                                          model.cov.info = cov.info.postmeno,
                                          model.snp.info = NULL,
                                          model.log.RR = logRR_postmeno,
                                          model.ref.dataset = ref_data_postmeno,
                                          model.disease.incidence.rates = bc_inci,
                                          model.competing.incidence.rates = mort_inci,
                                          model.bin.fh.name = NA,
                                          apply.age.start = age_postmeno,
                                          apply.age.interval.length = 10,
                                          apply.cov.profile = ref_data_postmeno,
                                          apply.snp.profile = NULL,
                                          return.lp = TRUE,
                                          return.refs.risk = TRUE)

output_postmeno_test_10y = computeAbsoluteRisk(model.formula = model.formula.postmeno,
                                               model.cov.info = cov.info.postmeno,
                                               model.snp.info = NULL,
                                               model.log.RR = logRR_postmeno,
                                               model.ref.dataset = ref_data_postmeno,
                                               model.disease.incidence.rates = bc_inci,
                                               model.competing.incidence.rates = mort_inci,
                                               model.bin.fh.name = NA,
                                               apply.age.start = age_postmeno_test,
                                               apply.age.interval.length = 10,
                                               apply.cov.profile = test_data_postmeno,
                                               apply.snp.profile = NULL,
                                               return.lp = TRUE,
                                               return.refs.risk = TRUE)

summary(output_postmeno_10y$risk * 100)
summary(output_postmeno_test_10y$risk * 100)

#pdf("Five_year_AR_postmenopausal.pdf")
plot(density(output_postmeno_10y$risk * 100),
     xlim = c(0,4), ylim = c(0,2),
     main = "Postmenopausal Women",
     xlab = "10-year absolute risk (%)", lwd = 3)
abline(v = output_postmeno_test_10y$risk * 100, col = "red")
text(x = 3.5, y = 1, 
     labels = paste("Mean AR (%):", 
                    round(mean(output_postmeno_10y$risk * 100),1)))
#dev.off()

```

## Risk projections in the Indian population

Load the census data
```{r}
setwd("/Users/anaghchattopadhyay/Documents/Nilanjan\ Project/TMC_BCProject/Data")

pop_data = read.xlsx("Indian population numbers by Age (Census-2011).xlsx")
names(pop_data) = c("Age","Population")
pop_data = pop_data[-1,]


pop_data$Age = as.numeric(pop_data$Age)
pop_data$Population = as.numeric(pop_data$Population)

pop_data = na.omit(pop_data)
```

Age at menopause category definitions
0: <= 40 years
1: 41-45 years
2: 46-50 years
3: >= 51 years

Look at the distribution of age at menopause to determine the age cutoff
```{r}
prop.table(table(ref_data_postmeno$catmenopause))
```
Median age at menopause roughly 45 years. Premenopausal group defined as 25-45 years and postmenopausal group defined as 45-68 years

Population size in the premenopausal group:
```{r}

pop_total_premeno = sum(pop_data$Population[pop_data$Age >= 25 & 
                                              pop_data$Age < 45], na.rm = TRUE)
pop_total_premeno
```

5-year risk projections

Approximate number of future cases (within 5 years) in the premenopausal group:
```{r}
cases_total_premeno = round(pop_total_premeno * round(mean(output_premeno_5y$risk),3),0)
cases_total_premeno
```


Percentage of premenopausal women above 1\% 5-year absolute risk
```{r}
round(mean(output_premeno_5y$risk >= 0.01) * 100,1)
```


Number of premenopausal women above 1\% 5-year absolute risk
```{r}
pop_total_above_0.01AR_premeno = round(round(mean(output_premeno_5y$risk >= 0.01),3) * 
                                         pop_total_premeno, 0)
pop_total_above_0.01AR_premeno
```



Number of future cases in the group above 1\% 5-year absolute risk
```{r}
cases_total_above_0.01AR_premeno = 
  round(round(mean(output_premeno_5y$risk[output_premeno_5y$risk >= 0.01]),3) * 
          pop_total_above_0.01AR_premeno,0)
cases_total_above_0.01AR_premeno
```
Percentage of future cases in the group above 1\% 5-year absolute risk
```{r}
round((cases_total_above_0.01AR_premeno/cases_total_premeno)*100,1)
```

Population size in the postmenopausal group:

```{r}
pop_total_postmeno = sum(pop_data$Population[pop_data$Age >= 45 & 
                                              pop_data$Age <= 68], na.rm = TRUE)
pop_total_postmeno
```

Approximate number of future cases (within 5 years) in the postmenopausal group:
```{r}
cases_total_postmeno = round(pop_total_postmeno * round(mean(output_postmeno_5y$risk),3),0)
cases_total_postmeno
```

Percentage of postmenopausal women above 1\% 5-year absolute risk
```{r}
round(mean(output_postmeno_5y$risk >= 0.01) * 100,1)
```

Number of postmenopausal women above 1\% 5-year absolute risk
```{r}
pop_total_above_0.01AR_postmeno = round(round(mean(output_postmeno_5y$risk >= 0.01),3) * 
                                          pop_total_postmeno, 0)
pop_total_above_0.01AR_postmeno
```

Number of future cases in the group above 1\% 5-year absolute risk
```{r}
cases_total_above_0.01AR_postmeno = 
  round(round(mean(output_postmeno_5y$risk[output_postmeno_5y$risk >= 0.01]),3) * 
          pop_total_above_0.01AR_postmeno,0)
cases_total_above_0.01AR_postmeno
```
Percentage of future cases in the group above 1\% 5-year absolute risk
```{r}
round((cases_total_above_0.01AR_postmeno/cases_total_postmeno)*100,1)
```



10-year risk projections

Approximate number of future cases (within 10 years) in the premenopausal group:
```{r}
cases_total_premeno = round(pop_total_premeno * round(mean(output_premeno_10y$risk),3),0)
cases_total_premeno
```


Percentage of premenopausal women above 2\% 10-year absolute risk
```{r}
round(mean(output_premeno_10y$risk >= 0.02) * 100,1)
```


Number of premenopausal women above 2\% 10-year absolute risk
```{r}
pop_total_above_0.02AR_premeno = round(round(mean(output_premeno_10y$risk >= 0.02),3) * 
                                         pop_total_premeno, 0)
pop_total_above_0.02AR_premeno
```

Number of future cases in the group above 2\% 10-year absolute risk
```{r}
cases_total_above_0.02AR_premeno = 
  round(round(mean(output_premeno_10y$risk[output_premeno_10y$risk >= 0.02]),3) * 
          pop_total_above_0.02AR_premeno,0)
cases_total_above_0.02AR_premeno
```
Percentage of future cases in the group above 2\% 10-year absolute risk
```{r}
round((cases_total_above_0.02AR_premeno/cases_total_premeno)*100,1)
```

Population size in the postmenopausal group:

```{r}
pop_total_postmeno = sum(pop_data$Population[pop_data$Age >= 45 & 
                                              pop_data$Age <= 68], na.rm = TRUE)
pop_total_postmeno
```

Approximate number of future cases (within 10 years) in the postmenopausal group:
```{r}
cases_total_postmeno = round(pop_total_postmeno * round(mean(output_postmeno_10y$risk),3),0)
cases_total_postmeno
```

Percentage of postmenopausal women above 2\% 10-year absolute risk
```{r}
round(mean(output_postmeno_10y$risk >= 0.02) * 100,1)
```

Number of postmenopausal women above 2\% 10-year absolute risk
```{r}
pop_total_above_0.02AR_postmeno = round(round(mean(output_postmeno_10y$risk >= 0.02),3) * 
                                          pop_total_postmeno, 0)
pop_total_above_0.02AR_postmeno
```

Number of future cases in the group above 2\% 10-year absolute risk
```{r}
cases_total_above_0.02AR_postmeno = 
  round(round(mean(output_postmeno_10y$risk[output_postmeno_10y$risk >= 0.02]),3) * 
          pop_total_above_0.02AR_postmeno,0)
cases_total_above_0.02AR_postmeno
```
Percentage of future cases in the group above 1\% 5-year absolute risk
```{r}
round((cases_total_above_0.02AR_postmeno/cases_total_postmeno)*100,1)
```


