---
title: "Breast Cancer Risk Prediction in Indian Population"
author: "Parichoy Pal Choudhury"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

## Install the required packages and libraries
```{r, message = FALSE}
# if (!require("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# 
# BiocManager::install("iCARE")
library(readstata13)
library(plotrix)
library(dplyr)
library(gdata)
library(openxlsx)
library(brglm2)
library(naniar)
library(iCARE)
library(purrr)
library(janitor)
```

## Set the working directory to the location of the data and read the data
```{r}

setwd("/Users/anaghchattopadhyay/Documents/Nilanjan\ Project/TMC_BCProject/Data")
data = read.dta13("/Users/anaghchattopadhyay/Documents/Nilanjan\ Project/TMC_BCProject/Data/Questionnaire_28.04.2017.dta",
                  generate.factors = TRUE, nonint.factors = TRUE)
data_new_meta = read.xlsx("/Users/anaghchattopadhyay/Documents/Nilanjan\ Project/TMC_BCProject/Data/Breast\ cancer\ data\ -\ CCE.TMC/Meta_Data.xlsx")
data_new_validation = read.xlsx("/Users/anaghchattopadhyay/Documents/Nilanjan\ Project/TMC_BCProject/Data/Breast\ cancer\ data\ -\ CCE.TMC/Breast\ cancer\ _Validation\ data\ set.xlsx")
data_DOE = read.xlsx("/Users/anaghchattopadhyay/Documents/Nilanjan\ Project/TMC_BCProject/Data/Breast\ cancer\ data\ -\ CCE.TMC/data_new_meta_ids.xlsx")
data_mmgy = read.xlsx("/Users/anaghchattopadhyay/Documents/Nilanjan\ Project/TMC_BCProject/Data/Breast\ cancer\ data\ -\ CCE.TMC/mamogram\ data\ _24.5.2025.xlsx")
#data_covariate_wise
data_new_alcohol = read.xlsx("/Users/anaghchattopadhyay/Documents/Nilanjan\ Project/TMC_BCProject/Data/Breast\ cancer\ data\ -\ CCE.TMC/28.03.2025_Risk\ prediction\ model\ data/Alcohol/Alcohol_25.09.2023.xlsx")
data_new_alcohol_exlusive = read.xlsx("/Users/anaghchattopadhyay/Documents/Nilanjan\ Project/TMC_BCProject/Data/Breast\ cancer\ data\ -\ CCE.TMC/28.03.2025_Risk\ prediction\ model\ data/Alcohol/Alcohol_exclusive_07.10.2023.xlsx")
data_new_alcohol_inclusive = read.xlsx("/Users/anaghchattopadhyay/Documents/Nilanjan\ Project/TMC_BCProject/Data/Breast\ cancer\ data\ -\ CCE.TMC/28.03.2025_Risk\ prediction\ model\ data/Alcohol/Alcohol_inclusive_07.10.2023.xlsx")
data_new_anthropometry_ank = read.xlsx("/Users/anaghchattopadhyay/Documents/Nilanjan\ Project/TMC_BCProject/Data/Breast\ cancer\ data\ -\ CCE.TMC/28.03.2025_Risk\ prediction\ model\ data/Anthropometry/Anthropometry_13072023\ -\ Ankita.xlsx")
data_new_anthropometry = read.xlsx("/Users/anaghchattopadhyay/Documents/Nilanjan\ Project/TMC_BCProject/Data/Breast\ cancer\ data\ -\ CCE.TMC/28.03.2025_Risk\ prediction\ model\ data/Anthropometry/Anthropometry_13072023.xlsx")
data_new_chewing = read.xlsx("/Users/anaghchattopadhyay/Documents/Nilanjan\ Project/TMC_BCProject/Data/Breast\ cancer\ data\ -\ CCE.TMC/28.03.2025_Risk\ prediction\ model\ data/Chewing/Chewing_inclusive_03.07.2024_final.xlsx")
data_new_demography = read.xlsx("/Users/anaghchattopadhyay/Documents/Nilanjan\ Project/TMC_BCProject/Data/Breast\ cancer\ data\ -\ CCE.TMC/28.03.2025_Risk\ prediction\ model\ data/Demographics/Demography_details_12034_17062023.xlsx")
data_new_FH1 = read.xlsx("/Users/anaghchattopadhyay/Documents/Nilanjan\ Project/TMC_BCProject/Data/Breast\ cancer\ data\ -\ CCE.TMC/28.03.2025_Risk\ prediction\ model\ data/Family_history_of_cancer/Fam_hist_CA_Br_core_10.08.2023..xlsx")
data_new_FH = read.xlsx("/Users/anaghchattopadhyay/Documents/Nilanjan\ Project/TMC_BCProject/Data/Breast\ cancer\ data\ -\ CCE.TMC/28.03.2025_Risk\ prediction\ model\ data/Family_history_of_cancer/family_history_of_cancer_10.08.2023.xlsx")
data_new_PMH = read.xlsx("/Users/anaghchattopadhyay/Documents/Nilanjan\ Project/TMC_BCProject/Data/Breast\ cancer\ data\ -\ CCE.TMC/28.03.2025_Risk\ prediction\ model\ data/Personal_medical/Personl_medical_history_27.03.2024_final.xlsx")
data_new_reproductive = read.xlsx("/Users/anaghchattopadhyay/Documents/Nilanjan\ Project/TMC_BCProject/Data/Breast\ cancer\ data\ -\ CCE.TMC/28.03.2025_Risk\ prediction\ model\ data/Reproductive/Reproductive_25.07.2023_final.xlsx")
data_new_residence1 = read.xlsx("/Users/anaghchattopadhyay/Documents/Nilanjan\ Project/TMC_BCProject/Data/Breast\ cancer\ data\ -\ CCE.TMC/28.03.2025_Risk\ prediction\ model\ data/Residential_details/Last_5_3_year_RU_05.10.2023.xlsx")
data_new_residence2 = read.xlsx("/Users/anaghchattopadhyay/Documents/Nilanjan\ Project/TMC_BCProject/Data/Breast\ cancer\ data\ -\ CCE.TMC/28.03.2025_Risk\ prediction\ model\ data/Residential_details/Last_5_3_year_state_04.10.2023.xlsx")
data_new_residence3 = read.xlsx("/Users/anaghchattopadhyay/Documents/Nilanjan\ Project/TMC_BCProject/Data/Breast\ cancer\ data\ -\ CCE.TMC/28.03.2025_Risk\ prediction\ model\ data/Residential_details/POB_duration_12.09.2023.xlsx")
data_new_residence4 = read.xlsx("/Users/anaghchattopadhyay/Documents/Nilanjan\ Project/TMC_BCProject/Data/Breast\ cancer\ data\ -\ CCE.TMC/28.03.2025_Risk\ prediction\ model\ data/Residential_details/Residential_birth_current_final_17.06.2023.xlsx")
data_new_smoking = read.xlsx("/Users/anaghchattopadhyay/Documents/Nilanjan\ Project/TMC_BCProject/Data/Breast\ cancer\ data\ -\ CCE.TMC/28.03.2025_Risk\ prediction\ model\ data/Smoking/Smoking_Incl_24.6.2024.xlsx")
```

```{r}
# ──────────────────────────────────────────────────────────────────────────────
# Assumptions (same as before):
# 1. data_DOE has columns “ID” and “DateOfInterview” (of class Date).
# 2. data_mmgy has column “ID” (subset of data_DOE$ID).
# ──────────────────────────────────────────────────────────────────────────────

# 1. Coerce DateOfInterview to Date if needed
if (!inherits(data_DOE$DateOfInterview, "Date")) {
  data_DOE$DateOfInterview <- as.Date(
    data_DOE$DateOfInterview,
    format = "%Y-%m-%d"  # adjust if your dates use a different format
  )
}

# 2. Subset data_DOE to just those with mammography records
doe_mmgy <- data_DOE[data_DOE$ID %in% data_mmgy$ID, ]

# 3. Order those mammography‐linked rows by DateOfInterview (newest first)
ord_mmgy_desc <- order(doe_mmgy$DateOfInterview, decreasing = TRUE)

# 4. Take the 400 latest mammography cases as validation
val_mmgy      <- doe_mmgy[ord_mmgy_desc[1:400], ]

# 5. Grab their IDs
val_ids       <- val_mmgy$ID

# 6. Define the new validation_set (all rows of data_DOE whose ID is in val_ids)
validation_set <- data_DOE[data_DOE$ID %in% val_ids, ]

# 7. Define the new training_set (all other rows of data_DOE)
training_set   <- data_DOE[!data_DOE$ID %in% val_ids, ]

```


```{r}
dataset_list <- list(
  alcohol = data_new_alcohol,
  alcohol_exclusive = data_new_alcohol_exlusive,
  alcohol_inclusive = data_new_alcohol_inclusive,
  anthropometry_ank = data_new_anthropometry_ank,
  anthropometry = data_new_anthropometry,
  chewing = data_new_chewing,
  demography = data_new_demography,
  FH1 = data_new_FH1,
  FH2 = data_new_FH,
  PMH = data_new_PMH,
  reproductive = data_new_reproductive,
  residence1 = data_new_residence1,
  residence2 = data_new_residence2,
  residence3 = data_new_residence3,
  residence4 = data_new_residence4,
  smoking = data_new_smoking
)

dataset_list <- map(dataset_list, ~clean_names(.x))
all_columns <- c("id")
repeated_columns <- list()
final_data <- NULL

for (name in names(dataset_list)) {
  df <- dataset_list[[name]]
  
  # Remove duplicated columns (already present in previous datasets except 'id')
  new_cols <- setdiff(names(df), "id")
  duplicate_cols <- intersect(new_cols, all_columns)
  
  if (length(duplicate_cols) > 0) {
    repeated_columns[[name]] <- duplicate_cols
    df <- df %>% select(-all_of(duplicate_cols))
  }
  
  all_columns <- unique(c(all_columns, names(df)))
  
  # First iteration
  if (is.null(final_data)) {
    final_data <- df
  } else {
    final_data <- full_join(final_data, df, by = "id")
  }
}

cat("🔁 Repeated columns (not bound multiple times):\n")
print(repeated_columns)

cat("\nFinal merged dataset dimensions:\n")
print(dim(final_data))

final_data$totalabortion_cat <- data_DOE$Count_ab[
  match(final_data$id, data_DOE$ID)
]
final_data$totalmiscarriage <- data_DOE$Count_mis[
  match(final_data$id, data_DOE$ID)
]

final_data <- final_data[ final_data$id %in% training_set$ID, ]
```

## Categories for age at first full term pregnancy
```{r}
data$ageatfirstfulltermpreg_cat = data$ageatfirstfulltermpreg
data$ageatfirstfulltermpreg_cat[which(data$ageatfirstfulltermpreg < 20)] = 1
data$ageatfirstfulltermpreg_cat[which(data$ageatfirstfulltermpreg == 20 
                                      | data$ageatfirstfulltermpreg == 21)] = 2
data$ageatfirstfulltermpreg_cat[which(data$ageatfirstfulltermpreg == 22 
                                      | data$ageatfirstfulltermpreg == 23)] = 3
data$ageatfirstfulltermpreg_cat[which(data$ageatfirstfulltermpreg == 24 
                                      | data$ageatfirstfulltermpreg == 25)] = 4
data$ageatfirstfulltermpreg_cat[which(data$ageatfirstfulltermpreg >= 26)] = 5
```

## Categories for interval between menarche and full term pregnancy
```{r}
data$interval_menarche_firstbirth = data$ageatfirstfulltermpreg - 
  data$whtagestartmensurating

data$interval_menarche_firstbirth_cat = data$interval_menarche_firstbirth
data$interval_menarche_firstbirth_cat[data$interval_menarche_firstbirth < 10] = 0
data$interval_menarche_firstbirth_cat[data$interval_menarche_firstbirth >= 10] = 1

```

## Categories for number of full term pregnancies
```{r}
data$fulltermpreg_cat = data$fulltermpreg
data$fulltermpreg_cat[data$fulltermpreg >= 4] = 4
```

## Categories for number of induced abortions
```{r}
data$totalabortion_cat = data$totalabortion
data$totalabortion_cat[data$totalabortion >= 2] = 2
```

## Categories for number of spontaneous abortions
```{r}
data$totalmiscarriage_cat = data$totalmiscarriage
data$totalmiscarriage_cat[data$totalmiscarriage >= 2] = 2
```

## Categories for duration of breastfeeding
```{r}
data$mean_brfeed_duration_cat = data$mean_brfeed_duration
data$mean_brfeed_duration_cat[data$mean_brfeed_duration < 12] = 1
data$mean_brfeed_duration_cat[data$mean_brfeed_duration >= 12 
                              & data$mean_brfeed_duration < 24] = 2
data$mean_brfeed_duration_cat[data$mean_brfeed_duration >= 24 
                              & data$mean_brfeed_duration < 36] = 3
data$mean_brfeed_duration_cat[data$mean_brfeed_duration >= 36] = 4
```

## Categories for duration of diabetes
```{r}
data$diabetic_years_cat = data$diabetic_years
data$diabetic_years_cat[data$diabetes_YN == 0] = 0
data$diabetic_years_cat[data$diabetic_years <= 5] = 1
data$diabetic_years_cat[data$diabetic_years >= 6 
                        & data$diabetic_years <= 10] = 2
data$diabetic_years_cat[data$diabetic_years >= 11] = 3
```

## Collapsing the <=18.5 and 18.5-24.9 categories for BMI
```{r}
data$bmicat[data$bmicat == 0 | data$bmicat == 1] = 1
```

#######################################################################################
## Categories for age at first full term pregnancy
```{r}
final_data$ageatfirstfulltermpreg_cat = final_data$age_fftp
final_data$ageatfirstfulltermpreg_cat[which(final_data$ageatfirstfulltermpreg < 20)] = 1
final_data$ageatfirstfulltermpreg_cat[which(final_data$ageatfirstfulltermpreg == 20 
                                      | final_data$ageatfirstfulltermpreg == 21)] = 2
final_data$ageatfirstfulltermpreg_cat[which(final_data$ageatfirstfulltermpreg == 22 
                                      | final_data$ageatfirstfulltermpreg == 23)] = 3
final_data$ageatfirstfulltermpreg_cat[which(final_data$ageatfirstfulltermpreg == 24 
                                      | final_data$ageatfirstfulltermpreg == 25)] = 4
final_data$ageatfirstfulltermpreg_cat[which(final_data$ageatfirstfulltermpreg >= 26)] = 5
```

## Categories for interval between menarche and full term pregnancy
```{r}
final_data$interval_menarche_firstbirth = final_data$age_fftp - 
  final_data$age_menarche

final_data$interval_menarche_firstbirth_cat = final_data$interval_menarche_firstbirth
final_data$interval_menarche_firstbirth_cat[final_data$interval_menarche_firstbirth < 10] = 0
final_data$interval_menarche_firstbirth_cat[final_data$interval_menarche_firstbirth >= 10] = 1

```

## Categories for number of full term pregnancies
```{r}
final_data$fulltermpreg_cat <- ifelse(final_data$parous == 0, 0, final_data$parity)
final_data$fulltermpreg_cat[final_data$fulltermpreg_cat >= 4] <- 4
```

## Categories for number of induced abortions (Not present)
```{r}
final_data$totalabortion_cat = final_data$totalabortion
final_data$totalabortion_cat[final_data$totalabortion >= 2] = 2
```

## Categories for number of spontaneous abortions (Not present)
```{r}
final_data$totalmiscarriage_cat = final_data$totalmiscarriage
final_data$totalmiscarriage_cat[final_data$totalmiscarriage >= 2] = 2
```

## Categories for duration of breastfeeding
```{r}
final_data$mean_brfeed_duration_cat = final_data$breast_mos/final_data$fulltermpreg_cat
final_data$mean_brfeed_duration_cat[final_data$mean_brfeed_duration < 12] = 1
final_data$mean_brfeed_duration_cat[final_data$mean_brfeed_duration >= 12 
                              & final_data$mean_brfeed_duration < 24] = 2
final_data$mean_brfeed_duration_cat[final_data$mean_brfeed_duration >= 24 
                              & final_data$mean_brfeed_duration < 36] = 3
final_data$mean_brfeed_duration_cat[final_data$mean_brfeed_duration >= 36] = 4
```

## Categories for duration of diabetes
```{r}
final_data$diabetic_years_cat = final_data$diabetes_duration
final_data$diabetic_years_cat[final_data$diabetes_yn == 0] = 0
final_data$diabetic_years_cat[final_data$diabetes_duration <= 5] = 1
final_data$diabetic_years_cat[final_data$diabetes_duration >= 6 
                        & final_data$diabetes_duration <= 10] = 2
final_data$diabetic_years_cat[final_data$diabetes_duration >= 11] = 3
```

## 2D:4D
```{r}
library(ggplot2)
library(dplyr)

# Create 2D:4D ratios
final_data <- final_data %>%
  mutate(
    TwoD4D_left = left_index_finger / left_ring_finger,
    TwoD4D_right = right_index_finger / right_ring_finger
  )

# Histogram for TwoD4D_left
ggplot(final_data, aes(x = TwoD4D_left, fill = case_site)) +
  geom_histogram(alpha = 0.6, bins = 100, position = "identity") +
  scale_fill_manual(values = c("Control" = "#1f77b4", "Breast Ca." = "#d62728")) +
  labs(title = "Histogram of 2D:4D Ratio (Left Hand)", x = "2D:4D (Left)", y = "Count") +
  theme_minimal()

# Histogram for TwoD4D_right
ggplot(final_data, aes(x = TwoD4D_right, fill = case_site)) +
  geom_histogram(alpha = 0.6, bins = 100, position = "identity") +
  scale_fill_manual(values = c("Control" = "#1f77b4", "Breast Ca." = "#d62728")) +
  labs(title = "Histogram of 2D:4D Ratio (Right Hand)", x = "2D:4D (Right)", y = "Count") +
  theme_minimal()

final_data <- final_data %>%
  mutate(
    TwoD4D_left_quart_cat = factor(ntile(TwoD4D_left, 4),
                                   labels = c("Q1", "Q2", "Q3", "Q4")),
    TwoD4D_right_quint_cat = factor(ntile(TwoD4D_right, 4),
                                    labels = c("Q1", "Q2", "Q3", "Q4"))
  )


```

```{r}
# family_history_of_beo
final_data$family_sitory_of_beo <- as.integer(
  final_data$f_his_fst_bc == 1 |
  final_data$f_his_fst_oc == 1 |
  final_data$ca_end_yn == 1
)
```


```{r}
#############
final_data$Avg_Height = ifelse(is.na(final_data$height_1), final_data$height_2, 
                                ifelse(is.na(final_data$height_2), final_data$height_1, 
                                       0.5 * (final_data$height_1 + final_data$height_2)))

final_data$catht = final_data$Avg_Height
final_data$catht[final_data$Avg_Height <= 150] = 0
final_data$catht[final_data$Avg_Height > 150 & 
                 final_data$Avg_Height <= 155] = 1
final_data$catht[final_data$Avg_Height > 155 & 
                 final_data$Avg_Height <= 160] = 2
final_data$catht[final_data$Avg_Height > 160] = 3


###############
final_data$ratio = final_data$waist_to_hip_ratio
final_data$ratio[final_data$waist_to_hip <= 0.84] = 0
final_data$ratio[final_data$waist_to_hip > 0.84 & 
                 final_data$waist_to_hip <= 0.94] = 1
final_data$ratio[final_data$waist_to_hip > 0.94] = 2

################
final_data$case_site = as.factor(final_data$case_site)

################
final_data$last_5_yr_res_ru[!(final_data$last_5_yr_res_ru %in% c(1, 2))] <- NA

################
final_data$oc_ever[!(final_data$oc_ever %in% c(1, 0))] <- NA

#################
final_data$education_self <- ifelse(final_data$education_self > 4, 2, 1)

#################
final_data$catmenopause = final_data$mens_age_last
final_data$catmenopause[final_data$mens_age_last <= 40] = 0
final_data$catmenopause[final_data$mens_age_last > 40 & 
                        final_data$mens_age_last <= 45] = 1
final_data$catmenopause[final_data$mens_age_last > 45 & 
                        final_data$mens_age_last <= 50] = 2
final_data$catmenopause[final_data$mens_age_last > 50] = 3

##################
final_data$bmicat = final_data$bmi
final_data$bmicat[final_data$bmi < 25] = 1
final_data$bmicat[final_data$bmi >= 25 & final_data$bmi < 30] = 2
final_data$bmicat[final_data$bmi >= 30] = 3

###################

unique(final_data$case_site)

final_data$case_site <- factor(
  final_data$case_site,
  levels = c("Control", "Breast Ca."),
  labels = c("Control", "Breast Cancer Case")
)

final_data$br <- final_data$case_site
final_data$oc <- final_data$oc_ever
final_data$menopause <- factor(
  final_data$meno_stat,
  levels = c(1, 2),
  labels = c("Premenopausal", "Postmenopausal")
)

final_data$family_sitory_of_beo <- factor(
  final_data$family_sitory_of_beo,
  levels = c(0, 1),
  labels = c("No", "Yes")
)

```
#######################################################################################

```{r}
# Define the columns to bind (must already exist in both data and final_data)
cols_to_bind <- c(
  "menopause",
  "br",
  "ageatfirstfulltermpreg_cat",
  "interval_menarche_firstbirth_cat",
  "fulltermpreg_cat",
  "oc",
  "mean_brfeed_duration_cat",
  "totalabortion_cat",
  "totalmiscarriage_cat",
  "catht",
  "bmicat",
  "ratio", "family_sitory_of_beo"
)

# Bind the rows: take only those columns from final_data and append to data
data <- rbind(
  data[, cols_to_bind, drop = FALSE],
  final_data[, cols_to_bind, drop = FALSE]
)
data <- unique(data)
```


```{r}
# Function to compute NA percentage for a given group
compute_na_pct <- function(data, variable, br_value, menopause_status) {
  subset_data <- data[data$br == br_value & data$menopause == menopause_status, ]
  total_n <- nrow(subset_data)
  na_n <- sum(is.na(subset_data[[variable]]))
  na_pct <- round(100 * na_n / total_n, 2)
  return(na_pct)
}

# Variables to check
vars <- c("ageatfirstfulltermpreg_cat", "interval_menarche_firstbirth_cat",
          "fulltermpreg_cat", "oc", "mean_brfeed_duration_cat",
          "totalabortion_cat", "totalmiscarriage_cat", "catht", "bmicat", "ratio", "family_sitory_of_beo")

# Initialize an empty data frame to store results
summary_df <- data.frame(variable = vars)

# Add columns for each group combination
summary_df$Control_Premenopausal <- sapply(vars, function(v) compute_na_pct(data, v, "Control", "premenopausal"))
summary_df$Control_Postmenopausal <- sapply(vars, function(v) compute_na_pct(data, v, "Control", "postmenopausal"))
summary_df$Case_Premenopausal <- sapply(vars, function(v) compute_na_pct(data, v, "Breast Cancer Case", "premenopausal"))
summary_df$Case_Postmenopausal <- sapply(vars, function(v) compute_na_pct(data, v, "Breast Cancer Case", "postmenopausal"))

# View the result
print(summary_df)

```

## Subset the data to pre-menopausal women
```{r}
data_premeno = data[which(as.character(data$menopause) == "premenopausal"),]
```

## Number of cases and controls
```{r}
table(data_premeno$br)
```

## Number and proportion of risk factors among controls and cases

## Age at first full term pregnancy
```{r}
table(data_premeno$ageatfirstfulltermpreg_cat
      [as.character(data_premeno$br) == "Control"])
sum(table(data_premeno$ageatfirstfulltermpreg_cat
          [as.character(data_premeno$br) == "Control"]))
prop.table(table(data_premeno$ageatfirstfulltermpreg_cat
                 [as.character(data_premeno$br) == "Control"]))

table(data_premeno$ageatfirstfulltermpreg_cat
      [as.character(data_premeno$br) == "Breast Cancer Case"])
sum(table(data_premeno$ageatfirstfulltermpreg_cat
          [as.character(data_premeno$br) == "Breast Cancer Case"]))
prop.table(table(data_premeno$ageatfirstfulltermpreg_cat
                 [as.character(data_premeno$br) == "Breast Cancer Case"]))

library(dplyr)
library(ggplot2)

data_premeno %>%
  filter(br %in% c("Breast Cancer Case", "Control")) %>%
  mutate(br = factor(br, levels = c("Control", "Breast Cancer Case"))) %>%  # Reversed
  count(br, ageatfirstfulltermpreg_cat) %>%
  group_by(br) %>%
  mutate(prop = n / sum(n)) %>%
  ggplot(aes(x = ageatfirstfulltermpreg_cat, y = prop, fill = br)) +
  geom_bar(stat = "identity", position = position_dodge(0.7), width = 0.6) +
  geom_text(aes(label = scales::percent(prop, accuracy = 0.1)),
            position = position_dodge(0.7), vjust = -0.4, size = 3.5) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(
    title = "Age at First Full-Term Pregnancy (premeno - Old)",
    x = "Age Category", y = "Proportion", fill = "Group"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    legend.position = "top"
  )


```

## Interval between menarche and first full term pregnancy
```{r}
table(data_premeno$interval_menarche_firstbirth_cat
      [as.character(data_premeno$br) == "Control"])
sum(table(data_premeno$interval_menarche_firstbirth_cat
          [as.character(data_premeno$br) == "Control"]))
prop.table(table(data_premeno$interval_menarche_firstbirth_cat
                 [as.character(data_premeno$br) == "Control"]))

table(data_premeno$interval_menarche_firstbirth_cat
      [as.character(data_premeno$br) == "Breast Cancer Case"])
sum(table(data_premeno$interval_menarche_firstbirth_cat
          [as.character(data_premeno$br) == "Breast Cancer Case"]))
prop.table(table(data_premeno$interval_menarche_firstbirth_cat
                 [as.character(data_premeno$br) == "Breast Cancer Case"]))

# Prepare the data
plot_data_premeno_br_interval <- data_premeno %>%
  filter(br %in% c("Control", "Breast Cancer Case")) %>%
  mutate(br = as.character(br)) %>%
  count(br, interval_menarche_firstbirth_cat) %>%
  group_by(br) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()

# Plot
ggplot(plot_data_premeno_br_interval, aes(x = interval_menarche_firstbirth_cat, y = prop, fill = br)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  geom_text(aes(label = scales::percent(prop, accuracy = 0.1)),
            position = position_dodge(width = 0.8),
            vjust = -0.3, size = 3.5) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = c("Breast Cancer Case" = "#1f77b4", "Control" = "#ff7f0e")) +  # Adjust colors as needed
  labs(
    title = "Interval between Menarche and First Birth (Pre Meno, Old)",
    x = "Interval Category", y = "Proportion", fill = "Group"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    legend.position = "top"
  )

```

## Number of full term pregnancies
```{r}
table(data_premeno$fulltermpreg_cat
      [as.character(data_premeno$br) == "Control"])
sum(table(data_premeno$fulltermpreg_cat
          [as.character(data_premeno$br) == "Control"]))
prop.table(table(data_premeno$fulltermpreg_cat
                 [as.character(data_premeno$br) == "Control"]))

table(data_premeno$fulltermpreg_cat
      [as.character(data_premeno$br) == "Breast Cancer Case"])
sum(table(data_premeno$fulltermpreg_cat
          [as.character(data_premeno$br) == "Breast Cancer Case"]))
prop.table(table(data_premeno$fulltermpreg_cat
                 [as.character(data_premeno$br) == "Breast Cancer Case"]))

# Prepare the data
plot_data_premeno_fullpreg <- data_premeno %>%
  filter(br %in% c("Control", "Breast Cancer Case")) %>%
  mutate(br = as.character(br)) %>%
  count(br, fulltermpreg_cat) %>%
  group_by(br) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()

# Plot
ggplot(plot_data_premeno_fullpreg, aes(x = fulltermpreg_cat, y = prop, fill = br)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  geom_text(aes(label = scales::percent(prop, accuracy = 0.1)),
            position = position_dodge(width = 0.8),
            vjust = -0.3, size = 3.5) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = c("Breast Cancer Case" = "#1f77b4", "Control" = "#ff7f0e")) +  # Adjust colors as needed
  labs(
    title = "Full-Term Pregnancy Category (Pre Meno, Old)",
    x = "Full-Term Pregnancy Category", y = "Proportion", fill = "Group"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    legend.position = "top"
  )

```

## OC Use
```{r}
table(data_premeno$oc
      [as.character(data_premeno$br) == "Control"])
sum(table(data_premeno$oc
          [as.character(data_premeno$br) == "Control"]))
prop.table(table(data_premeno$oc
                 [as.character(data_premeno$br) == "Control"]))

table(data_premeno$oc
      [as.character(data_premeno$br) == "Breast Cancer Case"])
sum(table(data_premeno$oc
          [as.character(data_premeno$br) == "Breast Cancer Case"]))
prop.table(table(data_premeno$oc
                 [as.character(data_premeno$br) == "Breast Cancer Case"]))

# Prepare the data
plot_data_premeno_oc <- data_premeno %>%
  filter(br %in% c("Control", "Breast Cancer Case")) %>%
  mutate(br = as.character(br)) %>%
  count(br, oc) %>%
  group_by(br) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()

# Plot
ggplot(plot_data_premeno_oc, aes(x = oc, y = prop, fill = br)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  geom_text(aes(label = scales::percent(prop, accuracy = 0.1)),
            position = position_dodge(width = 0.8),
            vjust = -0.3, size = 3.5) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = c("Breast Cancer Case" = "#1f77b4", "Control" = "#ff7f0e")) +  # Adjust colors as needed
  labs(
    title = "Oral Contraceptive Use (Pre Meno, Old)",
    x = "Oral Contraceptive Use", y = "Proportion", fill = "Group"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    legend.position = "top"
  )

```

## Duration of breast-feeding
```{r}
table(data_premeno$mean_brfeed_duration_cat
      [as.character(data_premeno$br) == "Control"])
sum(table(data_premeno$mean_brfeed_duration_cat
          [as.character(data_premeno$br) == "Control"]))
prop.table(table(data_premeno$mean_brfeed_duration_cat
                 [as.character(data_premeno$br) == "Control"]))

table(data_premeno$mean_brfeed_duration_cat
      [as.character(data_premeno$br) == "Breast Cancer Case"])
sum(table(data_premeno$mean_brfeed_duration_cat
          [as.character(data_premeno$br) == "Breast Cancer Case"]))
prop.table(table(data_premeno$mean_brfeed_duration_cat
                 [as.character(data_premeno$br) == "Breast Cancer Case"]))

# Prepare the data
plot_data_premeno_brfeed <- data_premeno %>%
  filter(br %in% c("Control", "Breast Cancer Case")) %>%
  mutate(br = as.character(br)) %>%
  count(br, mean_brfeed_duration_cat) %>%
  group_by(br) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()

# Plot
ggplot(plot_data_premeno_brfeed, aes(x = mean_brfeed_duration_cat, y = prop, fill = br)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  geom_text(aes(label = scales::percent(prop, accuracy = 0.1)),
            position = position_dodge(width = 0.8),
            vjust = -0.3, size = 3.5) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = c("Breast Cancer Case" = "#1f77b4", "Control" = "#ff7f0e")) +  # Adjust colors as needed
  labs(
    title = "Mean Breastfeeding Duration Category (Pre Meno, Old)",
    x = "Breastfeeding Duration Category", y = "Proportion", fill = "Group"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    legend.position = "top"
  )

```

## Number of induced abortions
```{r}
table(data_premeno$totalabortion_cat
      [as.character(data_premeno$br) == "Control"])
sum(table(data_premeno$totalabortion_cat
          [as.character(data_premeno$br) == "Control"]))
prop.table(table(data_premeno$totalabortion_cat
                 [as.character(data_premeno$br) == "Control"]))

table(data_premeno$totalabortion_cat
      [as.character(data_premeno$br) == "Breast Cancer Case"])
sum(table(data_premeno$totalabortion_cat
          [as.character(data_premeno$br) == "Breast Cancer Case"]))
prop.table(table(data_premeno$totalabortion_cat
                 [as.character(data_premeno$br) == "Breast Cancer Case"]))
```

## Number of spontaneous abortions
```{r}
table(data_premeno$totalmiscarriage_cat
      [as.character(data_premeno$br) == "Control"])
sum(table(data_premeno$totalmiscarriage_cat
          [as.character(data_premeno$br) == "Control"]))
prop.table(table(data_premeno$totalmiscarriage_cat
                 [as.character(data_premeno$br) == "Control"]))

table(data_premeno$totalmiscarriage_cat
      [as.character(data_premeno$br) == "Breast Cancer Case"])
sum(table(data_premeno$totalmiscarriage_cat
          [as.character(data_premeno$br) == "Breast Cancer Case"]))
prop.table(table(data_premeno$totalmiscarriage_cat
                 [as.character(data_premeno$br) == "Breast Cancer Case"]))
```
Q. (Doesn't make sense!)
## Family history of breast, endometrial and ovarian cancer 
```{r}
table(data_premeno$family_sitory_of_beo
      [as.character(data_premeno$br) == "Control"])
sum(table(data_premeno$family_sitory_of_beo
          [as.character(data_premeno$br) == "Control"]))
prop.table(table(data_premeno$family_sitory_of_beo
                 [as.character(data_premeno$br) == "Control"]))

table(data_premeno$family_sitory_of_beo
      [as.character(data_premeno$br) == "Breast Cancer Case"])
sum(table(data_premeno$family_sitory_of_beo
          [as.character(data_premeno$br) == "Breast Cancer Case"]))
prop.table(table(data_premeno$family_sitory_of_beo
                 [as.character(data_premeno$br) == "Breast Cancer Case"]))

library(ggplot2)
library(dplyr)



# Prepare data for family_sitory_of_beo (Premeno)
plot_data_family_history <- data_premeno %>%
  filter(br %in% c("Control", "Breast Cancer Case")) %>%
  mutate(br = as.character(br)) %>%
  count(br, family_sitory_of_beo) %>%
  group_by(br) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()

# Plot
ggplot(plot_data_family_history, aes(x = family_sitory_of_beo, y = prop, fill = br)) +
  geom_bar(stat = "identity", position = position_dodge(0.8), width = 0.7) +
  geom_text(aes(label = scales::percent(prop, accuracy = 0.1)),
            position = position_dodge(0.8), vjust = -0.3, size = 3.5) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = c("Breast Cancer Case" = "#1f77b4", "Control" = "#ff7f0e")) +
  labs(
    title = "Family History of Breast/Ovarian Cancer (Premeno - old)",
    x = "Family History", y = "Proportion", fill = "Group"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    legend.position = "top"
  )

```

## BMI
```{r}
table(data_premeno$bmicat
      [as.character(data_premeno$br) == "Control"])
sum(table(data_premeno$bmicat
          [as.character(data_premeno$br) == "Control"]))
prop.table(table(data_premeno$bmicat
                 [as.character(data_premeno$br) == "Control"]))

table(data_premeno$bmicat
      [as.character(data_premeno$br) == "Breast Cancer Case"])
sum(table(data_premeno$bmicat
          [as.character(data_premeno$br) == "Breast Cancer Case"]))
prop.table(table(data_premeno$bmicat
                 [as.character(data_premeno$br) == "Breast Cancer Case"]))

# Prepare data for bmicat (Premeno)
plot_data_bmicat <- data_premeno %>%
  filter(br %in% c("Control", "Breast Cancer Case")) %>%
  mutate(br = as.character(br)) %>%
  count(br, bmicat) %>%
  group_by(br) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()

# Plot
ggplot(plot_data_bmicat, aes(x = bmicat, y = prop, fill = br)) +
  geom_bar(stat = "identity", position = position_dodge(0.8), width = 0.7) +
  geom_text(aes(label = scales::percent(prop, accuracy = 0.1)),
            position = position_dodge(0.8), vjust = -0.3, size = 3.5) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = c("Breast Cancer Case" = "#1f77b4", "Control" = "#ff7f0e")) +
  labs(
    title = "BMI Category (Premeno - old)",
    x = "BMI Category", y = "Proportion", fill = "Group"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    legend.position = "top"
  )

```

## Height
```{r}
table(data_premeno$catht
      [as.character(data_premeno$br) == "Control"])
sum(table(data_premeno$catht
          [as.character(data_premeno$br) == "Control"]))
prop.table(table(data_premeno$catht
                 [as.character(data_premeno$br) == "Control"]))

table(data_premeno$catht
      [as.character(data_premeno$br) == "Breast Cancer Case"])
sum(table(data_premeno$catht
          [as.character(data_premeno$br) == "Breast Cancer Case"]))
prop.table(table(data_premeno$catht
                 [as.character(data_premeno$br) == "Breast Cancer Case"]))

# Prepare data for catht (Premeno)
plot_data_catht <- data_premeno %>%
  filter(br %in% c("Control", "Breast Cancer Case")) %>%
  mutate(br = as.character(br)) %>%
  count(br, catht) %>%
  group_by(br) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()

# Plot
ggplot(plot_data_catht, aes(x = catht, y = prop, fill = br)) +
  geom_bar(stat = "identity", position = position_dodge(0.8), width = 0.7) +
  geom_text(aes(label = scales::percent(prop, accuracy = 0.1)),
            position = position_dodge(0.8), vjust = -0.3, size = 3.5) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = c("Breast Cancer Case" = "#1f77b4", "Control" = "#ff7f0e")) +
  labs(
    title = "Height Category (Premeno - old)",
    x = "Height Category", y = "Proportion", fill = "Group"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    legend.position = "top"
  )

```

## Waist-to-hip ratio
```{r}
table(data_premeno$ratio
      [as.character(data_premeno$br) == "Control"])
sum(table(data_premeno$ratio
          [as.character(data_premeno$br) == "Control"]))
prop.table(table(data_premeno$ratio
                 [as.character(data_premeno$br) == "Control"]))

table(data_premeno$ratio
      [as.character(data_premeno$br) == "Breast Cancer Case"])
sum(table(data_premeno$ratio
          [as.character(data_premeno$br) == "Breast Cancer Case"]))
prop.table(table(data_premeno$ratio
                 [as.character(data_premeno$br) == "Breast Cancer Case"]))

# Prepare data for ratio (Premeno)
plot_data_ratio <- data_premeno %>%
  filter(br %in% c("Control", "Breast Cancer Case")) %>%
  mutate(br = as.character(br)) %>%
  count(br, ratio) %>%
  group_by(br) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()

# Plot
ggplot(plot_data_ratio, aes(x = ratio, y = prop, fill = br)) +
  geom_bar(stat = "identity", position = position_dodge(0.8), width = 0.7) +
  geom_text(aes(label = scales::percent(prop, accuracy = 0.1)),
            position = position_dodge(0.8), vjust = -0.3, size = 3.5) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = c("Breast Cancer Case" = "#1f77b4", "Control" = "#ff7f0e")) +
  labs(
    title = "WHR Category (Premeno - old)",  # Change title as needed
    x = "Waist-Hip Ratio Category", y = "Proportion", fill = "Group"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    legend.position = "top"
  )

```

## BMI and WHR interaction
```{r}
table(data_premeno$bmicat
      [as.character(data_premeno$br) == "Control"], 
      data_premeno$ratio
      [as.character(data_premeno$br) == "Control"])
sum(table(data_premeno$bmicat
          [as.character(data_premeno$br) == "Control"], 
          data_premeno$ratio
          [as.character(data_premeno$br) == "Control"]))
prop.table(table(data_premeno$bmicat
                 [as.character(data_premeno$br) == "Control"], 
                 data_premeno$ratio
                 [as.character(data_premeno$br) == "Control"]))


table(data_premeno$bmicat
      [as.character(data_premeno$br) == "Breast Cancer Case"], 
      data_premeno$ratio
      [as.character(data_premeno$br) == "Breast Cancer Case"])
sum(table(data_premeno$bmicat
          [as.character(data_premeno$br) == "Breast Cancer Case"], 
          data_premeno$ratio
          [as.character(data_premeno$br) == "Breast Cancer Case"]))
prop.table(table(data_premeno$bmicat
                 [as.character(data_premeno$br) == "Breast Cancer Case"], 
                 data_premeno$ratio
                 [as.character(data_premeno$br) == "Breast Cancer Case"]))
```

## Categorical age at enrollment
```{r}
# table(data_premeno$agecode
#       [as.character(data_premeno$br) == "Control"])
# sum(table(data_premeno$agecode
#           [as.character(data_premeno$br) == "Control"]))
# prop.table(table(data_premeno$agecode
#                  [as.character(data_premeno$br) == "Control"]))
# 
# table(data_premeno$agecode
#       [as.character(data_premeno$br) == "Breast Cancer Case"])
# sum(table(data_premeno$agecode
#           [as.character(data_premeno$br) == "Breast Cancer Case"]))
# prop.table(table(data_premeno$agecode
#                  [as.character(data_premeno$br) == "Breast Cancer Case"]))
```

## Geographical location
```{r}
# table(data_premeno$area
#       [as.character(data_premeno$br) == "Control"])
# sum(table(data_premeno$area
#           [as.character(data_premeno$br) == "Control"]))
# prop.table(table(data_premeno$area
#                  [as.character(data_premeno$br) == "Control"]))
# 
# table(data_premeno$area
#       [as.character(data_premeno$br) == "Breast Cancer Case"])
# sum(table(data_premeno$area
#           [as.character(data_premeno$br) == "Breast Cancer Case"]))
# prop.table(table(data_premeno$area
#                  [as.character(data_premeno$br) == "Breast Cancer Case"]))
# 
# # Prepare data for area (Premeno)
# plot_data_area <- data_premeno %>%
#   filter(br %in% c("Control", "Breast Cancer Case")) %>%
#   mutate(br = as.character(br)) %>%
#   count(br, area) %>%
#   group_by(br) %>%
#   mutate(prop = n / sum(n)) %>%
#   ungroup()
# 
# # Plot
# ggplot(plot_data_area, aes(x = area, y = prop, fill = br)) +
#   geom_bar(stat = "identity", position = position_dodge(0.8), width = 0.7) +
#   geom_text(aes(label = scales::percent(prop, accuracy = 0.1)),
#             position = position_dodge(0.8), vjust = -0.3, size = 3.5) +
#   scale_y_continuous(labels = scales::percent_format()) +
#   scale_fill_manual(values = c("Breast Cancer Case" = "#1f77b4", "Control" = "#ff7f0e")) +
#   labs(
#     title = "Residence Area (Premeno - old)",
#     x = "Area", y = "Proportion", fill = "Group"
#   ) +
#   theme_minimal(base_size = 14) +
#   theme(
#     plot.title = element_text(face = "bold", hjust = 0.5),
#     legend.position = "top"
#   )

```
(Important factor to avoid biased sample)
## Rural/urban based on first 20 years
```{r}
# table(data_premeno$residancefirsttwentyyears
#       [as.character(data_premeno$br) == "Control"])
# sum(table(data_premeno$residancefirsttwentyyears
#           [as.character(data_premeno$br) == "Control"]))
# prop.table(table(data_premeno$residancefirsttwentyyears
#                  [as.character(data_premeno$br) == "Control"]))
# 
# table(data_premeno$residancefirsttwentyyears
#       [as.character(data_premeno$br) == "Breast Cancer Case"])
# sum(table(data_premeno$residancefirsttwentyyears
#           [as.character(data_premeno$br) == "Breast Cancer Case"]))
# prop.table(table(data_premeno$residancefirsttwentyyears
#                  [as.character(data_premeno$br) == "Breast Cancer Case"]))
# 
# # Prepare data for residancefirsttwentyyears (Premeno)
# plot_data_residence <- data_premeno %>%
#   filter(br %in% c("Control", "Breast Cancer Case")) %>%
#   mutate(br = as.character(br)) %>%
#   count(br, residancefirsttwentyyears) %>%
#   group_by(br) %>%
#   mutate(prop = n / sum(n)) %>%
#   ungroup()
# 
# # Plot
# ggplot(plot_data_residence, aes(x = residancefirsttwentyyears, y = prop, fill = br)) +
#   geom_bar(stat = "identity", position = position_dodge(0.8), width = 0.7) +
#   geom_text(aes(label = scales::percent(prop, accuracy = 0.1)),
#             position = position_dodge(0.8), vjust = -0.3, size = 3.5) +
#   scale_y_continuous(labels = scales::percent_format()) +
#   scale_fill_manual(values = c("Breast Cancer Case" = "#1f77b4", "Control" = "#ff7f0e")) +
#   labs(
#     title = "Residence in First 20 Years (Premeno - old)",
#     x = "Residence Type", y = "Proportion", fill = "Group"
#   ) +
#   theme_minimal(base_size = 14) +
#   theme(
#     plot.title = element_text(face = "bold", hjust = 0.5),
#     legend.position = "top"
#   )

```

## Binary education status: <8 years and >= 8 years of schooling
```{r}
# table(data_premeno$ed
#       [as.character(data_premeno$br) == "Control"])
# sum(table(data_premeno$ed
#           [as.character(data_premeno$br) == "Control"]))
# prop.table(table(data_premeno$ed
#                  [as.character(data_premeno$br) == "Control"]))
# 
# table(data_premeno$ed
#       [as.character(data_premeno$br) == "Breast Cancer Case"])
# sum(table(data_premeno$ed
#           [as.character(data_premeno$br) == "Breast Cancer Case"]))
# prop.table(table(data_premeno$ed
#                  [as.character(data_premeno$br) == "Breast Cancer Case"]))
# 
# # Prepare data for education (ed) among Premeno
# plot_data_education <- data_premeno %>%
#   filter(br %in% c("Control", "Breast Cancer Case")) %>%
#   mutate(br = as.character(br)) %>%
#   count(br, ed) %>%
#   group_by(br) %>%
#   mutate(prop = n / sum(n)) %>%
#   ungroup()
# 
# # Plot
# ggplot(plot_data_education, aes(x = ed, y = prop, fill = br)) +
#   geom_bar(stat = "identity", position = position_dodge(0.8), width = 0.7) +
#   geom_text(aes(label = scales::percent(prop, accuracy = 0.1)),
#             position = position_dodge(0.8), vjust = -0.3, size = 3.5) +
#   scale_y_continuous(labels = scales::percent_format()) +
#   scale_fill_manual(values = c("Breast Cancer Case" = "#1f77b4", "Control" = "#ff7f0e")) +
#   labs(
#     title = "Education Level (Premeno - old)",
#     x = "Education Category", y = "Proportion", fill = "Group"
#   ) +
#   theme_minimal(base_size = 14) +
#   theme(
#     plot.title = element_text(face = "bold", hjust = 0.5),
#     legend.position = "top"
#   )

```

## Relative risk model for premenopausal women
```{r}
model_premeno = glm(br ~ as.factor(ageatfirstfulltermpreg_cat)
                    + interval_menarche_firstbirth_cat   
                    + as.factor(fulltermpreg_cat)
                    + oc + as.factor(mean_brfeed_duration_cat)
                    + as.factor(totalabortion_cat)
                    + as.factor(totalmiscarriage_cat)
                    + as.factor(catht) 
                    + as.factor(bmicat) * as.factor(ratio),
                    data = data_premeno, family = binomial(link=logit), 
                    na.action = na.omit)

# as.factor(agecode) + as.factor(area) 
#                     + as.factor(ed) + as.factor(residancefirsttwentyyears)
#                     + 

cbind(exp(coef(model_premeno)),exp(confint(model_premeno)))

model.frame(model_premeno) |> nrow()
nrow(data_premeno)

summary(model_premeno)
```

##Joint association of BMI and WHR#####
```{r}
#Log relative risks and variances
estimate = coef(model_premeno)[c("as.factor(bmicat)2","as.factor(bmicat)3",
                                 "as.factor(ratio)1","as.factor(ratio)2",
                                 "as.factor(bmicat)2:as.factor(ratio)1",
                                 "as.factor(bmicat)3:as.factor(ratio)1",
                                 "as.factor(bmicat)2:as.factor(ratio)2",
                                 "as.factor(bmicat)3:as.factor(ratio)2")]

variance = vcov(model_premeno)[c("as.factor(bmicat)2","as.factor(bmicat)3",
                                 "as.factor(ratio)1","as.factor(ratio)2",
                                 "as.factor(bmicat)2:as.factor(ratio)1",
                                 "as.factor(bmicat)3:as.factor(ratio)1",
                                 "as.factor(bmicat)2:as.factor(ratio)2",
                                 "as.factor(bmicat)3:as.factor(ratio)2"),
                               c("as.factor(bmicat)2","as.factor(bmicat)3",
                                 "as.factor(ratio)1","as.factor(ratio)2",
                                 "as.factor(bmicat)2:as.factor(ratio)1",
                                 "as.factor(bmicat)3:as.factor(ratio)1",
                                 "as.factor(bmicat)2:as.factor(ratio)2",
                                 "as.factor(bmicat)3:as.factor(ratio)2")]

#BMI < 24.9, WHR < 0.84

RR1 = 1

ci1 = c(1,1)

#BMI < 24.9, WHR 0.85-0.94

RR2 = exp(coef(model_premeno))["as.factor(ratio)1"]

ci2 = exp(confint(model_premeno))["as.factor(ratio)1",]
#BMI < 24.9, WHR >= 0.95

RR3 = exp(coef(model_premeno))["as.factor(ratio)2"]

ci3 = exp(confint(model_premeno))["as.factor(ratio)2",]

#BMI 25-29.9, WHR < 0.84

RR4 = exp(coef(model_premeno))["as.factor(bmicat)2"]

ci4 = exp(confint(model_premeno))["as.factor(bmicat)2",]

#BMI 25-29.9, WHR 0.85-0.94

RR5 = exp(coef(model_premeno))["as.factor(bmicat)2"] * 
  exp(coef(model_premeno))["as.factor(ratio)1"] * 
  exp(coef(model_premeno))["as.factor(bmicat)2:as.factor(ratio)1"]

v5 = c(1,0,1,0,1,0,0,0)

RR5 = exp(v5 %*% estimate)

sd5 = sqrt(v5 %*% variance %*% v5)

ci5 = exp(c(log(RR5) - 1.96 * sd5, log(RR5) + 1.96 * sd5))

#BMI 25-29.9, WHR >= 0.95

RR6 = exp(coef(model_premeno))["as.factor(bmicat)2"] * 
  exp(coef(model_premeno))["as.factor(ratio)2"] * 
  exp(coef(model_premeno))["as.factor(bmicat)2:as.factor(ratio)2"]

v6 = c(1,0,0,1,0,0,1,0)

RR6 = exp(v6 %*% estimate)

sd6 = sqrt(v6 %*% variance %*% v6)

ci6 = exp(c(log(RR6) - 1.96 * sd6, log(RR6) + 1.96 * sd6))

#BMI >= 30, WHR < 0.84

RR7 = exp(coef(model_premeno))["as.factor(bmicat)3"]

ci7 = exp(confint(model_premeno))["as.factor(bmicat)3",]

#BMI >= 30, WHR 0.85-0.94

RR8 = exp(coef(model_premeno))["as.factor(bmicat)3"] * 
  exp(coef(model_premeno))["as.factor(ratio)1"] * 
  exp(coef(model_premeno))["as.factor(bmicat)3:as.factor(ratio)1"]

v8 = c(0,1,1,0,0,1,0,0)

RR8 = exp(v8 %*% estimate)

sd8 = sqrt(v8 %*% variance %*% v8)

ci8 = exp(c(log(RR8) - 1.96 * sd8, log(RR8) + 1.96 * sd8))

#BMI >= 30, WHR >= 0.95

RR9 = exp(coef(model_premeno))["as.factor(bmicat)3"] * 
  exp(coef(model_premeno))["as.factor(ratio)2"] * 
  exp(coef(model_premeno))["as.factor(bmicat)3:as.factor(ratio)2"]

v9 = c(0,1,0,1,0,0,0,1)

RR9 = exp(v9 %*% estimate)

sd9 = sqrt(v9 %*% variance %*% v9)

ci9 = exp(c(log(RR9) - 1.96 * sd9, log(RR9) + 1.96 * sd9))


BMI_WHR_Premeno = cbind(c(RR1,RR2,RR3,RR4,RR5,RR6,RR7,RR8,RR9),
                        rbind(ci1,ci2,ci3,ci4,ci5,ci6,ci7,ci8,ci9))

rownames(BMI_WHR_Premeno)[c(1,5,6,8,9)] = c("Ref","as.factor(bmicat)2:as.factor(ratio)1",
                                         "as.factor(bmicat)2:as.factor(ratio)2",
                                         "as.factor(bmicat)3:as.factor(ratio)1",
                                         "as.factor(bmicat)2:as.factor(ratio)2")

#pdf(file = "BMI_WHR_Interaction_Premeno")
plotCI(x = BMI_WHR_Premeno[,1], ui = BMI_WHR_Premeno[,3], li = BMI_WHR_Premeno[,2],xlab = "",
       ylab = "Joint Relative Risk",pch=16,lwd = 2, xaxt = "n", 
       col = c(rep("red",3),rep("blue",3),rep("green",3)))

mtext(expression("BMI"<="24.9"), side = 3, line = 1, at = 2, col = "red")
mtext(expression("BMI 25-29.9"), side = 3, line = 1, at = 5, col = "blue")
mtext(expression("BMI">="30"), side = 3, line = 1, at = 8, col = "green")

mtext(rep(expression("WHR"<=0.84),3), side = 1, line = 0.2, las = 2, 
      at = c(1,4,7), col = c("red","blue","green"), cex = 0.7)
mtext(rep("WHR 0.85-0.94",3), side = 1, line = 0.2, las = 2, 
      at = c(2,5,8), col = c("red","blue","green"), cex = 0.7)
mtext(rep(expression("WHR">=0.95),3), side = 1, line = 0.2, las = 2, 
      at = c(3,6,9), col = c("red","blue","green"), cex = 0.7)
#dev.off()

```

## Subset the data to post-menopausal women
```{r}
data_postmeno = data[which(as.character(data$menopause) == "postmenopausal"),]
```
## Number of cases and controls
```{r}
table(data_postmeno$br)
```

## Number and proportion of risk factors among controls

## Age at first full term pregnancy
```{r}
table(data_postmeno$ageatfirstfulltermpreg_cat
      [as.character(data_postmeno$br) == "Control"])
sum(table(data_postmeno$ageatfirstfulltermpreg_cat
          [as.character(data_postmeno$br) == "Control"]))
prop.table(table(data_postmeno$ageatfirstfulltermpreg_cat
                 [as.character(data_postmeno$br) == "Control"]))

table(data_postmeno$ageatfirstfulltermpreg_cat
      [as.character(data_postmeno$br) == "Breast Cancer Case"])
sum(table(data_postmeno$ageatfirstfulltermpreg_cat
          [as.character(data_postmeno$br) == "Breast Cancer Case"]))
prop.table(table(data_postmeno$ageatfirstfulltermpreg_cat
                 [as.character(data_postmeno$br) == "Breast Cancer Case"]))

library(dplyr)
library(ggplot2)

# Prepare the data
plot_data_post_br <- data_postmeno %>%
  filter(br %in% c("Control", "Breast Cancer Case")) %>%
  mutate(br = as.character(br)) %>%
  count(br, ageatfirstfulltermpreg_cat) %>%
  group_by(br) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()

# Plot
ggplot(plot_data_post_br, aes(x = ageatfirstfulltermpreg_cat, y = prop, fill = br)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  geom_text(aes(label = scales::percent(prop, accuracy = 0.1)),
            position = position_dodge(width = 0.8),
            vjust = -0.3, size = 3.5) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = c("Breast Cancer Case" = "#1f77b4", "Control" = "#ff7f0e")) +  # Swap colors if needed
  labs(
    title = "Age at First Full-Term Pregnancy(Postmeno - Old)",
    x = "Age Category", y = "Proportion", fill = "Group"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    legend.position = "top"
  )

```

## Interval between menarche and first full term pregnancy
```{r}
table(data_postmeno$interval_menarche_firstbirth_cat
      [as.character(data_postmeno$br) == "Control"])
sum(table(data_postmeno$interval_menarche_firstbirth_cat
          [as.character(data_postmeno$br) == "Control"]))
prop.table(table(data_postmeno$interval_menarche_firstbirth_cat
                 [as.character(data_postmeno$br) == "Control"]))

table(data_postmeno$interval_menarche_firstbirth_cat
      [as.character(data_postmeno$br) == "Breast Cancer Case"])
sum(table(data_postmeno$interval_menarche_firstbirth_cat
          [as.character(data_postmeno$br) == "Breast Cancer Case"]))
prop.table(table(data_postmeno$interval_menarche_firstbirth_cat
                 [as.character(data_postmeno$br) == "Breast Cancer Case"]))

# Prepare the data
plot_data_post_br_interval <- data_postmeno %>%
  filter(br %in% c("Control", "Breast Cancer Case")) %>%
  mutate(br = as.character(br)) %>%
  count(br, interval_menarche_firstbirth_cat) %>%
  group_by(br) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()

# Plot
ggplot(plot_data_post_br_interval, aes(x = interval_menarche_firstbirth_cat, y = prop, fill = br)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  geom_text(aes(label = scales::percent(prop, accuracy = 0.1)),
            position = position_dodge(width = 0.8),
            vjust = -0.3, size = 3.5) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = c("Breast Cancer Case" = "#1f77b4", "Control" = "#ff7f0e")) +  # Adjust colors as needed
  labs(
    title = "Interval between Menarche and First Birth (PostMeno, Old)",
    x = "Interval Category", y = "Proportion", fill = "Group"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    legend.position = "top"
  )

```

## Number of full term pregnancies
```{r}
table(data_postmeno$fulltermpreg_cat
      [as.character(data_postmeno$br) == "Control"])
sum(table(data_postmeno$fulltermpreg_cat
          [as.character(data_postmeno$br) == "Control"]))
prop.table(table(data_postmeno$fulltermpreg_cat
                 [as.character(data_postmeno$br) == "Control"]))

table(data_postmeno$fulltermpreg_cat
      [as.character(data_postmeno$br) == "Breast Cancer Case"])
sum(table(data_postmeno$fulltermpreg_cat
          [as.character(data_postmeno$br) == "Breast Cancer Case"]))
prop.table(table(data_postmeno$fulltermpreg_cat
                 [as.character(data_postmeno$br) == "Breast Cancer Case"]))

# Prepare the data
plot_data_postmeno_fullterm <- data_postmeno %>%
  filter(br %in% c("Control", "Breast Cancer Case")) %>%
  mutate(br = as.character(br)) %>%
  count(br, fulltermpreg_cat) %>%
  group_by(br) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()

# Plot
ggplot(plot_data_postmeno_fullterm, aes(x = fulltermpreg_cat, y = prop, fill = br)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  geom_text(aes(label = scales::percent(prop, accuracy = 0.1)),
            position = position_dodge(width = 0.8),
            vjust = -0.3, size = 3.5) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = c("Breast Cancer Case" = "#1f77b4", "Control" = "#ff7f0e")) +
  labs(
    title = "Number of Full-Term Pregnancies (Postmeno, Old)",
    x = "Full-Term Pregnancy Category", y = "Proportion", fill = "Group"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    legend.position = "top"
  )

```

## OC Use
```{r}
table(data_postmeno$oc
      [as.character(data_postmeno$br) == "Control"])
sum(table(data_postmeno$oc
          [as.character(data_postmeno$br) == "Control"]))
prop.table(table(data_postmeno$oc
                 [as.character(data_postmeno$br) == "Control"]))

table(data_postmeno$oc
      [as.character(data_postmeno$br) == "Breast Cancer Case"])
sum(table(data_postmeno$oc
          [as.character(data_postmeno$br) == "Breast Cancer Case"]))
prop.table(table(data_postmeno$oc
                 [as.character(data_postmeno$br) == "Breast Cancer Case"]))

# Prepare data
plot_data_postmeno_oc <- data_postmeno %>%
  filter(br %in% c("Control", "Breast Cancer Case")) %>%
  mutate(br = as.character(br)) %>%
  count(br, oc) %>%
  group_by(br) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()

# Plot
ggplot(plot_data_postmeno_oc, aes(x = oc, y = prop, fill = br)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  geom_text(aes(label = scales::percent(prop, accuracy = 0.1)),
            position = position_dodge(width = 0.8),
            vjust = -0.3, size = 3.5) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = c("Breast Cancer Case" = "#1f77b4", "Control" = "#ff7f0e")) +
  labs(
    title = "Oral Contraceptive Use (Postmeno, Old)",
    x = "OC Use", y = "Proportion", fill = "Group"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    legend.position = "top"
  )

```

##Age at menopause
```{r}
# table(data_postmeno$catmenopause
#       [as.character(data_postmeno$br) == "Control"])
# sum(table(data_postmeno$catmenopause
#           [as.character(data_postmeno$br) == "Control"]))
# prop.table(table(data_postmeno$catmenopause
#                  [as.character(data_postmeno$br) == "Control"]))
# 
# table(data_postmeno$catmenopause
#       [as.character(data_postmeno$br) == "Breast Cancer Case"])
# sum(table(data_postmeno$catmenopause
#           [as.character(data_postmeno$br) == "Breast Cancer Case"]))
# prop.table(table(data_postmeno$catmenopause
#                  [as.character(data_postmeno$br) == "Breast Cancer Case"]))
# 
# # Prepare data
# plot_data_catmeno <- data_postmeno %>%
#   filter(br %in% c("Control", "Breast Cancer Case")) %>%
#   mutate(br = as.character(br)) %>%
#   count(br, catmenopause) %>%
#   group_by(br) %>%
#   mutate(prop = n / sum(n)) %>%
#   ungroup()
# 
# # Plot
# ggplot(plot_data_catmeno, aes(x = catmenopause, y = prop, fill = br)) +
#   geom_bar(stat = "identity", position = position_dodge(0.8), width = 0.7) +
#   geom_text(aes(label = scales::percent(prop, accuracy = 0.1)),
#             position = position_dodge(0.8), vjust = -0.3, size = 3.5) +
#   scale_y_continuous(labels = scales::percent_format()) +
#   scale_fill_manual(values = c("Breast Cancer Case" = "#1f77b4", "Control" = "#ff7f0e")) +
#   labs(
#     title = "Type of Menopause (Postmeno, Old)",
#     x = "Menopause Category", y = "Proportion", fill = "Group"
#   ) +
#   theme_minimal(base_size = 14) +
#   theme(
#     plot.title = element_text(face = "bold", hjust = 0.5),
#     legend.position = "top"
#   )

```

## Duration of breast-feeding
```{r}
table(data_postmeno$mean_brfeed_duration_cat
      [as.character(data_postmeno$br) == "Control"])
sum(table(data_postmeno$mean_brfeed_duration_cat
          [as.character(data_postmeno$br) == "Control"]))
prop.table(table(data_postmeno$mean_brfeed_duration_cat
                 [as.character(data_postmeno$br) == "Control"]))

table(data_postmeno$mean_brfeed_duration_cat
      [as.character(data_postmeno$br) == "Breast Cancer Case"])
sum(table(data_postmeno$mean_brfeed_duration_cat
          [as.character(data_postmeno$br) == "Breast Cancer Case"]))
prop.table(table(data_postmeno$mean_brfeed_duration_cat
                 [as.character(data_postmeno$br) == "Breast Cancer Case"]))

# Prepare data
plot_data_bf <- data_postmeno %>%
  filter(br %in% c("Control", "Breast Cancer Case")) %>%
  mutate(br = as.character(br)) %>%
  count(br, mean_brfeed_duration_cat) %>%
  group_by(br) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()

# Plot
ggplot(plot_data_bf, aes(x = mean_brfeed_duration_cat, y = prop, fill = br)) +
  geom_bar(stat = "identity", position = position_dodge(0.8), width = 0.7) +
  geom_text(aes(label = scales::percent(prop, accuracy = 0.1)),
            position = position_dodge(0.8), vjust = -0.3, size = 3.5) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = c("Breast Cancer Case" = "#1f77b4", "Control" = "#ff7f0e")) +
  labs(
    title = "Mean Breastfeeding Duration (Postmeno, Old)",
    x = "Breastfeeding Duration Category", y = "Proportion", fill = "Group"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    legend.position = "top"
  )

```

## Number of induced abortions
```{r}
table(data_postmeno$totalabortion_cat
      [as.character(data_postmeno$br) == "Control"])
sum(table(data_postmeno$totalabortion_cat
          [as.character(data_postmeno$br) == "Control"]))
prop.table(table(data_postmeno$totalabortion_cat
                 [as.character(data_postmeno$br) == "Control"]))

table(data_postmeno$totalabortion_cat
      [as.character(data_postmeno$br) == "Breast Cancer Case"])
sum(table(data_postmeno$totalabortion_cat
          [as.character(data_postmeno$br) == "Breast Cancer Case"]))
prop.table(table(data_postmeno$totalabortion_cat
                 [as.character(data_postmeno$br) == "Breast Cancer Case"]))
```

## Number of spontaneous abortions
```{r}
table(data_postmeno$totalmiscarriage_cat
      [as.character(data_postmeno$br) == "Control"])
sum(table(data_postmeno$totalmiscarriage_cat
          [as.character(data_postmeno$br) == "Control"]))
prop.table(table(data_postmeno$totalmiscarriage_cat
                 [as.character(data_postmeno$br) == "Control"]))

table(data_postmeno$totalmiscarriage_cat
      [as.character(data_postmeno$br) == "Breast Cancer Case"])
sum(table(data_postmeno$totalmiscarriage_cat
          [as.character(data_postmeno$br) == "Breast Cancer Case"]))
prop.table(table(data_postmeno$totalmiscarriage_cat
                 [as.character(data_postmeno$br) == "Breast Cancer Case"]))
```
(why no differential family history)
## Family history of breast, endometrial and ovarian cancer 
```{r}
table(data_postmeno$family_sitory_of_beo
      [as.character(data_postmeno$br) == "Control"])
sum(table(data_postmeno$family_sitory_of_beo
          [as.character(data_postmeno$br) == "Control"]))
prop.table(table(data_postmeno$family_sitory_of_beo
                 [as.character(data_postmeno$br) == "Control"]))

table(data_postmeno$family_sitory_of_beo
      [as.character(data_postmeno$br) == "Breast Cancer Case"])
sum(table(data_postmeno$family_sitory_of_beo
          [as.character(data_postmeno$br) == "Breast Cancer Case"]))
prop.table(table(data_postmeno$family_sitory_of_beo
                 [as.character(data_postmeno$br) == "Breast Cancer Case"]))

# Prepare data for family history (Postmenopausal - Old)
plot_data_family_history_old <- data_postmeno %>%
  filter(br %in% c("Control", "Breast Cancer Case")) %>%
  mutate(br = as.character(br)) %>%
  count(br, family_sitory_of_beo) %>%
  group_by(br) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()

# Plot
ggplot(plot_data_family_history_old, aes(x = family_sitory_of_beo, y = prop, fill = br)) +
  geom_bar(stat = "identity", position = position_dodge(0.8), width = 0.7) +
  geom_text(aes(label = scales::percent(prop, accuracy = 0.1)),
            position = position_dodge(0.8), vjust = -0.3, size = 3.5) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = c("Breast Cancer Case" = "#1f77b4", "Control" = "#ff7f0e")) +
  labs(
    title = "Family History of Breast/Ovarian Cancer (Postmenopausal - Old)",
    x = "Family History", y = "Proportion", fill = "Group"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    legend.position = "top"
  )

```

## BMI
```{r}
table(data_postmeno$bmicat
      [as.character(data_postmeno$br) == "Control"])
sum(table(data_postmeno$bmicat
          [as.character(data_postmeno$br) == "Control"]))
prop.table(table(data_postmeno$bmicat
                 [as.character(data_postmeno$br) == "Control"]))

table(data_postmeno$bmicat
      [as.character(data_postmeno$br) == "Breast Cancer Case"])
sum(table(data_postmeno$bmicat
          [as.character(data_postmeno$br) == "Breast Cancer Case"]))
prop.table(table(data_postmeno$bmicat
                 [as.character(data_postmeno$br) == "Breast Cancer Case"]))

# Prepare data for BMI category (Postmenopausal - Old)
plot_data_bmicat_old <- data_postmeno %>%
  filter(br %in% c("Control", "Breast Cancer Case")) %>%
  mutate(br = as.character(br)) %>%
  count(br, bmicat) %>%
  group_by(br) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()

# Plot
ggplot(plot_data_bmicat_old, aes(x = bmicat, y = prop, fill = br)) +
  geom_bar(stat = "identity", position = position_dodge(0.8), width = 0.7) +
  geom_text(aes(label = scales::percent(prop, accuracy = 0.1)),
            position = position_dodge(0.8), vjust = -0.3, size = 3.5) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = c("Breast Cancer Case" = "#1f77b4", "Control" = "#ff7f0e")) +
  labs(
    title = "BMI Category (Postmenopausal - Old)",
    x = "BMI Category", y = "Proportion", fill = "Group"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    legend.position = "top"
  )

```

## Height
```{r}
table(data_postmeno$catht
      [as.character(data_postmeno$br) == "Control"])
sum(table(data_postmeno$catht
          [as.character(data_postmeno$br) == "Control"]))
prop.table(table(data_postmeno$catht
                 [as.character(data_postmeno$br) == "Control"]))

table(data_postmeno$catht
      [as.character(data_postmeno$br) == "Breast Cancer Case"])
sum(table(data_postmeno$catht
          [as.character(data_postmeno$br) == "Breast Cancer Case"]))
prop.table(table(data_postmeno$catht
                 [as.character(data_postmeno$br) == "Breast Cancer Case"]))

# Prepare data for catht (Postmenopausal - Old)
plot_data_catht_old <- data_postmeno %>%
  filter(br %in% c("Control", "Breast Cancer Case")) %>%
  mutate(br = as.character(br)) %>%
  count(br, catht) %>%
  group_by(br) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()

# Plot
ggplot(plot_data_catht_old, aes(x = catht, y = prop, fill = br)) +
  geom_bar(stat = "identity", position = position_dodge(0.8), width = 0.7) +
  geom_text(aes(label = scales::percent(prop, accuracy = 0.1)),
            position = position_dodge(0.8), vjust = -0.3, size = 3.5) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = c("Breast Cancer Case" = "#1f77b4", "Control" = "#ff7f0e")) +
  labs(
    title = "Cat-height Category (Postmenopausal - Old)",
    x = "Cat-height Category", y = "Proportion", fill = "Group"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    legend.position = "top"
  )

```

## Waist-to-hip ratio
```{r}
table(data_postmeno$ratio
      [as.character(data_postmeno$br) == "Control"])
sum(table(data_postmeno$ratio
          [as.character(data_postmeno$br) == "Control"]))
prop.table(table(data_postmeno$ratio
                 [as.character(data_postmeno$br) == "Control"]))

table(data_postmeno$ratio
      [as.character(data_postmeno$br) == "Breast Cancer Case"])
sum(table(data_postmeno$ratio
          [as.character(data_postmeno$br) == "Breast Cancer Case"]))
prop.table(table(data_postmeno$ratio
                 [as.character(data_postmeno$br) == "Breast Cancer Case"]))

# Prepare data for ratio (Postmenopausal - Old)
plot_data_ratio_old <- data_postmeno %>%
  filter(br %in% c("Control", "Breast Cancer Case")) %>%
  mutate(br = as.character(br)) %>%
  count(br, ratio) %>%
  group_by(br) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()

# Plot
ggplot(plot_data_ratio_old, aes(x = ratio, y = prop, fill = br)) +
  geom_bar(stat = "identity", position = position_dodge(0.8), width = 0.7) +
  geom_text(aes(label = scales::percent(prop, accuracy = 0.1)),
            position = position_dodge(0.8), vjust = -0.3, size = 3.5) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = c("Breast Cancer Case" = "#1f77b4", "Control" = "#ff7f0e")) +
  labs(
    title = "Ratio Distribution (Postmenopausal - Old)",
    x = "Ratio", y = "Proportion", fill = "Group"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    legend.position = "top"
  )

```

<!-- ## Duration of diabetes -->
<!-- ```{r} -->
<!-- table(data_postmeno$diabetic_years_cat -->
<!--       [as.character(data_postmeno$br) == "Control"]) -->
<!-- sum(table(data_postmeno$diabetic_years_cat -->
<!--           [as.character(data_postmeno$br) == "Control"])) -->
<!-- prop.table(table(data_postmeno$diabetic_years_cat -->
<!--                  [as.character(data_postmeno$br) == "Control"])) -->

<!-- table(data_postmeno$diabetic_years_cat -->
<!--       [as.character(data_postmeno$br) == "Breast Cancer Case"]) -->
<!-- sum(table(data_postmeno$diabetic_years_cat -->
<!--           [as.character(data_postmeno$br) == "Breast Cancer Case"])) -->
<!-- prop.table(table(data_postmeno$diabetic_years_cat -->
<!--                  [as.character(data_postmeno$br) == "Breast Cancer Case"])) -->
<!-- ``` -->
<!-- ## New variable for duration of diabetes collapsing the levels 0 and 1 -->
<!-- ```{r} -->
<!-- data_postmeno$diabetes_duration = data_postmeno$diabetic_years_cat -->
<!-- data_postmeno$diabetes_duration[data_postmeno$diabetes_duration <= 1] = 1 -->
<!-- ``` -->

<!-- ## Diabetes -->
<!-- ```{r} -->
<!-- table(data_postmeno$diabetes_YN -->
<!--       [as.character(data_postmeno$br) == "Control"]) -->
<!-- sum(table(data_postmeno$diabetes_YN -->
<!--           [as.character(data_postmeno$br) == "Control"])) -->
<!-- prop.table(table(data_postmeno$diabetes_YN -->
<!--                  [as.character(data_postmeno$br) == "Control"])) -->

<!-- table(data_postmeno$diabetes_YN -->
<!--       [as.character(data_postmeno$br) == "Breast Cancer Case"]) -->
<!-- sum(table(data_postmeno$diabetes_YN -->
<!--           [as.character(data_postmeno$br) == "Breast Cancer Case"])) -->
<!-- prop.table(table(data_postmeno$diabetes_YN -->
<!--                  [as.character(data_postmeno$br) == "Breast Cancer Case"])) -->

<!-- # Prepare data for diabetes_YN (Postmenopausal - Old) -->
<!-- plot_data_diabetes_old <- data_postmeno %>% -->
<!--   filter(br %in% c("Control", "Breast Cancer Case")) %>% -->
<!--   mutate(br = as.character(br)) %>% -->
<!--   count(br, diabetes_YN) %>% -->
<!--   group_by(br) %>% -->
<!--   mutate(prop = n / sum(n)) %>% -->
<!--   ungroup() -->

<!-- # Plot -->
<!-- ggplot(plot_data_diabetes_old, aes(x = diabetes_YN, y = prop, fill = br)) + -->
<!--   geom_bar(stat = "identity", position = position_dodge(0.8), width = 0.7) + -->
<!--   geom_text(aes(label = scales::percent(prop, accuracy = 0.1)), -->
<!--             position = position_dodge(0.8), vjust = -0.3, size = 3.5) + -->
<!--   scale_y_continuous(labels = scales::percent_format()) + -->
<!--   scale_fill_manual(values = c("Breast Cancer Case" = "#1f77b4", "Control" = "#ff7f0e")) + -->
<!--   labs( -->
<!--     title = "Diabetes Y/N Distribution (Postmenopausal - Old)", -->
<!--     x = "Diabetes (Y/N)", y = "Proportion", fill = "Group" -->
<!--   ) + -->
<!--   theme_minimal(base_size = 14) + -->
<!--   theme( -->
<!--     plot.title = element_text(face = "bold", hjust = 0.5), -->
<!--     legend.position = "top" -->
<!--   ) -->

<!-- ``` -->

<!-- ## Family history of diabetes -->
<!-- ```{r} -->
<!-- table(data_postmeno$F_H_diabetes -->
<!--       [as.character(data_postmeno$br) == "Control"]) -->
<!-- sum(table(data_postmeno$F_H_diabetes -->
<!--           [as.character(data_postmeno$br) == "Control"])) -->
<!-- prop.table(table(data_postmeno$F_H_diabetes -->
<!--                  [as.character(data_postmeno$br) == "Control"])) -->

<!-- table(data_postmeno$F_H_diabetes -->
<!--       [as.character(data_postmeno$br) == "Breast Cancer Case"]) -->
<!-- sum(table(data_postmeno$F_H_diabetes -->
<!--           [as.character(data_postmeno$br) == "Breast Cancer Case"])) -->
<!-- prop.table(table(data_postmeno$F_H_diabetes -->
<!--                  [as.character(data_postmeno$br) == "Breast Cancer Case"])) -->

<!-- # Prepare data for F_H_diabetes (Postmenopausal - Old) -->
<!-- plot_data_fh_diabetes_old <- data_postmeno %>% -->
<!--   filter(br %in% c("Control", "Breast Cancer Case")) %>% -->
<!--   mutate(br = as.character(br)) %>% -->
<!--   count(br, F_H_diabetes) %>% -->
<!--   group_by(br) %>% -->
<!--   mutate(prop = n / sum(n)) %>% -->
<!--   ungroup() -->

<!-- # Plot -->
<!-- ggplot(plot_data_fh_diabetes_old, aes(x = F_H_diabetes, y = prop, fill = br)) + -->
<!--   geom_bar(stat = "identity", position = position_dodge(0.8), width = 0.7) + -->
<!--   geom_text(aes(label = scales::percent(prop, accuracy = 0.1)), -->
<!--             position = position_dodge(0.8), vjust = -0.3, size = 3.5) + -->
<!--   scale_y_continuous(labels = scales::percent_format()) + -->
<!--   scale_fill_manual(values = c("Breast Cancer Case" = "#1f77b4", "Control" = "#ff7f0e")) + -->
<!--   labs( -->
<!--     title = "Family History of Diabetes (Postmenopausal - Old)", -->
<!--     x = "Family History of Diabetes", y = "Proportion", fill = "Group" -->
<!--   ) + -->
<!--   theme_minimal(base_size = 14) + -->
<!--   theme( -->
<!--     plot.title = element_text(face = "bold", hjust = 0.5), -->
<!--     legend.position = "top" -->
<!--   ) -->

<!-- ``` -->


## BMI and WHR interaction
```{r}
table(data_postmeno$bmicat
      [as.character(data_postmeno$br) == "Control"], 
      data_postmeno$ratio
      [as.character(data_postmeno$br) == "Control"])
sum(table(data_postmeno$bmicat
          [as.character(data_postmeno$br) == "Control"], 
          data_postmeno$ratio
          [as.character(data_postmeno$br) == "Control"]))
prop.table(table(data_postmeno$bmicat
                 [as.character(data_postmeno$br) == "Control"], 
                 data_postmeno$ratio
                 [as.character(data_postmeno$br) == "Control"]))


table(data_postmeno$bmicat
      [as.character(data_postmeno$br) == "Breast Cancer Case"], 
      data_postmeno$ratio
      [as.character(data_postmeno$br) == "Breast Cancer Case"])
sum(table(data_postmeno$bmicat
          [as.character(data_postmeno$br) == "Breast Cancer Case"], 
          data_postmeno$ratio
          [as.character(data_postmeno$br) == "Breast Cancer Case"]))
prop.table(table(data_postmeno$bmicat
                 [as.character(data_postmeno$br) == "Breast Cancer Case"], 
                 data_postmeno$ratio
                 [as.character(data_postmeno$br) == "Breast Cancer Case"]))
```

<!-- ## BMI and diabetes interaction -->
<!-- ```{r} -->
<!-- table(data_postmeno$bmicat -->
<!--       [as.character(data_postmeno$br) == "Control"],  -->
<!--       data_postmeno$diabetes_YN -->
<!--       [as.character(data_postmeno$br) == "Control"]) -->
<!-- sum(table(data_postmeno$bmicat -->
<!--           [as.character(data_postmeno$br) == "Control"],  -->
<!--           data_postmeno$diabetes_YN -->
<!--           [as.character(data_postmeno$br) == "Control"])) -->
<!-- prop.table(table(data_postmeno$bmicat -->
<!--                  [as.character(data_postmeno$br) == "Control"],  -->
<!--                  data_postmeno$diabetes_YN -->
<!--                  [as.character(data_postmeno$br) == "Control"])) -->


<!-- table(data_postmeno$bmicat -->
<!--       [as.character(data_postmeno$br) == "Breast Cancer Case"],  -->
<!--       data_postmeno$diabetes_YN -->
<!--       [as.character(data_postmeno$br) == "Breast Cancer Case"]) -->
<!-- sum(table(data_postmeno$bmicat -->
<!--           [as.character(data_postmeno$br) == "Breast Cancer Case"],  -->
<!--           data_postmeno$diabetes_YN -->
<!--           [as.character(data_postmeno$br) == "Breast Cancer Case"])) -->
<!-- prop.table(table(data_postmeno$bmicat -->
<!--                  [as.character(data_postmeno$br) == "Breast Cancer Case"],  -->
<!--                  data_postmeno$diabetes_YN -->
<!--                  [as.character(data_postmeno$br) == "Breast Cancer Case"])) -->

<!-- ``` -->

<!-- ## Categorical age at enrollment -->
<!-- ```{r} -->
<!-- table(data_postmeno$agecode -->
<!--       [as.character(data_postmeno$br) == "Control"]) -->
<!-- sum(table(data_postmeno$agecode -->
<!--           [as.character(data_postmeno$br) == "Control"])) -->
<!-- prop.table(table(data_postmeno$agecode -->
<!--                  [as.character(data_postmeno$br) == "Control"])) -->

<!-- table(data_postmeno$agecode -->
<!--       [as.character(data_postmeno$br) == "Breast Cancer Case"]) -->
<!-- sum(table(data_postmeno$agecode -->
<!--           [as.character(data_postmeno$br) == "Breast Cancer Case"])) -->
<!-- prop.table(table(data_postmeno$agecode -->
<!--                  [as.character(data_postmeno$br) == "Breast Cancer Case"])) -->
<!-- ``` -->

<!-- ## Geographical location -->
<!-- ```{r} -->
<!-- table(data_postmeno$area -->
<!--       [as.character(data_postmeno$br) == "Control"]) -->
<!-- sum(table(data_postmeno$area -->
<!--           [as.character(data_postmeno$br) == "Control"])) -->
<!-- prop.table(table(data_postmeno$area -->
<!--                  [as.character(data_postmeno$br) == "Control"])) -->

<!-- table(data_postmeno$area -->
<!--       [as.character(data_postmeno$br) == "Breast Cancer Case"]) -->
<!-- sum(table(data_postmeno$area -->
<!--           [as.character(data_postmeno$br) == "Breast Cancer Case"])) -->
<!-- prop.table(table(data_postmeno$area -->
<!--                  [as.character(data_postmeno$br) == "Breast Cancer Case"])) -->

<!-- # Prepare data for area (Postmenopausal - Old) -->
<!-- plot_data_area_old <- data_postmeno %>% -->
<!--   filter(br %in% c("Control", "Breast Cancer Case")) %>% -->
<!--   mutate(br = as.character(br)) %>% -->
<!--   count(br, area) %>% -->
<!--   group_by(br) %>% -->
<!--   mutate(prop = n / sum(n)) %>% -->
<!--   ungroup() -->

<!-- # Plot -->
<!-- ggplot(plot_data_area_old, aes(x = area, y = prop, fill = br)) + -->
<!--   geom_bar(stat = "identity", position = position_dodge(0.8), width = 0.7) + -->
<!--   geom_text(aes(label = scales::percent(prop, accuracy = 0.1)), -->
<!--             position = position_dodge(0.8), vjust = -0.3, size = 3.5) + -->
<!--   scale_y_continuous(labels = scales::percent_format()) + -->
<!--   scale_fill_manual(values = c("Breast Cancer Case" = "#1f77b4", "Control" = "#ff7f0e")) + -->
<!--   labs( -->
<!--     title = "Area of Residence (Postmenopausal - Old)", -->
<!--     x = "Area of Residence", y = "Proportion", fill = "Group" -->
<!--   ) + -->
<!--   theme_minimal(base_size = 14) + -->
<!--   theme( -->
<!--     plot.title = element_text(face = "bold", hjust = 0.5), -->
<!--     legend.position = "top" -->
<!--   ) -->

<!-- ``` -->

<!-- ## Rural/urban based on first 20 years -->
<!-- ```{r} -->
<!-- table(data_postmeno$residancefirsttwentyyears -->
<!--       [as.character(data_postmeno$br) == "Control"]) -->
<!-- sum(table(data_postmeno$residancefirsttwentyyears -->
<!--           [as.character(data_postmeno$br) == "Control"])) -->
<!-- prop.table(table(data_postmeno$residancefirsttwentyyears -->
<!--                  [as.character(data_postmeno$br) == "Control"])) -->

<!-- table(data_postmeno$residancefirsttwentyyears -->
<!--       [as.character(data_postmeno$br) == "Breast Cancer Case"]) -->
<!-- sum(table(data_postmeno$residancefirsttwentyyears -->
<!--           [as.character(data_postmeno$br) == "Breast Cancer Case"])) -->
<!-- prop.table(table(data_postmeno$residancefirsttwentyyears -->
<!--                  [as.character(data_postmeno$br) == "Breast Cancer Case"])) -->

<!-- # Prepare data for residence in first 20 years (Postmenopausal - Old) -->
<!-- plot_data_residency_old <- data_postmeno %>% -->
<!--   filter(br %in% c("Control", "Breast Cancer Case")) %>% -->
<!--   mutate(br = as.character(br)) %>% -->
<!--   count(br, residancefirsttwentyyears) %>% -->
<!--   group_by(br) %>% -->
<!--   mutate(prop = n / sum(n)) %>% -->
<!--   ungroup() -->

<!-- # Plot -->
<!-- ggplot(plot_data_residency_old, aes(x = residancefirsttwentyyears, y = prop, fill = br)) + -->
<!--   geom_bar(stat = "identity", position = position_dodge(0.8), width = 0.7) + -->
<!--   geom_text(aes(label = scales::percent(prop, accuracy = 0.1)), -->
<!--             position = position_dodge(0.8), vjust = -0.3, size = 3.5) + -->
<!--   scale_y_continuous(labels = scales::percent_format()) + -->
<!--   scale_fill_manual(values = c("Breast Cancer Case" = "#1f77b4", "Control" = "#ff7f0e")) + -->
<!--   labs( -->
<!--     title = "Residence in First 20 Years (Postmenopausal - Old)", -->
<!--     x = "Residence in First 20 Years", y = "Proportion", fill = "Group" -->
<!--   ) + -->
<!--   theme_minimal(base_size = 14) + -->
<!--   theme( -->
<!--     plot.title = element_text(face = "bold", hjust = 0.5), -->
<!--     legend.position = "top" -->
<!--   ) -->

<!-- ``` -->

<!-- ## Binary education status: <8 years and >= 8 years of schooling -->
<!-- ```{r} -->
<!-- table(data_postmeno$ed -->
<!--       [as.character(data_postmeno$br) == "Control"]) -->
<!-- sum(table(data_postmeno$ed -->
<!--           [as.character(data_postmeno$br) == "Control"])) -->
<!-- prop.table(table(data_postmeno$ed -->
<!--                  [as.character(data_postmeno$br) == "Control"])) -->

<!-- table(data_postmeno$ed -->
<!--       [as.character(data_postmeno$br) == "Breast Cancer Case"]) -->
<!-- sum(table(data_postmeno$ed -->
<!--           [as.character(data_postmeno$br) == "Breast Cancer Case"])) -->
<!-- prop.table(table(data_postmeno$ed -->
<!--                  [as.character(data_postmeno$br) == "Breast Cancer Case"])) -->

<!-- # Prepare data for education level (Postmenopausal - Old) -->
<!-- plot_data_ed_old <- data_postmeno %>% -->
<!--   filter(br %in% c("Control", "Breast Cancer Case")) %>% -->
<!--   mutate(br = as.character(br)) %>% -->
<!--   count(br, ed) %>% -->
<!--   group_by(br) %>% -->
<!--   mutate(prop = n / sum(n)) %>% -->
<!--   ungroup() -->

<!-- # Plot -->
<!-- ggplot(plot_data_ed_old, aes(x = ed, y = prop, fill = br)) + -->
<!--   geom_bar(stat = "identity", position = position_dodge(0.8), width = 0.7) + -->
<!--   geom_text(aes(label = scales::percent(prop, accuracy = 0.1)), -->
<!--             position = position_dodge(0.8), vjust = -0.3, size = 3.5) + -->
<!--   scale_y_continuous(labels = scales::percent_format()) + -->
<!--   scale_fill_manual(values = c("Breast Cancer Case" = "#1f77b4", "Control" = "#ff7f0e")) + -->
<!--   labs( -->
<!--     title = "Education Level (Postmenopausal - Old)", -->
<!--     x = "Education Level", y = "Proportion", fill = "Group" -->
<!--   ) + -->
<!--   theme_minimal(base_size = 14) + -->
<!--   theme( -->
<!--     plot.title = element_text(face = "bold", hjust = 0.5), -->
<!--     legend.position = "top" -->
<!--   ) -->

<!-- ``` -->

##Relative risk model for postmenopausal women
```{r}
model_postmeno = glm(br ~ as.factor(ageatfirstfulltermpreg_cat)
                    + interval_menarche_firstbirth_cat
                    + as.factor(fulltermpreg_cat)
                    + oc + as.factor(mean_brfeed_duration_cat)
                    + as.factor(totalabortion_cat)
                    + as.factor(totalmiscarriage_cat)
                    + as.factor(catht) + as.factor(bmicat) * as.factor(ratio),
                    data = data_postmeno, family = binomial(link=logit),
                    na.action = na.omit)

# as.factor(agecode) + as.factor(area) 
#                     + as.factor(ed) + as.factor(residancefirsttwentyyears) + as.factor(catmenopause)
#                     + 

cbind(exp(coef(model_postmeno)),exp(confint(model_postmeno)))

summary(model_postmeno)
```

## Joint association of BMI and WHR#####
```{r}
#Log relative risks and variances
estimate = coef(model_postmeno)[c("as.factor(bmicat)2","as.factor(bmicat)3",
                                 "as.factor(ratio)1","as.factor(ratio)2",
                                 "as.factor(bmicat)2:as.factor(ratio)1",
                                 "as.factor(bmicat)3:as.factor(ratio)1",
                                 "as.factor(bmicat)2:as.factor(ratio)2",
                                 "as.factor(bmicat)3:as.factor(ratio)2")]

variance = vcov(model_postmeno)[c("as.factor(bmicat)2","as.factor(bmicat)3",
                                 "as.factor(ratio)1","as.factor(ratio)2",
                                 "as.factor(bmicat)2:as.factor(ratio)1",
                                 "as.factor(bmicat)3:as.factor(ratio)1",
                                 "as.factor(bmicat)2:as.factor(ratio)2",
                                 "as.factor(bmicat)3:as.factor(ratio)2"),
                               c("as.factor(bmicat)2","as.factor(bmicat)3",
                                 "as.factor(ratio)1","as.factor(ratio)2",
                                 "as.factor(bmicat)2:as.factor(ratio)1",
                                 "as.factor(bmicat)3:as.factor(ratio)1",
                                 "as.factor(bmicat)2:as.factor(ratio)2",
                                 "as.factor(bmicat)3:as.factor(ratio)2")]

#BMI < 24.9, WHR < 0.84

RR1 = 1

ci1 = c(1,1)

#BMI < 24.9, WHR 0.85-0.94

RR2 = exp(coef(model_postmeno))["as.factor(ratio)1"]

ci2 = exp(confint(model_postmeno))["as.factor(ratio)1",]

#BMI < 24.9, WHR >= 0.95

RR3 = exp(coef(model_postmeno))["as.factor(ratio)2"]

ci3 = exp(confint(model_postmeno))["as.factor(ratio)2",]

#BMI 25-29.9, WHR < 0.84

RR4 = exp(coef(model_postmeno))["as.factor(bmicat)2"]

ci4 = exp(confint(model_postmeno))["as.factor(bmicat)2",]

#BMI 25-29.9, WHR 0.85-0.94

RR5 = exp(coef(model_postmeno))["as.factor(bmicat)2"] * 
  exp(coef(model_postmeno))["as.factor(ratio)1"] * 
  exp(coef(model_postmeno))["as.factor(bmicat)2:as.factor(ratio)1"]

v5 = c(1,0,1,0,1,0,0,0)

RR5 = exp(v5 %*% estimate)

sd5 = sqrt(v5 %*% variance %*% v5)

ci5 = exp(c(log(RR5) - 1.96 * sd5, log(RR5) + 1.96 * sd5))


#BMI 25-29.9, WHR >= 0.95

RR6 = exp(coef(model_postmeno))["as.factor(bmicat)2"] * 
  exp(coef(model_postmeno))["as.factor(ratio)2"] * 
  exp(coef(model_postmeno))["as.factor(bmicat)2:as.factor(ratio)2"]

v6 = c(1,0,0,1,0,0,1,0)

RR6 = exp(v6 %*% estimate)

sd6 = sqrt(v6 %*% variance %*% v6)

ci6 = exp(c(log(RR6) - 1.96 * sd6, log(RR6) + 1.96 * sd6))

#BMI >= 30, WHR < 0.84

RR7 = exp(coef(model_postmeno))["as.factor(bmicat)3"]

ci7 = exp(confint(model_postmeno))["as.factor(bmicat)3",]

#BMI >= 30, WHR 0.85-0.94

RR8 = exp(coef(model_postmeno))["as.factor(bmicat)3"] * 
  exp(coef(model_postmeno))["as.factor(ratio)1"] * 
  exp(coef(model_postmeno))["as.factor(bmicat)3:as.factor(ratio)1"]

v8 = c(0,1,1,0,0,1,0,0)

RR8 = exp(v8 %*% estimate)

sd8 = sqrt(v8 %*% variance %*% v8)

ci8 = exp(c(log(RR8) - 1.96 * sd8, log(RR8) + 1.96 * sd8))

#BMI >= 30, WHR >= 0.95

RR9 = exp(coef(model_postmeno))["as.factor(bmicat)3"] * 
  exp(coef(model_postmeno))["as.factor(ratio)2"] * 
  exp(coef(model_postmeno))["as.factor(bmicat)3:as.factor(ratio)2"]

v9 = c(0,1,0,1,0,0,0,1)

RR9 = exp(v9 %*% estimate)

sd9 = sqrt(v9 %*% variance %*% v9)

ci9 = exp(c(log(RR9) - 1.96 * sd9, log(RR9) + 1.96 * sd9))

BMI_WHR_Postmeno = cbind(c(RR1,RR2,RR3,RR4,RR5,RR6,RR7,RR8,RR9),
                        rbind(ci1,ci2,ci3,ci4,ci5,ci6,ci7,ci8,ci9))

rownames(BMI_WHR_Postmeno)[c(1,5,6,8,9)] = c("Ref","as.factor(bmicat)2:as.factor(ratio)1",
                                            "as.factor(bmicat)2:as.factor(ratio)2",
                                            "as.factor(bmicat)3:as.factor(ratio)1",
                                            "as.factor(bmicat)2:as.factor(ratio)2")


#pdf(file = "BMI_WHR_Interaction_Postmeno")
plotCI(x = BMI_WHR_Postmeno[,1], ui = BMI_WHR_Postmeno[,3], li = BMI_WHR_Postmeno[,2],
       xlab = "", ylab = "Joint Relative Risk",pch=16,lwd = 2,xaxt = "n",
       col = c(rep("red",3),rep("blue",3),rep("green",3)))

mtext(expression("BMI"<="24.9"), side = 3, line = 1, at = 2, col = "red")
mtext(expression("BMI 25-29.9"), side = 3, line = 1, at = 5, col = "blue")
mtext(expression("BMI">="30"), side = 3, line = 1, at = 8, col = "green")

mtext(rep(expression("WHR"<=0.84),3), side = 1, line = 0.2, las = 2, 
      at = c(1,4,7), col = c("red","blue","green"), cex = 0.7)
mtext(rep("WHR 0.85-0.94",3), side = 1, line = 0.2, las = 2, 
      at = c(2,5,8), col = c("red","blue","green"), cex = 0.7)
mtext(rep(expression("WHR">=0.95),3), side = 1, line = 0.2, las = 2, 
      at = c(3,6,9), col = c("red","blue","green"), cex = 0.7)
#dev.off()
```


# Absolute risk prediction using iCARE

## Function to format incidence and mortality rates
```{r}
formatted_rates = function(mat){
  if(ncol(mat)==3){
    start <- mat[,1]
    end   <- mat[,2]
    rate  <- mat[,3]
    
    integer_ages <- seq(min(start), max(end))
    formatted    <- cbind( integer_ages, rep(NA, length(integer_ages)))
    for(i in 1:nrow(mat)){
      these              <- which(formatted[,1]>=start[i] & 
                                    formatted[,1]<=end[i])
      #formatted[these,2] <- rate[i]/length(these)
      formatted[these,2] <- rate[i]
    }
    colnames(formatted) <- c("ages", "rates")
    formatted
  }else{
    formatted = mat
  }
  formatted
}
```

## Formatting breast cancer incidence rate 
```{r}
setwd("/Users/anaghchattopadhyay/Documents/Nilanjan\ Project/TMC_BCProject/Data")

bc_inci = read.xlsx("BC_Incidence_Rates.xlsx")
bc_inci = formatted_rates(bc_inci)
bc_inci[,2] = bc_inci[,2]/100000
```

## Formatting overall mortality rate
```{r}
setwd("/Users/anaghchattopadhyay/Documents/Nilanjan\ Project/TMC_BCProject/Data")
mort_inci = read.xlsx("Overall_Mortality_Rates.xlsx")
mort_inci = formatted_rates(mort_inci)
mort_inci[,2] = mort_inci[,2]/1000
```

## Creating the reference dataset
```{r}
setwd("/Users/anaghchattopadhyay/Documents/Nilanjan\ Project/TMC_BCProject/Data")
ref_data = read.xlsx("Breast_1.12.22.xlsx")
test_data = read.xlsx("parichoy_br_data_6.2.23.xlsx")
```

## Categories for age at first full term pregnancy
```{r}
ref_data$ageatfirstfulltermpreg_cat = 
  ref_data$age_at_first_fullterm_preg
ref_data$ageatfirstfulltermpreg_cat[which(ref_data$age_at_first_fullterm_preg < 20)] = 1
ref_data$ageatfirstfulltermpreg_cat[which(ref_data$age_at_first_fullterm_preg == 20 
                                          | ref_data$age_at_first_fullterm_preg == 21)] = 2
ref_data$ageatfirstfulltermpreg_cat[which(ref_data$age_at_first_fullterm_preg == 22 
                                          | ref_data$age_at_first_fullterm_preg == 23)] = 3
ref_data$ageatfirstfulltermpreg_cat[which(ref_data$age_at_first_fullterm_preg == 24 
                                          | ref_data$age_at_first_fullterm_preg == 25)] = 4
ref_data$ageatfirstfulltermpreg_cat[which(ref_data$age_at_first_fullterm_preg >= 26)] = 5


test_data$ageatfirstfulltermpreg_cat = 
  test_data$age_at_first_fullterm_preg
test_data$ageatfirstfulltermpreg_cat[which(test_data$age_at_first_fullterm_preg < 20)] = 1
test_data$ageatfirstfulltermpreg_cat[which(test_data$age_at_first_fullterm_preg == 20 
                                          | test_data$age_at_first_fullterm_preg == 21)] = 2
test_data$ageatfirstfulltermpreg_cat[which(test_data$age_at_first_fullterm_preg == 22 
                                          | test_data$age_at_first_fullterm_preg == 23)] = 3
test_data$ageatfirstfulltermpreg_cat[which(test_data$age_at_first_fullterm_preg == 24 
                                          | test_data$age_at_first_fullterm_preg == 25)] = 4
test_data$ageatfirstfulltermpreg_cat[which(test_data$age_at_first_fullterm_preg >= 26)] = 5

table(test_data$ageatfirstfulltermpreg_cat)
sum(is.na(test_data$age_at_first_fullterm_preg))

```

##Categories of interval between menarche and full term pregnancy
```{r}
ref_data$interval_menarche_firstbirth_cat = 
  ref_data$durn_menarch_first_fullterm_preg
ref_data$interval_menarche_firstbirth_cat[ref_data$durn_menarch_first_fullterm_preg < 10] = 0
ref_data$interval_menarche_firstbirth_cat[ref_data$durn_menarch_first_fullterm_preg >= 10] = 1

test_data$interval_menarche_firstbirth_cat = 
  test_data$durn_menarch_first_fullterm_preg
test_data$interval_menarche_firstbirth_cat[test_data$durn_menarch_first_fullterm_preg < 10] = 0
test_data$interval_menarche_firstbirth_cat[test_data$durn_menarch_first_fullterm_preg >= 10] = 1

table(test_data$interval_menarche_firstbirth_cat)
sum(is.na(test_data$interval_menarche_firstbirth_cat))
```

## Categories for number of full term pregnancies
```{r}
ref_data$fulltermpreg_cat = ref_data$No_of_live_birth
ref_data$fulltermpreg_cat[ref_data$No_of_live_birth == 0] = 1
ref_data$fulltermpreg_cat[ref_data$No_of_live_birth >= 4] = 4

test_data$fulltermpreg_cat = test_data$No_of_live_birth
test_data$fulltermpreg_cat[test_data$No_of_live_birth == 0] = 1
test_data$fulltermpreg_cat[test_data$No_of_live_birth >= 4] = 4

table(test_data$fulltermpreg_cat)
sum(is.na(test_data$fulltermpreg_cat))

```

## OC use
```{r}
ref_data$oc = NA
ref_data$oc[ref_data$oc_use_yn == 2] = 0
ref_data$oc[ref_data$oc_use_yn == 1] = 1

test_data$oc = NA
test_data$oc[test_data$oc_use_yn == 2] = 0
test_data$oc[test_data$oc_use_yn == 1] = 1

table(test_data$oc)
sum(is.na(test_data$oc))
```

## Categories for number of induced abortions
```{r}
ref_data$totalabortion_cat = ref_data$count_abortion
ref_data$totalabortion_cat[ref_data$count_abortion >= 2] = 2

test_data$totalabortion_cat = test_data$count_abortion
test_data$totalabortion_cat[test_data$count_abortion >= 2] = 2

table(test_data$totalabortion_cat)
sum(is.na(test_data$totalabortion_cat))
```

## Categories for number of spontaneous abortions
```{r}
ref_data$totalmiscarriage_cat = ref_data$count_miscarriage
ref_data$totalmiscarriage_cat[ref_data$count_miscarriage >= 2] = 2

test_data$totalmiscarriage_cat = test_data$count_miscarriage
test_data$totalmiscarriage_cat[test_data$count_miscarriage >= 2] = 2

table(test_data$totalmiscarriage_cat)
sum(is.na(test_data$totalmiscarriage_cat))
```

## Categories for duration of breastfeeding
```{r}
ref_data$mean_brfeed_duration_cat = ref_data$avg_dur_breastfeeding
ref_data$mean_brfeed_duration_cat[ref_data$avg_dur_breastfeeding < 12] = 1
ref_data$mean_brfeed_duration_cat[ref_data$avg_dur_breastfeeding >= 12 
                                  & ref_data$avg_dur_breastfeeding < 24] = 2
ref_data$mean_brfeed_duration_cat[ref_data$avg_dur_breastfeeding >= 24 
                                  & ref_data$avg_dur_breastfeeding < 36] = 3
ref_data$mean_brfeed_duration_cat[ref_data$avg_dur_breastfeeding >= 36] = 4

test_data$mean_brfeed_duration_cat = test_data$avg_dur_breastfeeding
test_data$mean_brfeed_duration_cat[test_data$avg_dur_breastfeeding < 12] = 1
test_data$mean_brfeed_duration_cat[test_data$avg_dur_breastfeeding >= 12 
                                  & test_data$avg_dur_breastfeeding < 24] = 2
test_data$mean_brfeed_duration_cat[test_data$avg_dur_breastfeeding >= 24 
                                  & test_data$avg_dur_breastfeeding < 36] = 3
test_data$mean_brfeed_duration_cat[test_data$avg_dur_breastfeeding >= 36] = 4

table(test_data$mean_brfeed_duration_cat)
sum(is.na(test_data$mean_brfeed_duration_cat))
```

## BMI
```{r}
ref_data$bmicat = ref_data$BMI
ref_data$bmicat[ref_data$BMI < 25] = 1
ref_data$bmicat[ref_data$BMI >= 25 & ref_data$BMI < 30] = 2
ref_data$bmicat[ref_data$BMI >= 30] = 3

test_data$bmicat = test_data$BMI
test_data$bmicat[test_data$BMI < 25] = 1
test_data$bmicat[test_data$BMI >= 25 & test_data$BMI < 30] = 2
test_data$bmicat[test_data$BMI >= 30] = 3

table(test_data$bmicat )
sum(is.na(test_data$bmicat))
```

## Height
```{r}
ref_data$catht = ref_data$Avg_Height
ref_data$catht[ref_data$Avg_Height <= 150] = 0
ref_data$catht[ref_data$Avg_Height > 150 & 
                 ref_data$Avg_Height <= 155] = 1
ref_data$catht[ref_data$Avg_Height > 155 & 
                 ref_data$Avg_Height <= 160] = 2
ref_data$catht[ref_data$Avg_Height > 160] = 3

test_data$catht = test_data$Avg_Height
test_data$catht[test_data$Avg_Height <= 150] = 0
test_data$catht[test_data$Avg_Height > 150 & 
                 test_data$Avg_Height <= 155] = 1
test_data$catht[test_data$Avg_Height > 155 & 
                 test_data$Avg_Height <= 160] = 2
test_data$catht[test_data$Avg_Height > 160] = 3

table(test_data$catht)
sum(is.na(test_data$catht))
```

## WHR
```{r}
ref_data$ratio = ref_data$waist_to_hip
ref_data$ratio[ref_data$waist_to_hip <= 0.84] = 0
ref_data$ratio[ref_data$waist_to_hip > 0.84 & 
                 ref_data$waist_to_hip <= 0.94] = 1
ref_data$ratio[ref_data$waist_to_hip > 0.94] = 2

test_data$ratio = test_data$waist_to_hip
test_data$ratio[test_data$waist_to_hip <= 0.84] = 0
test_data$ratio[test_data$waist_to_hip > 0.84 & 
                 test_data$waist_to_hip <= 0.94] = 1
test_data$ratio[test_data$waist_to_hip > 0.94] = 2

table(test_data$ratio)
sum(is.na(test_data$ratio))
```

## Age at menopause
```{r}
ref_data$catmenopause = ref_data$Age_at_men_stop
ref_data$catmenopause[ref_data$Age_at_men_stop <= 40] = 0
ref_data$catmenopause[ref_data$Age_at_men_stop > 40 & 
                        ref_data$Age_at_men_stop <= 45] = 1
ref_data$catmenopause[ref_data$Age_at_men_stop > 45 & 
                        ref_data$Age_at_men_stop <= 50] = 2
ref_data$catmenopause[ref_data$Age_at_men_stop > 50] = 3

test_data$catmenopause = test_data$Age_at_men_stop
test_data$catmenopause[test_data$Age_at_men_stop <= 40] = 0
test_data$catmenopause[test_data$Age_at_men_stop > 40 & 
                        test_data$Age_at_men_stop <= 45] = 1
test_data$catmenopause[test_data$Age_at_men_stop > 45 & 
                        test_data$Age_at_men_stop <= 50] = 2
test_data$catmenopause[test_data$Age_at_men_stop > 50] = 3

table(test_data$catmenopause)
sum(is.na(test_data$catmenopause))
```

## Reference dataset for premenopausal women
```{r}
ref_data_premeno = ref_data[ref_data$mentruation_status == 1,]
ref_data_premeno = ref_data_premeno[,c("age","ageatfirstfulltermpreg_cat",
                                       "interval_menarche_firstbirth_cat",
                                       "fulltermpreg_cat",
                                       "oc", "mean_brfeed_duration_cat",
                                       "totalabortion_cat",
                                       "totalmiscarriage_cat",
                                       "catht", "bmicat", "ratio")]

ref_data_premeno = na.omit(ref_data_premeno)

#Study entry age for the premenopausal women
age_premeno = ref_data_premeno$age

ref_data_premeno = ref_data_premeno[,!names(ref_data_premeno) %in% "age"]

dim(ref_data_premeno)


##Setting up the test data premenopausal women
test_data_premeno = test_data[test_data$mentruation_status == 1,]
test_data_premeno = test_data_premeno[,c("age","ageatfirstfulltermpreg_cat",
                                       "interval_menarche_firstbirth_cat",
                                       "fulltermpreg_cat",
                                       "oc", "mean_brfeed_duration_cat",
                                       "totalabortion_cat",
                                       "totalmiscarriage_cat",
                                       "catht", "bmicat", "ratio")]

#test_data_premeno = na.omit(test_data_premeno)

#Study entry age for the premenopausal women
age_premeno_test = test_data_premeno$age

test_data_premeno = test_data_premeno[,!names(test_data_premeno) %in% "age"]
```

## Model formula for premenopausal women
```{r}
model.formula.premeno = br ~ as.factor(ageatfirstfulltermpreg_cat) + 
  interval_menarche_firstbirth_cat + as.factor(fulltermpreg_cat) + 
  oc + as.factor(mean_brfeed_duration_cat) +
  as.factor(totalabortion_cat) + as.factor(totalmiscarriage_cat) + 
  as.factor(catht) + as.factor(bmicat) * as.factor(ratio)
```

## Model covariate information for premenopausal women
```{r}
v1 = list()
v1$name = "ageatfirstfulltermpreg_cat"
v1$type = "factor"
v1$levels = c("1","2","3","4","5")

v2 = list()
v2$name = "interval_menarche_firstbirth_cat"
v2$type = "continuous"

v3 = list()
v3$name = "fulltermpreg_cat"
v3$type = "factor"
v3$levels = c("1","2","3","4")

v4 = list()
v4$name = "oc"
v4$type = "continuous"

v5 = list()
v5$name = "mean_brfeed_duration_cat"
v5$type = "factor"
v5$levels = c("1","2","3","4")

v6 = list()
v6$name = "totalabortion_cat"
v6$type = "factor"
v6$levels = c("0","1","2")

v7 = list()
v7$name = "totalmiscarriage_cat"
v7$type = "factor"
v7$levels = c("0","1","2")

v8 = list()
v8$name = "catht"
v8$type = "factor"
v8$levels = c("0","1","2","3")

v9 = list()
v9$name = "bmicat"
v9$type = "factor"
v9$levels = c("1","2","3")

v10 = list()
v10$name = "ratio"
v10$type = "factor"
v10$levels = c("0","1","2")

cov.info.premeno = list(v1,v2,v3,v4,v5,v6,v7,v8,v9,v10)
```

##Model log relative risks for premenopausal women
```{r}
logRR_premeno = coef(model_premeno)[-c(1:10)]
```


## Absolute risk distribution for premenopausal women
```{r}
## 5-year risk
output_premeno_5y = computeAbsoluteRisk(model.formula = model.formula.premeno,
                                        model.cov.info = cov.info.premeno,
                                        model.snp.info = NULL,
                                        model.log.RR = logRR_premeno,
                                        model.ref.dataset = ref_data_premeno,
                                        model.disease.incidence.rates = bc_inci,
                                        model.competing.incidence.rates = mort_inci,
                                        model.bin.fh.name = NA,
                                        apply.age.start = age_premeno,
                                        apply.age.interval.length = 5,
                                        apply.cov.profile = ref_data_premeno,
                                        apply.snp.profile = NULL,
                                        return.lp = TRUE,
                                        return.refs.risk = TRUE)

output_premeno_test_5y = computeAbsoluteRisk(model.formula = model.formula.premeno,
                                             model.cov.info = cov.info.premeno,
                                             model.snp.info = NULL,
                                             model.log.RR = logRR_premeno,
                                             model.ref.dataset = ref_data_premeno,
                                             model.disease.incidence.rates = bc_inci,
                                             model.competing.incidence.rates = mort_inci,
                                             model.bin.fh.name = NA,
                                             apply.age.start = age_premeno_test,
                                             apply.age.interval.length = 5,
                                             apply.cov.profile = test_data_premeno,
                                             apply.snp.profile = NULL,
                                             return.lp = TRUE,
                                             return.refs.risk = TRUE)

summary(output_premeno_5y$risk * 100)
summary(output_premeno_test_5y$risk * 100)

#pdf("Five_year_AR_premenopausal.pdf")
plot(density(output_premeno_5y$risk * 100),
     xlim = c(0,4), ylim = c(0,4),
     main = "Premenopausal Women",
     xlab = "5-year absolute risk (%)", lwd = 3)
abline(v = output_premeno_test_5y$risk * 100, col = "red")
text(x = 3, y = 1, 
     labels = paste("Mean AR (%):", 
                    round(mean(output_premeno_5y$risk * 100),1)))
#dev.off()

## 10-year risk
output_premeno_10y = computeAbsoluteRisk(model.formula = model.formula.premeno,
                                         model.cov.info = cov.info.premeno,
                                         model.snp.info = NULL,
                                         model.log.RR = logRR_premeno,
                                         model.ref.dataset = ref_data_premeno,
                                         model.disease.incidence.rates = bc_inci,
                                         model.competing.incidence.rates = mort_inci,
                                         model.bin.fh.name = NA,
                                         apply.age.start = age_premeno,
                                         apply.age.interval.length = 10,
                                         apply.cov.profile = ref_data_premeno,
                                         apply.snp.profile = NULL,
                                         return.lp = TRUE,
                                         return.refs.risk = TRUE)

output_premeno_test_10y = computeAbsoluteRisk(model.formula = model.formula.premeno,
                                              model.cov.info = cov.info.premeno,
                                              model.snp.info = NULL,
                                              model.log.RR = logRR_premeno,
                                              model.ref.dataset = ref_data_premeno,
                                              model.disease.incidence.rates = bc_inci,
                                              model.competing.incidence.rates = mort_inci,
                                              model.bin.fh.name = NA,
                                              apply.age.start = age_premeno_test,
                                              apply.age.interval.length = 10,
                                              apply.cov.profile = test_data_premeno,
                                              apply.snp.profile = NULL,
                                              return.lp = TRUE,
                                              return.refs.risk = TRUE)

summary(output_premeno_10y$risk * 100)
summary(output_premeno_test_10y$risk * 100)

#pdf("Five_year_AR_premenopausal.pdf")
plot(density(output_premeno_10y$risk * 100),
     xlim = c(0,4), ylim = c(0,2),
     main = "Premenopausal Women",
     xlab = "10-year absolute risk (%)", lwd = 3)
abline(v = output_premeno_test_10y$risk * 100, col = "red")
text(x = 3, y = 1, 
     labels = paste("Mean AR (%):", 
                    round(mean(output_premeno_10y$risk * 100),1)))
#dev.off()
```


## Reference dataset for postmenopausal women
```{r}
ref_data_postmeno = ref_data[ref_data$mentruation_status == 2,]
ref_data_postmeno = ref_data_postmeno[,c("age","ageatfirstfulltermpreg_cat",
                                         "interval_menarche_firstbirth_cat",
                                         "fulltermpreg_cat", "catmenopause",
                                         "oc", "mean_brfeed_duration_cat",
                                         "totalabortion_cat",
                                         "totalmiscarriage_cat",
                                         "catht", "bmicat", "ratio")]

ref_data_postmeno = na.omit(ref_data_postmeno)

#Study entry age for postmenopausal women
age_postmeno = ref_data_postmeno$age

ref_data_postmeno = ref_data_postmeno[,!names(ref_data_postmeno) %in% "age"]


test_data_postmeno = test_data[test_data$mentruation_status == 2,]
test_data_postmeno = test_data_postmeno[,c("age","ageatfirstfulltermpreg_cat",
                                         "interval_menarche_firstbirth_cat",
                                         "fulltermpreg_cat", "catmenopause",
                                         "oc", "mean_brfeed_duration_cat",
                                         "totalabortion_cat",
                                         "totalmiscarriage_cat",
                                         "catht", "bmicat", "ratio")]

#test_data_postmeno = na.omit(test_data_postmeno)

#Study entry age for postmenopausal women
age_postmeno_test = test_data_postmeno$age

test_data_postmeno = test_data_postmeno[,!names(test_data_postmeno) %in% "age"]
```

## Model formula for postmenopausal women
```{r}
model.formula.postmeno = br ~ as.factor(ageatfirstfulltermpreg_cat) + 
  interval_menarche_firstbirth_cat + as.factor(fulltermpreg_cat) + 
  as.factor(catmenopause) + oc + as.factor(mean_brfeed_duration_cat) + 
  as.factor(totalabortion_cat) + as.factor(totalmiscarriage_cat) + 
  as.factor(catht) + as.factor(bmicat) * as.factor(ratio)
```

## Model covariate information for postmenopausal women
```{r}
v1 = list()
v1$name = "ageatfirstfulltermpreg_cat"
v1$type = "factor"
v1$levels = c("1","2","3","4","5")

v2 = list()
v2$name = "interval_menarche_firstbirth_cat"
v2$type = "continuous"

v3 = list()
v3$name = "fulltermpreg_cat"
v3$type = "factor"
v3$levels = c("1","2","3","4")


v4 = list()
v4$name = "catmenopause"
v4$type = "factor"
v4$levels = c("0","1","2","3")

v5 = list()
v5$name = "oc"
v5$type = "continuous"

v6 = list()
v6$name = "mean_brfeed_duration_cat"
v6$type = "factor"
v6$levels = c("1","2","3","4")

v7 = list()
v7$name = "totalabortion_cat"
v7$type = "factor"
v7$levels = c("0","1","2")

v8 = list()
v8$name = "totalmiscarriage_cat"
v8$type = "factor"
v8$levels = c("0","1","2")

v9 = list()
v9$name = "catht"
v9$type = "factor"
v9$levels = c("0","1","2","3")

v10 = list()
v10$name = "bmicat"
v10$type = "factor"
v10$levels = c("1","2","3")

v11 = list()
v11$name = "ratio"
v11$type = "factor"
v11$levels = c("0","1","2")


cov.info.postmeno = list(v1,v2,v3,v4,v5,v6,v7,v8,v9,v10,v11)
```

## Model log relative risks for postmenopausal women
```{r}
logRR_postmeno = coef(model_postmeno)[-c(1:10)]
```

## Absolute risk distribution for postmenopausal women
```{r}
## 5-year risk
output_postmeno_5y = computeAbsoluteRisk(model.formula = model.formula.postmeno,
                                         model.cov.info = cov.info.postmeno,
                                         model.snp.info = NULL,
                                         model.log.RR = logRR_postmeno,
                                         model.ref.dataset = ref_data_postmeno,
                                         model.disease.incidence.rates = bc_inci,
                                         model.competing.incidence.rates = mort_inci,
                                         model.bin.fh.name = NA,
                                         apply.age.start = age_postmeno,
                                         apply.age.interval.length = 5,
                                         apply.cov.profile = ref_data_postmeno,
                                         apply.snp.profile = NULL,
                                         return.lp = TRUE,
                                         return.refs.risk = TRUE)

output_postmeno_test_5y = computeAbsoluteRisk(model.formula = model.formula.postmeno,
                                              model.cov.info = cov.info.postmeno,
                                              model.snp.info = NULL,
                                              model.log.RR = logRR_postmeno,
                                              model.ref.dataset = ref_data_postmeno,
                                              model.disease.incidence.rates = bc_inci,
                                              model.competing.incidence.rates = mort_inci,
                                              model.bin.fh.name = NA,
                                              apply.age.start = age_postmeno_test,
                                              apply.age.interval.length = 5,
                                              apply.cov.profile = test_data_postmeno,
                                              apply.snp.profile = NULL,
                                              return.lp = TRUE,
                                              return.refs.risk = TRUE)

summary(output_postmeno_5y$risk * 100)
summary(output_postmeno_test_5y$risk * 100)

#pdf("Five_year_AR_postmenopausal.pdf")
plot(density(output_postmeno_5y$risk * 100),
     xlim = c(0,4), ylim = c(0,4),
     main = "Postmenopausal Women",
     xlab = "5-year absolute risk (%)", lwd = 3)
abline(v = output_postmeno_test_5y$risk * 100, col = "red")
text(x = 3, y = 1, 
     labels = paste("Mean AR (%):", 
                    round(mean(output_postmeno_5y$risk * 100),1)))
#dev.off()

##10-year risk
output_postmeno_10y = computeAbsoluteRisk(model.formula = model.formula.postmeno,
                                          model.cov.info = cov.info.postmeno,
                                          model.snp.info = NULL,
                                          model.log.RR = logRR_postmeno,
                                          model.ref.dataset = ref_data_postmeno,
                                          model.disease.incidence.rates = bc_inci,
                                          model.competing.incidence.rates = mort_inci,
                                          model.bin.fh.name = NA,
                                          apply.age.start = age_postmeno,
                                          apply.age.interval.length = 10,
                                          apply.cov.profile = ref_data_postmeno,
                                          apply.snp.profile = NULL,
                                          return.lp = TRUE,
                                          return.refs.risk = TRUE)

output_postmeno_test_10y = computeAbsoluteRisk(model.formula = model.formula.postmeno,
                                               model.cov.info = cov.info.postmeno,
                                               model.snp.info = NULL,
                                               model.log.RR = logRR_postmeno,
                                               model.ref.dataset = ref_data_postmeno,
                                               model.disease.incidence.rates = bc_inci,
                                               model.competing.incidence.rates = mort_inci,
                                               model.bin.fh.name = NA,
                                               apply.age.start = age_postmeno_test,
                                               apply.age.interval.length = 10,
                                               apply.cov.profile = test_data_postmeno,
                                               apply.snp.profile = NULL,
                                               return.lp = TRUE,
                                               return.refs.risk = TRUE)

summary(output_postmeno_10y$risk * 100)
summary(output_postmeno_test_10y$risk * 100)

#pdf("Five_year_AR_postmenopausal.pdf")
plot(density(output_postmeno_10y$risk * 100),
     xlim = c(0,4), ylim = c(0,2),
     main = "Postmenopausal Women",
     xlab = "10-year absolute risk (%)", lwd = 3)
abline(v = output_postmeno_test_10y$risk * 100, col = "red")
text(x = 3.5, y = 1, 
     labels = paste("Mean AR (%):", 
                    round(mean(output_postmeno_10y$risk * 100),1)))
#dev.off()

```

## Risk projections in the Indian population

Load the census data
```{r}
setwd("/Users/anaghchattopadhyay/Documents/Nilanjan\ Project/TMC_BCProject/Data")

pop_data = read.xlsx("Indian population numbers by Age (Census-2011).xlsx")
names(pop_data) = c("Age","Population")
pop_data = pop_data[-1,]


pop_data$Age = as.numeric(pop_data$Age)
pop_data$Population = as.numeric(pop_data$Population)

pop_data = na.omit(pop_data)
```

Age at menopause category definitions
0: <= 40 years
1: 41-45 years
2: 46-50 years
3: >= 51 years

Look at the distribution of age at menopause to determine the age cutoff
```{r}
prop.table(table(ref_data_postmeno$catmenopause))
```
Median age at menopause roughly 45 years. Premenopausal group defined as 25-45 years and postmenopausal group defined as 45-68 years

Population size in the premenopausal group:
```{r}

pop_total_premeno = sum(pop_data$Population[pop_data$Age >= 25 & 
                                              pop_data$Age < 45], na.rm = TRUE)
pop_total_premeno
```

5-year risk projections

Approximate number of future cases (within 5 years) in the premenopausal group:
```{r}
cases_total_premeno = round(pop_total_premeno * round(mean(output_premeno_5y$risk),3),0)
cases_total_premeno
```


Percentage of premenopausal women above 1\% 5-year absolute risk
```{r}
round(mean(output_premeno_5y$risk >= 0.01) * 100,1)
```


Number of premenopausal women above 1\% 5-year absolute risk
```{r}
pop_total_above_0.01AR_premeno = round(round(mean(output_premeno_5y$risk >= 0.01),3) * 
                                         pop_total_premeno, 0)
pop_total_above_0.01AR_premeno
```



Number of future cases in the group above 1\% 5-year absolute risk
```{r}
cases_total_above_0.01AR_premeno = 
  round(round(mean(output_premeno_5y$risk[output_premeno_5y$risk >= 0.01]),3) * 
          pop_total_above_0.01AR_premeno,0)
cases_total_above_0.01AR_premeno
```
Percentage of future cases in the group above 1\% 5-year absolute risk
```{r}
round((cases_total_above_0.01AR_premeno/cases_total_premeno)*100,1)
```

Population size in the postmenopausal group:

```{r}
pop_total_postmeno = sum(pop_data$Population[pop_data$Age >= 45 & 
                                              pop_data$Age <= 68], na.rm = TRUE)
pop_total_postmeno
```

Approximate number of future cases (within 5 years) in the postmenopausal group:
```{r}
cases_total_postmeno = round(pop_total_postmeno * round(mean(output_postmeno_5y$risk),3),0)
cases_total_postmeno
```

Percentage of postmenopausal women above 1\% 5-year absolute risk
```{r}
round(mean(output_postmeno_5y$risk >= 0.01) * 100,1)
```

Number of postmenopausal women above 1\% 5-year absolute risk
```{r}
pop_total_above_0.01AR_postmeno = round(round(mean(output_postmeno_5y$risk >= 0.01),3) * 
                                          pop_total_postmeno, 0)
pop_total_above_0.01AR_postmeno
```

Number of future cases in the group above 1\% 5-year absolute risk
```{r}
cases_total_above_0.01AR_postmeno = 
  round(round(mean(output_postmeno_5y$risk[output_postmeno_5y$risk >= 0.01]),3) * 
          pop_total_above_0.01AR_postmeno,0)
cases_total_above_0.01AR_postmeno
```
Percentage of future cases in the group above 1\% 5-year absolute risk
```{r}
round((cases_total_above_0.01AR_postmeno/cases_total_postmeno)*100,1)
```



10-year risk projections

Approximate number of future cases (within 10 years) in the premenopausal group:
```{r}
cases_total_premeno = round(pop_total_premeno * round(mean(output_premeno_10y$risk),3),0)
cases_total_premeno
```


Percentage of premenopausal women above 2\% 10-year absolute risk
```{r}
round(mean(output_premeno_10y$risk >= 0.02) * 100,1)
```


Number of premenopausal women above 2\% 10-year absolute risk
```{r}
pop_total_above_0.02AR_premeno = round(round(mean(output_premeno_10y$risk >= 0.02),3) * 
                                         pop_total_premeno, 0)
pop_total_above_0.02AR_premeno
```

Number of future cases in the group above 2\% 10-year absolute risk
```{r}
cases_total_above_0.02AR_premeno = 
  round(round(mean(output_premeno_10y$risk[output_premeno_10y$risk >= 0.02]),3) * 
          pop_total_above_0.02AR_premeno,0)
cases_total_above_0.02AR_premeno
```
Percentage of future cases in the group above 2\% 10-year absolute risk
```{r}
round((cases_total_above_0.02AR_premeno/cases_total_premeno)*100,1)
```

Population size in the postmenopausal group:

```{r}
pop_total_postmeno = sum(pop_data$Population[pop_data$Age >= 45 & 
                                              pop_data$Age <= 68], na.rm = TRUE)
pop_total_postmeno
```

Approximate number of future cases (within 10 years) in the postmenopausal group:
```{r}
cases_total_postmeno = round(pop_total_postmeno * round(mean(output_postmeno_10y$risk),3),0)
cases_total_postmeno
```

Percentage of postmenopausal women above 2\% 10-year absolute risk
```{r}
round(mean(output_postmeno_10y$risk >= 0.02) * 100,1)
```

Number of postmenopausal women above 2\% 10-year absolute risk
```{r}
pop_total_above_0.02AR_postmeno = round(round(mean(output_postmeno_10y$risk >= 0.02),3) * 
                                          pop_total_postmeno, 0)
pop_total_above_0.02AR_postmeno
```

Number of future cases in the group above 2\% 10-year absolute risk
```{r}
cases_total_above_0.02AR_postmeno = 
  round(round(mean(output_postmeno_10y$risk[output_postmeno_10y$risk >= 0.02]),3) * 
          pop_total_above_0.02AR_postmeno,0)
cases_total_above_0.02AR_postmeno
```
Percentage of future cases in the group above 1\% 5-year absolute risk
```{r}
round((cases_total_above_0.02AR_postmeno/cases_total_postmeno)*100,1)
```


